### 2014-01-27 notes/notes.tex
Adding notes

\documentclass[12pt]{article}
\usepackage{parskip}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}

\title{Test statistic selection for randomization inference under interference}
\author{Peter Aronow and Jake Bowers}
\date{\today}

\begin{document}
\maketitle

%% 4 pages

 \citep{bowers2013sutva} followed Rosenbaum by recommending effect increasing
 statistic.

 controls Type I error, but power can be very poor for these types of models.
 We explain why and propose a refinement. Multivariate dependence of
 parameters and outcome space. A weighted E-statistic.

Fisher's sharp null hypothesis of no effects combined with his
randomization-based hypothesis testing framework provided a very clear and
clean vision of what it means to do statistical inference \citep[Chap
2]{fisher:1935}. In recent years, a number of scholars have realized that this
clarity also implied great flexibility in regard to the scientific questions
that one might engage: for example, one could use the sharp null itself to
detect interference among units \citep{aronow-general}, and sharp hypotheses
about effects (in contrast to hypotheses about no effects) have been developed
to assess models of interference \citep{bowers2013sutva}, [ include Rosenbaum
2007, Bowers and Hansen 2009, and a bunch of other stuff by Rosenbaum ].

One problem that has become apparent, as scholars have extended Fisher's basic
ideas, is that as the models of effects have become more complex, the power of
the testing procedure itself has become more important. A simple test of the
sharp null can be executed with power that varies as a function of the design
of the study (proportion treated, blocking structure, etc..), characteristics
of the outcome, and the way that a test statistic summarizes the outcome. Yet,
when the models are complex,  \citet{bowers2013sutva} showed that sometimes a
test will have (1) no power to address the model at all (a function, in part
of the model itself --- for example ...) such that all hypothesized parameters
would receive the same implausibility assessment; (2) or might reject all such
hypothesized parameters. Thus, although in theory one may assess sharp
hypotheses about multiparameter settings, in practice one may not learn much
from such tests --- or, in the absence of simulation studies assessing the
model and test statistic, one may be mislead.

In this paper we use a very simple causal model and a simple test statistic to
first show how the combination of model and test statistic is a crucial choice
--- and how more attention ought to be paid to both the model and the test
statistic.

Then we suggest one test statistic that we think should have good properties
when the number of parameters in a model is low (i.e. $k << n$). And we should
this to be the case --- that the robust F-stat from a linear model is a better
way to go than the KS-test and certainly better than any single measure of
distributional difference.

Finally, we highlight the similarities between testing the sharp hypothesis
and the weak hypothesis --- which in turn allows even more flexibility and
perhaps better performing testing procedures.

Then we suggest a few alternative avenues to pursue.

Some models are well powered for some designs and not other designs.

Using same simulation study using same data from PA study. Better than KS
test.

The F-test is model consistent. By design it controls Type I error, but if the
model is correct, it has some optimality criterion. If you had control
variables, for example, then it would be consistent against mean differences.

Under the model, model consistent.

\bibliographystyle{apsr}
\bibliography{/Users/jwbowers/Documents/PROJECTS/BIB/big}

\end{document}

### 2014-09-11 paper/comment-on-bfp.tex
Moving files into place

\documentclass[12pt]{article}

% Graphics
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{psfrag}
\usepackage{wrapfig}
\usepackage[all]{xy}

% Author
\usepackage[blocks]{authblk}
\renewcommand\Affilfont{\small}
\setlength{\affilsep}{0em}

% Document formatting
\usepackage{fullpage}
\pagestyle{plain}
\usepackage{setspace}
%\usepackage{mathptmx}
\usepackage{url}
\usepackage{times}

% Bibliography formatting
\bibliographystyle{chicago}
\usepackage{natbib}

% Equation formatting
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{bbm}
\usepackage{array}

\newtheorem{lemma}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{thm}{Theorem}

% Shortcuts
\newcommand{\nn}{\nonumber}

% Define new characters
\def\Var{{\rm Var}\,}
\def\E{{\rm E}\,}
\def\arg{{\rm arg}\,}
\def\Cov{{\rm Cov}\,}
\def\N{{\rm N}\,}
%--------------------------------------------------------------------------
% Math boldface shortcuts, etc. ----------------------------------
%--------------------------------------------------------------------------
\newcommand{\X}{\mathbf{X}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\Z}{\mathbf{Z}}
\newcommand{\I}{{\rm I}}
\newcommand{\V}{\mathbf{V}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\R}{\mathbf{R}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\F}{{\rm F}}
\renewcommand{\O}{\mathbf{O}}
\renewcommand{\H}{\mathbf{H}}
\newcommand{\g}{\mathbf{g}}
\renewcommand{\r}{\mathbf{r}}
\newcommand{\s}{\mathbf{s}}
\newcommand{\uu}{\mathbf{u}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\Beta}{\boldsymbol{\beta}}
\newcommand{\btheta}{\boldsymbol{\theta}}
\newcommand{\bgamma}{\boldsymbol{\gamma}}
\newcommand{\bpi}{\boldsymbol{\pi}}
\newcommand{\arrowp}{\stackrel{p}{\rightarrow}}
\newcommand{\0}{\mathbf{0}}
\newcommand{\one}{\mathbf{1}}
\newcommand{\bP}{\mathbf{P}}
\newcommand{\plim}{{\rm plim}}

\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}

\renewcommand\refname{Additional references}

\begin{document}

\sloppy

%\addtolength{\oddsidemargin}{-.075in}
%\addtolength{\evensidemargin}{-.075in}
%\addtolength{\textwidth}{.15in}
\addtolength{\topmargin}{-.025in}
\addtolength{\textheight}{.05in}

\author{Peter M. Aronow\thanks{Peter M. Aronow is Assistant Professor, Department of Political Science, Yale University, New Haven, CT (Email: peter.aronow@yale.edu).}}

% TITLE
\title{Comment: Reasoning about Interference Between Units}
\maketitle

%\doublespace

Bowers, Fredrickson and Panagopoulos (BFP) should be commended for an illuminating paper, one that highlights how randomization provides a basis for inference, even when considering models that include interference between units. BFP articulates and convincingly advocates for a principled, Fisherian approach to inference.

The goal of this comment is to complement BFP by discussing limitations, extensions and an alternative mode of inference. Using a simple running example, 
this comment makes three points:

\begin{enumerate}
\item The power of the procedures proposed by BFP depends on the choice of test statistic. The class of ``effect increasing" test statistics used by BFP -- which focuses on distributional differences between the hypothesized potential outcomes for units assigned to treatment and those assigned to control -- may be suboptimal for models that include interference. Although type I error rates are controlled well, type II error rates are not. I suggest alternative test statistics that may utilize the data to a fuller extent. 
\item The low-dimensional models considered by BFP may yield misleading conclusions about causal effects. In an experiment with heterogeneous treatment effects but no spillover, researchers may nevertheless reject all {\it sharp null} hypotheses that there are no spillover effects (conditional on a given causal model). The proposed method cannot test hypotheses about spillover without also jointly hypothesizing that treatment effect heterogeneity is completely modeled.
\item Estimation and testing are possible under semiparametric models that do not presuppose low-dimensional models of causal effects. These approaches obviate both of the above concerns but introduce asymptotic approximations to facilitate hypothesis testing. I briefly outline one such approach.%, and refer the reader to readings on the topic.
\end{enumerate}
None of these points undermine the primary contribution of BFP -- an %technically correct and 
insightful piece that opens new avenues of research. Rather, contextualizing the piece permits greater appreciation of its contribution and highlights open areas of inquiry.
%Rather, contextualizing the piece emphasizes the open areas of inquiry.

\section*{Running example}

Consider the following setting: a randomized experiment is performed on a finite population $U$ consisting of $N$ individuals living in $N/2$ two-person households. Index each individual $ij$, where $i \in 1,...,(N/2)$ households and $j \in \{-1,1\}$, so that individual $i1$ is the first person in household $i$ and individual $i(-1)$ is the second individual in household $i$. Assume simple randomization of individuals to treatment: a fair coin is independently flipped to determine the treatment condition of each individual $ij$. Associated with each individual $ij$ is an outcome $y_{ij}$ and a treatment condition $Z_{ij}$. Further assume the possibility of partial interference (Sobel, 2006)%cite
, so that treatment effects may spillover within households, but not across households.

% Coinflip randomization

%The randomization proceeds as follows: $m$ of $N$ households are chosen with equal probability to treatment eligibility. If household $i$ is chosen for eligibility, then one individual $i1$ or $i2$ is randomly selected to receive treatment. Denote the treatment assignment vector $\Z = \{Z_{11},...,Z_{N1},{Z_{1(-1)},...,Z_{N(-1)}}\}$, where $Z_{ij} \in \{0,1\}$ indicates if unit $ij$ was assigned to treatment. Thus all individuals have equal probability $p \equiv (\frac{m}{2N})$ of receiving the treatment $Z_{ij}=1$, and no more than one individual in each household receives treatment.

\section*{Test statistic selection}

BFP recommends choosing ``effect increasing" test statistics ``that will be small when the treated and control distributions in the adjusted data ... are similar, and large when the distributions diverge." However, test statistics  that only compare distributional features of hypothesized potential outcomes between individuals assigned to treatment and those assigned to control fail to fully exploit the data at hand. The problem is that these test statistics are  not necessarily increasing in spillover effects.

Consider the following causal model:
\begin{equation}\label{causal}
y_{ij}(\Z) = y_{ij}(\0) + \tau_1 Z_{ij} + \tau_2 Z_{i(-j)},
\end{equation}
where $y_{ij}(\Z)$ is unit $ij$'s potential outcome under the treatment vector $\Z$, $Z_{ij} \in \{0,1\}$ indicates if unit $ij$ was assigned to treatment in $\Z$, and $Z_{i(-j)}$ indicates if the housemate of unit $ij$ was assigned to treatment, and $y_{ij}(\0)$ is the potential outcome for individual $ij$ when no one is assigned to treatment (known as a ``uniformity trial").  Before proceeding, it is helpful to establish some additional notation. Denote $\Z \equiv (Z_{11}, ..., Z_{(N/2){(-1)}})^T$, $\Z^I \equiv (Z_{1(-1)},...,Z_{(N/2)1})^T$ (reversing the within-household indices), $\0 \equiv (0,...,0)^T$, and $\one \equiv (1,...,1)^T$. Define the hypothesized $y_{ij}(\0)$ that would be obtained under $\Z$, $\widetilde y_{ij}(\0,\Z) = y_{ij}(\Z) -  \widetilde \tau_1 Z_{ij} + \widetilde \tau_2 Z_{i(-j)}$, where $\widetilde \tau_1$ and $\widetilde \tau_2$ are the hypothesized values of $\tau_1$ and $\tau_2$. Finally, define the mean of some feature $X_{ij}$ over all units $ij$ such that $Z_{ij} = t$, $\mu_t(X_{ij},\Z) \equiv  {\sum_{ij \in U} \I[Z_{ij} = t] X_{ij}} / {\sum_{ij \in U} \I[Z_{ij} = t]}$. %Finally, define a test statistic $\T(\widetilde y_{ij}(\0,\Z), \Z')$.

%A simple intuition for this result follows. Under a radically incorrect hypothesis, such a test amounts to a test of the difference in proportions of treated roommates for treated and control individuals. Given simple randomization, there is no difference in the expected proportions of treated roommates. Furthermore, as $N$ grows large, the randomization variance of the difference in proportions converges at the same rate as does the hypothesized randomization variance. It follows that a test based on an effect increasing statistic will fail to reject the radically incorrect hypothesis at proper rates.

%In fact, this second result does not depend on example in the setting.
% I will show that a test based on the absolute difference-in-means, coupled with a linear model of the form \eqref{causal}, may not be able to rule any hypothesized value of $\tau_2$ at any significance level $\alpha < 1$.  Second, I will show that, given the running example, a test based on any effect increasing statistic will be inconsistent against a radically incorrect hypothesis about $\tau_2$.
%(In fact, this first result does not depend on the running example, but is only relevant for the difference-in-means as a test statistic.)

\subsection*{Results for effect increasing test statistics}

I demonstrate two results, each showing how effect increasing test statistics, as defined by BFP, may provide little to no information about the spillover parameter $\tau_2$. 
Let us first consider the case when the test statistic is the absolute mean difference between hypothesized uniformity trial outcomes for treated and control individuals: 
$\T(\widetilde y_{ij}(\0,\Z), \Z') = 
\left\vert \mu_1 ( \widetilde y_{ij}(\0,\Z), \Z' ) - \mu_0 ( \widetilde y_{ij}(\0,\Z), \Z' )
\right\vert.
$
Fix a realization of $\Z = \z$, and further assume noncollinearity: $ \z \notin \{\z^I, (1-\z^I)\}$. We can see that, for any hypothesized $\widetilde\tau_2$, there always exists a $\widetilde\tau_1$ such that the $p$-value is 1. Expanding the test statistic, 
$\T(\widetilde y_{ij}(\0,\z), \z) = 
\left\vert \mu_1 (y_{ij}(\z) - \widetilde\tau_1 - \tau_2 z_{i(-j)}, \z) - \mu_0(\widetilde y_{ij}(\z) - \widetilde\tau_2 z_{i(-j)} , \z)
\right\vert
=
\left\vert \mu_1 (y_{ij}(\z) - \widetilde\tau_2 z_{i(-j)}, \z)- \mu_0 (\widetilde y_{ij}(\z) - \widetilde\tau_2 z_{i(-j)}, \z)- \widetilde\tau_1 
\right\vert.
$ 
Thus, regardless of the value of $\widetilde\tau_2$, we can always find a $\widetilde\tau_1$ that sets $\T(\widetilde y_{ij}(\0,\z), \z) = 0$. Given that $\T(\widetilde y_{ij}(\0,\z), \Z')$ can only take values in $\mathbb{R}^+$, then $p=1$, as any other possible randomization $\Z'$ will yield a value of $\T(\widetilde y_{ij}(\0,\z), \Z') \geq 0$. It follows that, given this test, we cannot rule out any value of $\tau_2$ as being implausible.\footnote{In fact, this result does not depend on example in the setting, and holds for any randomization scheme coupled with a linear causal model of the form \eqref{causal}.}

Second, I show that, given the running example, a test based on any effect increasing statistic is inconsistent against a radically incorrect hypothesis about $\tau_2$. Let us assume that $\tau_1$ is properly hypothesized at its true value, and that the hypothesized value of $\tau_2$ ($\widetilde \tau_2$) is radically incorrect: 
$\vert \widetilde\tau_2 - \tau_2 \vert \gg  \max(y_{ij}(\0)) - \min(y_{ij}(\0))$. Denote $\beta \equiv \tau_2 - \widetilde\tau_2$, and assume $\vert \beta \vert < \infty$.
Under the assumption that $\tau_2 = \widetilde \tau_2$:
$\widetilde y_{ij}(\0,\Z) = y_{ij}(\0) + \beta Z_{i(-j)}$. 
Assume a sequence of nested finite populations $\{U_k\} = U_1, U_2,...$, growing in $k$ households, so that $U_k$ consists of $N_k = 2k$ individuals, each with an associated random assignment $\Z_k$ \citep{isakifuller}. (For notational parsimony, I drop explicit dependence on the subscript $k$.)  Given an incorrect hypothesis, a necessary, but not sufficient, condition for a test's consistency is that $\lim_{k \rightarrow \infty} \Pr(p > \alpha) < \epsilon$, for any $\alpha,\epsilon > 0$.

Let an effect increasing test statistic $\T(\widetilde y_{ij}(\0,\Z),\Z')$ be some measure of the absolute divergence in empirical distributions of $\widetilde y_{ij}(\0,\Z)$ for treated and control individuals under the treatment vector $\Z'$. 
Since distributional differences in $y_{ij}(\0)$ are small relative to shifts induced by differences in proportions of treated housemates,  $\T(\widetilde y_{ij}(\0,\Z),\Z')$ will be increasing in 
 $ \left\vert \mu_1(Z_{i(-j)},\Z') -   \mu_0(Z_{i(-j)},\Z') \right\vert$. %, 
For a realization $\Z = \z$, 
the hypothesized randomization distribution of
 $\mu_1(z_{i(-j)},\Z') - \mu_0(z_{i(-j)},\Z')$ is asymptotically normal with mean $0$ and variance $V_{x(\z)} = 4 x(\z) \left[1-x(\z)\right] / N$, where $x(\z) \equiv \sum_{ij\in U}z _{ij} /N$,
if $0 < x(\z) <  1$. % (by the Lindeberg CLT, Taylor linearization and Slutsky's theorem). 
 Thus the $p$-value associated with randomization $\z$, $p = 2 \Phi(-|\mu_1(z_{i(-j)},\z) - \mu_0(z_{i(-j)},\z)|/V_{x(\z)}^{1/2})$. It follows that  
 $|\mu_1(z_{i(-j)},\z) - \mu_0(z_{i(-j)},\z)\vert < -\Phi^{-1}(\alpha/2) V_{x(\z)}^{1/2} \Leftrightarrow p > \alpha$. 
By similar calculations at the household level, the %(actual, not hypothesized) 
randomization distribution of 
$\mu_1(Z_{i(-j)},\Z) - \mu_0(Z_{i(-j)},\Z)$ is asymptotically mean $0$ with variance $2 V_{1/2} =  2/N$.\footnote{This result follows from reframing the problem as randomly assigning households into 4 conditions \{00,01,10,11\} with equal probability, then linearizing $\mu_1(Z_{i(-j)},\Z) - \mu_0(Z_{i(-j)},\Z)$.} By the LLN %,  $\plim_{k \rightarrow \infty} x = 1/2$,
%and so, by 
%the 
and continuous mapping theorem, $\lim_{k \rightarrow \infty} \Pr(\vert NV_{x(\Z)} - NV_{1/2}\vert > \epsilon_3) = 0$.
  Then, by Chebyshev's inequality, $\lim_{k \rightarrow \infty}  \Pr(p > \alpha) > 1- 2\left[\Phi^{-1}(\alpha/2)\right]^{-2}$, 
  %$\epsilon = 1- 2\left[\Phi^{-1}(\alpha/2)\right]^{-2}$ proportion of possible $\Z \in \Omega$ have $p > \alpha$, 
  for any $0 < \alpha < 2\Phi(-\sqrt{2})$.  Tests based on effect increasing statistics are therefore not generally consistent.

\subsection*{Alternative test statistics}

There are usually better ways to evaluate the fit of a model than comparing two marginal distributions. One simple test statistic would be the sum-of-squared-residuals (SSR) from a least squares regression of $\widetilde y_{ij}(\0,\Z)$ on $Z_{ij}$ and $Z_{i(-j)}$. Since the model in \eqref{causal} is linear, $\T(\widetilde y_{ij}(\0,\Z), \Z) = \sum_{ij \in U} \left[ \widetilde \Y(\0) - (\X^T\X)^{-1}\X^T \widetilde \Y(\0) \right]^2$, where 
$$
\widetilde \Y(\0) \equiv \left( \begin{array}{c} \widetilde y_{11}(0) \\ \vdots \\ y_{(N/2)(-1)}(0) \end{array} \right), 
\X \equiv  \left( \begin{array}{ccc} 1 & Z_{11} & Z_{1(-1)} \\ \vdots & \vdots & \vdots  \\ 1 & Z_{(N/2)(-1)} & Z_{(N/2)1} \end{array} \right) = ({\bf 1} \; \Z \; \Z^I).
$$   
Assume now that either the hypothesized $\tau_1$ or $\tau_2$ is radically incorrect: i.e., $\vert \widetilde\tau_1 - \tau_1 \vert$ or $\vert \widetilde\tau_2 - \tau_2 \vert \gg \max(y_{ij}(\0)) - \min(y_{ij}(\0))$.
%When either $\tau_1$ or $\tau_2$ is radically incorrect (), 
%
 %$\tau_1$ is correctly hypothesized and $\tau_2$ is radically incorrect, 
Fixing $\Z = \z$, for any randomization $\Z' \notin \{\z, (1-\z), \z^I, (1-\z^I),
 \0, \one \}$,
$\T(\widetilde y_{ij}(\0,\z),\Z') > \T(\widetilde y_{ij}(\0,\z),\z)$. Assuming noncollinearity,
%$\Z \notin \{\Z^I, (1-\Z^I)\}$,
%, (0,...,0), (1,...,1)\}$)
  $p \leq 6  \left( 2^{-N} \right)$, rejecting the hypothesis with $\alpha = 0.05$ for any $N \geq 8$. %Consistency follows straightforwardly.

In the example above, use of SSR as a test statistic has maximal power to reject radically incorrect hypotheses, whereas the test statistics proposed by BFP are inconsistent. But SSR from an ordinary least squares regression is not always appropriate: for example, when the probability of exposure to spillover is heterogeneous across individuals, we may wish to apply inverse probability weights so as to ensure representative samples of potential outcomes. %This suggests a conjecture: that the $SSR$ from an {\it inverse-probability-weighted} least squares regression is more generally a sensible test statistic for models that include interference.
(The nature of these weights will be discussed when discussing alternative modes of inference.)
Additionally, when nonlinear deviations from model predictions are of concern, a weighted variant of the Brownian distance covariance \citep{rizzo} or other {\it E}-statistic may be more sensible than SSR.

Test statistic selection is not trivial and BFP's contribution has effectively opened a new line of research. Via counterexample, we can see that effect increasing test statistics are not always appropriate for assessing hypotheses about interference. BFP is correct in the assertion that, regardless of the choice of test statistic selection, a set of implausible hypotheses is identified by the procedure. But we should not be led to believe that, for any given test statistic, that some hypotheses are more plausible than others. Such inferences -- comparing hypotheses -- may depend on the test statistic used, and not necessarily reflect the plausibility of the model at hand. This point will gain importance as we now consider the limitations of sharp null hypotheses.

\section*{Sharp null hypotheses}

The Fisherian mode of inference presented here is useful for testing sharp null hypotheses, wherein all potential outcomes -- or at a minimum those under a uniformity trial -- can be exactly hypothesized. 
%The most common sharp null hypothesis is one of no effect whatsoever: $H_0: y_i(t) = y_i(t'), \forall i \in 1,...,N; t \in \Omega$. 
Note what this implies though: we cannot ``reason about interference" without also reasoning about the exact form of the individualistic effects of the treatment.

Consider the following causal model:
$
y_i(\Z) =  y_{ij}(\0) + \tau_{1ij} Z_{ij} + \tau_2 Z_{i(-j)},
$
such that $\exists ij, ij'$ where $\tau_{1ij} \neq \tau_{1ij'}$.
The only difference between this causal model and the one in \eqref{causal} is that now treatment effects are heterogeneous. As a result, though, the hypothesis $H_0: y_i(\Z) =  y_{ij}(\0) + \tau_{1} Z_{ij} + \tau_2 Z_{i(-j)}$ is false for  any values of $\tau_{1}, \tau_{2}$. An consistent test against $H_0$ would therefore always reject with large $N$. We could entertain any number of sharp null hypotheses: including modeled heterogeneity or dilated effects. But unless these models were exactly correct, a consistent test would reject all hypotheses of the form $H_0$. In many ways, this is a {\it feature} of BFP's analytic framework: entire classes of causal models may be rejected as implausible, regardless of the values of the hypothesized parameters. But it can also be a limitation. BFP's method cannot test any hypothesis about the spillover effects of a treatment without also jointly hypothesizing that all sources of treatment effect heterogeneity are modeled.

These issues are compounded when our tests are not consistent against incorrect hypotheses. For example, given modest regularity conditions, when $\T(\widetilde y_{ij}(\0,\Z), \Z') = 
\left\vert \mu_1 ( \widetilde y_{ij}(\0,\Z), \Z' ) - \mu_0 ( \widetilde y_{ij}(\0,\Z), \Z' )
\right\vert
$, we can always find a set of hypotheses are are prima facie plausible and we can reject other hypotheses at some level $\alpha$. It is then imperative for the researcher to carefully interpret what rejection of a sharp null hypothesis implies. BFP is careful to distinguish between the failure to reject and  the acceptance of a null hypothesis. This point should not be neglected by researchers. Although some hypotheses can be rejected as implausible, this does not speak to the plausibility of the remaining hypotheses.

\section*{An alternative mode of inference}

An alternative to the Fisherian mode of inference finds its origins in Neyman (1923 [1990]).  Here, causal effects are defined individualistically but the researcher remains agnostic as to their exact distribution. Instead, the inferential target is average differences between potential outcomes and therefore average causal effects.

To see how this might proceed in the running example, assume that each individual $ij$ has four potential outcomes for each possible exposure to the treatment:
\begin{align*}
y_{ij}(00) &= y_{ij}(Z_{ij} = 0, Z_{i(-j)} = 0), \\
y_{ij}(10) &= y_{ij}(Z_{ij} = 1, Z_{i(-j)} = 0), \\
y_{ij}(01) &= y_{ij}(Z_{ij} = 0, Z_{i(-j)} = 1), \\
y_{ij}(11) &= y_{ij}(Z_{ij} = 1, Z_{i(-j)} = 1).
\end{align*}
This {\it exposure mapping} (Aronow and Samii, 2012a) includes the causal model \eqref{causal} as a special case, and is a fully general representation of the partial interference setup.

The average potential outcome associated with each exposure $t \in \{00,10,01,11\}$, $\mu(t) \equiv  \sum_{ij \in U} y_{ij}(t)/N$. Average causal effects are then defined as $\tau(t,t') \equiv \mu(t) - \mu(t') =
  \sum_{ij \in U} \left[y_{ij}(t') -  y_{ij}(t)\right]/N$, for some $t \neq t'$.  Define a variable $D_{ij}$ that takes on the value $(10 Z_{ij} + Z_{i(-j)})$, so that the observed outcome $y_{ij} = y_{ij}(D_{ij})$ and $\I \left[D_i = 01\right] \Leftrightarrow \I \left[Z_{ij} = 0 \cap Z_{i(-j)} = 1\right]$. Since labels $ij$ are exchangeable, each sample mean
$$
\widehat\mu(t) =\frac{\sum_{ij \in U} \I \left[D_{ij} = t\right] y_{ij} }{\sum_{ij \in U} \I \left[D_{ij} = t\right]}
$$
is unbiased for $\mu(t)$ (conditional on at least one unit being observed in the condition $t$). Then as a linear combination of two unbiased estimators, $\widehat\tau(t,t') = \widehat\mu(t) - \widehat\mu(t')$ is unbiased for $\tau(t,t')$. %Although BFP claims that this approach requires ``a decomposition of average treatment effects into parts due to interference and parts not due to interference," we can see that this is not the case: average causal effects are simply defined as differences in potential outcomes.

Under suitable regularity conditions (e.g., bounded potential outcomes), $\widehat\tau(t,t')$ is consistent and asymptotically normal by the LLN, Slutsky's Theorem and the CLT. Aronow and Samii (2012a, 2012c) %cite
demonstrate that a conservative variance estimator exists and is convergent, thus providing a basis for confidence intervals that asymptotically cover $\tau(t,t')$ with at least proper coverage.  Tests of {\it weak} null hypotheses are therefore possible. For example, if the confidence interval for $\tau(10,00)$ does not include $0$, we can reject the null hypothesis that the average effect of the treatment, given a housemate assigned to control, is zero. Importantly, rejection of the weak null hypothesis that $\tau(10,00) = 0$ implies rejection of the sharp null hypothesis $H_0: y_{ij}(10) - y_{ij}(00) = 0, \forall ij \in U$. Or, put simply, if the average effect is nonzero, it must be the case that at least one individual has been affected. This inferential framework thus provides valid tests of many of the same hypotheses that are entertained by BFP.

There are two important caveats to this approach. First, given a nonuniform network structure (such as those entertained by BFP), simple differences in means are rarely unbiased for $\tau(t,t')$. When the probability of receiving a given level of exposure is heterogeneous, statistical adjustments are required to derive an unbiased or consistent estimator of $\tau(t,t')$. For example, when estimating the spillover effect of having a neighbor treated, those with more neighbors are more likely to have treated neighbors. With simultaneous knowledge of the randomization scheme and the exposure model, it is possible to back out the implied probabilities of exposure for each unit. Unbiased or consistent estimation then proceeds using some form of an inverse-probability-weighted estimator \citep{horvitz52}, with the probabilities for each unit $ij$ being $\Pr(D_{ij} = t)$. Second, testing in this framework requires asymptotic approximations, the validity of which depends on the channels through which spillover occurs.
 Aronow and Samii (2012a) %(2012a \citet{aronowsamii}
 demonstrates that confidence intervals are justified when the scope of interference tends to be small relative to the size of the entire dataset (analogous to $m$-dependence). In small datasets, particularly ones where interference is highly nonlocalized, BFP's procedures may be more sensible for testing hypotheses about causal effects.

\section*{Concluding remarks}

BFP has introduced a set of tools for testing sharp null hypotheses that include interference between units. These tools will undoubtedly prove useful to researchers across the social sciences. However, it is important to understand the limitations of such an approach, and to place its contributions into a broader context.

Sensible test statistic selection remains an open field of inquiry. Until progress is made in this domain, it will be difficult to assess whether it is a model that is plausible, or if a test is simply ill-equipped to detect deviations from model predictions. Nevertheless, randomization guarantees that rejection of sharp null hypotheses remains valid. BFP's elucidation of this fact should not be ignored. 
In addition, political scientists should be aware that it is possible to estimate causal effects (and test hypotheses) under weaker assumptions than are presented in BFP. Rejection of a sharp null hypothesis may yield little information on either the direct effects of a treatment or its spillover effects. Where the recent movement in causal inference has been focused on unmodeled treatment effect heterogeneity \citep[e.g.,][]{angristpischke}, a semiparametric approach that does not require that causal effects have been completely modeled may be preferable. That being said, when dealing with data where asymptotic approximations are unlikely to hold (e.g, small experiments with wide-ranging network effects), BFP's approach may prove to be a more reliable basis for inference.

\begin{thebibliography}{}
%\bibitem[\protect\citeauthoryear{Angrist et~al.}{1996}]{air96} Angrist, Joshua D., Guido W. Imbens and Donald B. Rubin. 1996. Identification of Causal Effects Using Instrumental Variables. {\it Journal of the American Statistical Association}. 91: 444--55.
\bibitem[\protect\citeauthoryear{Angrist and Pischke}{2009}]{angristpischke} Angrist, J. D. and J.-S. Pischke. 2009.  {\it Mostly Harmless Econometrics: An Empiricist's Companion.}  Princeton, NJ: Princeton University Press.
\bibitem[\protect\citeauthoryear{Aronow and Samii}{2012}]{aronowsamii} Aronow, P. M. and C. Samii. 2012c. Conservative variance estimation for sampling designs with zero pairwise inclusion probabilities. {\it Survey Methodology}. In press.
%\bibitem[\protect\citeauthoryear{Hoeffding and Robbins}{1948}]{hoeffding48} Hoeffding, W. and H. Robbins (1948). The central limit theorem for dependent random variables. Duke Mathematical Journal 15(3), 773--780.
\bibitem[\protect\citeauthoryear{Horvitz and Thompson}{1952}]{horvitz52} Horvitz, D. G. and D. J. Thompson. 1952. A generalization of sampling without replacement from a finite universe. {\it J. Amer. Statist. Assoc.} 47, 663--684.
\bibitem[\protect\citeauthoryear{Isaki and Fuller}{1982}]{isakifuller} Isaki, C. T. and W. A. Fuller. 1982. Survey design under the regression superpopulation model. {\it J. Amer. Statist. Assoc.}  77, 89--96.
\bibitem[\protect\citeauthoryear{Szekely and Rizzo}{2009}]{rizzo}  Szekely, G. J. and M. L. Rizzo. 2009. Brownian distance covariance. {\it Ann. Appl. Stat.} 3, 1236--1265.
%%\bibitem[\protect\citeauthoryear{Green et~al.}{2012}]{reviewessay} Green, Donald P., Peter M. Aronow and Mary C. McGrath. 2012. Field Experiments and the Study of Voter Turnout. Forthcoming at {\it Journal of Elections, Public Opinion and Parties}.
%\bibitem[\protect\citeauthoryear{Imbens and Rubin}{1997}]{imbensrubin} Imbens, Guido W. and Donald B. Rubin. 1997. Estimating Outcome Distributions for Compliers in Instrumental Variables Models. {\it Review of Economic Studies} (4): 555--574.
%\bibitem[\protect\citeauthoryear{Mullainathan et~al.}{2010}]{Mullainathan} Mullainathan, Sendhil, Ebonya Washington and Julia R. Azari. 2010. The Impact of Electoral Debate on Public Opinions: An Experimental Investigation of the 2005 New York City Mayoral Election. In {\it Political Representation}, eds. Ian Shapiro, Susan Stokes, Elizabeth Wood and Alexander S. Kirshner. New York: Cambridge University Press.
\end{thebibliography}

\end{document}

### 2014-09-11 figures/teststat-parallel-summary.R
Adding figure creation file:

@@ -1,0 +1,80 @@
+
+source("simulation/setup.R")
+source("code/plotting.R")
+
+load("simulation/teststat-parallel-results.rda")
+
+searchParams<-expand.grid(SEARCH)
+
+res<-cbind(data.frame(powResults),searchParams)
+
+## Good Type I error for all tests
+res[res$beta==TRUTH$beta & res$tau==TRUTH$tau,]
+
+oneDplot<-function(x,f,...){
+	## f is a vector of TRUE/FALSE selecting rows of res to plot
+	## x is the name of the variable defining the x-axis
+	plot(range(res[[x]]),c(0,1),type="n",ylab="Proportion p<.05",...)
+	with(res[f,],{
+		 lines(eval(parse(text=x)),ssrTestStat,col="red")
+		 lines(eval(parse(text=x)),Mean.Diff,col="blue")
+		 lines(eval(parse(text=x)),KS.Test,col="black")
+		 lines(eval(parse(text=x)),Mann.Whitney.U,col="green") })
+	ll<-TRUTH[[x]]-.1*diff(range(res[[x]]))
+	ul<-TRUTH[[x]]+.1*diff(range(res[[x]]))
+	segments(rep(ll,3),
+			 .05+c(0,c(-2,2)*sim.se(REPETITIONS)),
+			 rep(ul,3),
+			 .05+c(0,c(-2,2)*sim.se(REPETITIONS)),col="gray",lwd=.5,lty=2)
+	points(TRUTH[[x]],.05,pch=19,cex=2)
+	legend("bottomright",
+		   col=c("red","blue","black","green"),
+		   legend=names(res)[1:4],
+		   lwd=2,
+		   bty="n")
+}
+
+
+## Power comparison holding tau fixed
+par(mfrow=c(1,3),oma=rep(0,4),mgp=c(1.5,.5,0),pty="s")
+
+oneDplot(x="beta",f=res$tau==TRUTH$tau,xlab="Beta")
+oneDplot(x="beta",f=res$tau==min(SEARCH$tau),xlab="Beta")
+oneDplot(x="beta",f=res$tau==max(SEARCH$tau),xlab="Beta")
+
+
+## Power comparison holding beta fixed
+par(mfrow=c(1,3),oma=rep(0,4),mgp=c(1.5,.5,0),pty="s")
+
+oneDplot(x="tau",f=res$beta==TRUTH$beta,xlab="Tau")
[…]

### 2014-09-11 code/covadjfns.R
Adding code dir

@@ -1,0 +1,136 @@
+##Functions for covariance adjustment
+
+### Used in interference.Rnw
+
+interference.hypothesis.fn <- function(ys,z,tau0,w) {
+  # a helper function to create constant additive models of effect
+  ##for each tau0 return a list of functions adjusting outcomes in accord with the interference hypotheses
+  ##assuming constant, additive treatment effects and constant spillover when spillover is hypothesized to occur
+  fn1<-function(ys,z,tau0){ ys - (z * tau0) }
+  ##adding (1-zi) into next just to ensure it only is picked up when zi is control
+  fn2<-function(ys,zi,zj,tau0,w){ ys - (zi * tau0) - (w * tau0 * zj*(1-zi))  }
+  ##rownnames(z)<-rownames(ys) ##make sure that treatment assignment has names
+  targets<-c("Richland","Midland","Saginaw")
+  sources<-c("Yakima","Saginaw","Midland");names(sources)<-targets
+  ##print(tau0)
+  adjys<-sapply(rownames(z),function(nm){ ##I bet we could use outer() to make this much faster
+    ##print(tau0)
+    if(nm %in% targets){
+      newy<-fn2(ys[nm,],zi=z[nm,],zj=z[sources[nm],],tau0=tau0,w=w)
+    }
+    if(!(nm %in% targets)){
+      newy<-fn1(ys[nm,],z[nm,],tau0=tau0)
+    }
+    return(newy)
+  })
+  names(adjys)<-rownames(z)
+  return(adjys)
+}
+
+
+
+covadjI.fn<-function(thefmla,tau0,thedata,X,lam,R,w,z=thez,teststats=test.statistics.aligned){ ##I for interference
+    ## X are covariates, Z is binary treatment, s is strata, R is outcome 
+    ## lam is tuning parameter for lasso
+    ## tau0 is a hypothesized effect (here entertaining constant treatment effect)
+    ## thefmla is a covariance adjustment formula.
+    ###stopifnot(length(R)==length(Z)) ## add other checks, be better if we could use names
+  ## w is a weight (how much spillover)
+  ##testsstats is a list of test statistic functions
+    
+  adjr<-interference.hypothesis.fn(ys=R,z=z,tau0=tau0,w=w)
+
[…]

### 2014-09-11 code/networkfns.R

@@ -1,0 +1,55 @@
+randomize <- function(n = 100, k = floor(n/2)) {
+  z <- vector("logical", n)
+  z[sample.int(n, k)] <- T
+  return(z)
+}
+
+network <- function(n = 100, p = 0.25){
+
+  # Generate matrix values, sampling 0 or 1 with given probabilities
+  matvals = sample (c(0,1), n*(n-1)/2, replace = TRUE, prob = c(1-p, p))
+
+  # From the values above, generate a symmetric matrix 
+    networkmat = matrix(rep(0, n*n), ncol = n)
+    mv = 1
+    for (i in 1:n){
+      for (j in 1:n){
+        if (i>j) {
+          networkmat [i,j] = matvals[mv]
+            networkmat [j,i] = matvals[mv]
+            mv = mv + 1
+        }
+      }
+    }
+  return(networkmat)
+}
+
+covariates <- function(n = 100) {
+  x1 <- rnorm(n, 0, 10)
+  x2 <- rbeta(n, 7, 1)
+  data.frame(x1, x2)
+}
+
+within.block.network <- function(B) {
+  n <- length(B) 
+  S <- outer(X =B, Y =B, `==`)
+  S <- replace(S, 1:n^2, as.numeric(S))
+  diag(S) <- 0
+  return(S)
+}
+
+# if both n and average.edges are odd, the average degree will be close but
+# not exactly the requested value
+average.edge.network <- function(n = 100, average.edges = 5) {
+  total.edges <- floor(n * average.edges / 2)
+  total.possible <- (n * (n - 1)) / 2
+  edge.numbers <- sample.int(total.possible, total.edges)
+  edges <- numeric(total.possible)
+  edges[edge.numbers] <- 1
+  S <- matrix(0, nrow = n, ncol = n)
+  S[upper.tri(S)] <- edges
+  tmp <- t(S)
+  S[lower.tri(S)] <- tmp[lower.tri(tmp)]
+  return(S)
+}
+

### 2014-09-11 code/operchar.R

@@ -1,0 +1,194 @@
+
+## Functions to inspect the operating characteristics of our
+## simulations and testStatistic
+
+## Type I error rate: is the size of the test no greater than the
+## level?
+
+## ??Bias?? : Across randomizations is the mean of the point estimate
+## equal to the true value.
+require(RItools)
+
+get.p.truth<-function(obj,true.params,alphas="auto"){
+  
+  theps<-p.values(obj)
+  par.names<-names(obj@params)
+  
+  p.true.params<-theps[apply(theps[,par.names,drop=FALSE],1,function(x){
+    identical(as.numeric(x),true.params)
+  }),]
+
+  return(p.true.params)
+}
+
+test.size<-function(sim.obj,truth,alphas="auto"){
+  if(alphas[1]=="auto"){
+    alphas<-sort(unique(unlist(lapply(sim.obj,function(obj){p.values(obj)$p}))))
+  }
+  p.truths<-sapply(sim.obj,get.p.truth,true.params=truth)
+  
+  res<-sapply(alphas,function(a){c(nominal.alpha=a,
+                                   realized.alpha.rand=mean(p.truths<=a)
+                                   )
+                               })
+  return(res)
+}
+
+  
+
+get.2d.mn<-function(obj){ ##get 2d mean
+  ##This function is useful for getting HL estimates (approx) with 2 params
+  require(depth)
+  pts<-na.omit(point(obj))
+  par.nms<-grep("p",names(pts),invert=TRUE,value=TRUE)
+  if(nrow(pts)!=1){
+  mn<-trmean(pts[,par.nms],alpha=0)}else{
+    mn<-as.numeric(pts[,par.nms])
+  }
+  
+  return(mn)
+}
+
+
+get.2d.mn2<-function(obj){ ##get 2d mean of highest p-value hyps
+  pts<-na.omit(obj)
+  pts<-pts[pts[,"p"]==max(pts[,"p"]),,drop=FALSE]
+  par.nms<-grep("p",names(pts),invert=TRUE,value=TRUE)
+  if(length(par.nms)>1){
+    if(nrow(pts)!=1){
+      mn<-trmean(pts[,par.nms],alpha=0)
+    }else{
+      mn<-as.numeric(pts[,par.nms])
+    }
+  }else{
+    mn<-mean(pts[,par.nms])
+  }
+  return(mn)
+}
+
+get.2d.mn3<-function(pts){
+  require(depth)
[…]

### 2014-09-11 code/plotting.R

@@ -1,0 +1,451 @@
+################################################################################
+# Plotting/graphing Routines
+################################################################################
+par(oma=rep(0,4))
+
+library(igraph)
+library(colorspace)
+
+plotexperiment <- function(s, z, mode = "undirected",
+       vertex.shape = ifelse(z, "circle", "square"),
+       layout = layout.fruchterman.reingold,
+       vertex.size = 6,
+       vertex.color = ifelse(z, "#252525", "#A0A0A0"),
+       vertex.label = NA,
+       edge.color = "black",
+       ...) {
+
+  tmp <- graph.adjacency(s, mode = mode)
+
+  plot(tmp,
+       vertex.shape = vertex.shape,
+       layout = layout,
+       vertex.size = vertex.size,
+       vertex.color = vertex.color,
+       vertex.label = vertex.label,
+       edge.color = edge.color,
+       margin=0,
+       ...)
+}
+
+
+make.filled.contour.plot<-function(mat,main,levs=10,col=rev(terrain_hcl(levs)),
+                                   addTruth=TRUE,autolab=TRUE,
+                                   thelevels=NULL,
+                                   true.params, ##names and values of the parameters
+                                   thez="p",
+                                   ...){
+
+  params<-names(true.params)
+
+  mat.p<- matrix(mat[,thez],
+                 nrow=length(unique(mat[,params[1]])),
+                 ncol=length(unique(mat[,params[2]])),
+                 dimnames=list(sort(unique(mat[,params[1]])),
+                   sort(unique(mat[,params[2]]))
+                   )
+                 )
+  theps<-sort(unique(as.numeric(mat.p)))
+  p.line.levels<-quantile(theps,seq(0,1,length=levs) )
+  if(addTruth){
+    p.truth<-mat[simp.eq.test(mat[,params[1]],true.params[[1]])&simp.eq.test(mat[,params[2]],true.params[[2]]),thez]
+    ##main<-paste(main,"(",thealpha,"test level=",round(p.truth,3),")")
+  }
[…]

### 2014-09-11 code/setup-clusters.R

@@ -1,0 +1,43 @@
+
+.libPaths(".libraries")
+library(Rmpi)
+library(parallel)
+
+if (Sys.getenv("CORES") == "") {
+  numcores <- detectCores() ##as.numeric(system("sysctl hw.ncpu | awk '{print $2}'",intern=TRUE)) 
+} else {
+  numcores <- as.numeric(Sys.getenv("CORES")[[1]])  
+}
+
+if(!exists("clusterswitch")){ clusterswitch<-Sys.getenv("CLUSTER") }
+stopifnot(clusterswitch %in% c("","none","ec2","dkh","taub","keeling"))
+
+if (clusterswitch == "none") {
+  ##thecluster<-rep("localhost",numcores)
+  ##cl<-makeCluster(thecluster)
+}
+if (clusterswitch == "") {
+  thecluster<-rep("localhost",numcores)
+  cl<-makeCluster(thecluster)
+}
+if (clusterswitch == "ec2"){
+  nodes <- system("grep node[0-9] /etc/hosts | cut -d ' ' -f 2",intern=TRUE)
+  mastercores<-length(system("ssh master  cat /proc/cpuinfo | grep processor | cut -d ':' -f 2",intern=TRUE))
+  nodecores<-length(system("ssh node001 cat /proc/cpuinfo | grep processor | cut -d ':' -f 2",intern=TRUE))
+  thecluster <- c(rep("master",mastercores),rep(nodes,each=nodecores))
+  cl<-makeCluster(thecluster)
+}
+if (clusterswitch == "dkh"){
+  nodes<-c(rep("localhost",numcores),rep("jwbowers.pol.illinois.edu",8))
+  thecluster <- nodes
+  cl<-makeCluster(thecluster)
+}
+if (clusterswitch == "taub" & Sys.getenv("PBS_NODEFILE") != "") {
+  thecluster <- system("cat $PBS_NODEFILE|wc -l",intern=TRUE)
+  cl<-makeCluster(thecluster[1],type="MPI")
+}
+if (clusterswitch == "keeling") {
+  cl<-makeCluster(numcores,type="MPI")
+}
+
+

### 2014-09-11 code/sweave-settings.R

@@ -1,0 +1,50 @@
+options(SweaveHooks = list(
+  # Sweave related options
+  fig = function() {
+         par(mar=c(3.5, 3, 1.1, 0),
+         pty="s",
+         mgp=c(1.5,0.5,0),
+         oma=c(0,0,0,0))}),
+
+  # echo = function() { 
+  #         options(continue=" ") ##Don't show "+" prompts,
+  #         options(prompt=" ")}),
+
+  # error=function(){ ##show "+" prompts after errors
+  #         options(prompt="> ",continue="+ ");NULL},
+
+  # general options
+  digits = 4,
+  scipen = 5,
+  show.signif.stars = FALSE)
+
+set.seed(20110515)
+
+# has the number of simulation samples been set in the evironment?
+if (Sys.getenv("SAMPLES") == "") {
+  SAMPLES <- 10000  
+} else {
+  SAMPLES <- as.numeric(Sys.getenv("SAMPLES")[[1]])  
+}
+
+# has the number of cores been set in the system evironment or in the running R workspace?
+if (!exists("CORES")&Sys.getenv("CORES") == "") {
+  CORES <- 2  ##default is 2 cores
+}
+
+if (!exists("CORES")&Sys.getenv("CORES") != "") {
+  CORES <- as.numeric(Sys.getenv("CORES")[[1]])  
+}
+
+# a little helper to do sci. notation from
+# http://stackoverflow.com/questions/8366324/r-sweave-formatting-numbers-with-sexpr-in-scientific-notation
+sn <- function(x,digits)
+{
+  if (x==0) return("0")
+  ord <- floor(log(abs(x),10))
+  x <- x / 10^ord
+  if (!missing(digits)) x <- format(x,digits=digits)
+  if (ord==0) return(as.character(x))
+  return(paste(x,"\\\\times 10^{",ord,"}",sep=""))
+}
+

### 2014-09-11 simulation/teststat-parallel-results.R
Add simulation files

@@ -1,0 +1,24 @@
+## A file to extract power and error rate from the huge list of RItest objects
+
+library(RItools,lib.loc=".libraries")
+
+source("simulation/setup.R")
+
+load("simulation/teststat-parallel.rda")
+
+simPow2<-function(lst){
+		   m <- sapply(lst,function(l){ l[-1,"p.value"]})
+           m <- m <= ALPHA
+		   rowMeans(m)
+}
+
+powResults<-lapply(testStatResults,function(obj){ simPow2(obj) })
+
+searchParams<-expand.grid(SEARCH)
+searchParams2<-testStatResults[[1]][[1]]@params
+stopifnot(all.equal(searchParams,searchParams2))
+
+save(powResults,file="simulation/teststat-parallel-results.rda")
+
+
+

### 2014-09-11 simulation/teststat-parallel.R

@@ -1,0 +1,105 @@
+#.libPaths("../.libraries")
+masterdir<-getwd()
+
+library(RItools,lib.loc=".libraries")
+options("RItools-lapply"=lapply)
+options("RItools-sapply"=sapply)
+
+source("simulation/setup.R")
+
+# A simulation using different uniformity trials
+n <- 256
+e <- n * DENSITY
+simsamples<-10000
+##REPTITIONS<-100
+
+sampler <- simpleRandomSampler(total = n, treated = n/2)
+Zs <- sampler(REPETITIONS)$samples # discard the weight element
+
+# the basic data has S, the network, and y0, the unif trial data used in the other sims
+uniformityData <- simulationData(n,e)
+##num1hopPeers<-colSums(uniformityData$S)
+themodel <- interference.model.maker(uniformityData$S)
+ssrTestStat<-SSRTmaker(uniformityData$S)
+
+dotestMaker<-function(model,y0,truth,TZ,thegrid,simsamples){
+     force(model);force(y0);force(truth);force(TZ);force(thegrid);force(simsamples)
+     function(z){
+          y <- invertModel(model, y0, z, truth$beta, truth$tau)
+          rit<-RItest(y,z, TZ, model, thegrid,samples=simsamples)
+	  ## cbind(rit@params, p = rit[-1, "p.value"])
+	 return(rit[-1, "p.value"])
+     }
+}
+
+
+dotest<-dotestMaker(model=themodel,
+                    y0=uniformityData$data$y0,
+                    truth=TRUTH,
+                    TZ=ssrTestStat,
+                    thegrid=SEARCH,
+                    simsamples=simsamples)
+
+source("code/setup-clusters.R")
+##library(parallel)
+##thecluster<-rep("localhost",8) ##rep(c("localhost","jwbowers.pol.illinois.edu"),c(8,8))
+##thecluster<-rep(c("localhost","jwbowers.pol.illinois.edu"),c(12,8))
+##cl<-makeCluster(thecluster,type="PSOCK")
+
+## system.time(tmp<-dotest(Zs[,1]))
+## Rprof(line.profiling=TRUE)
+##tmp<-dotest(Zs[,1])
+## Rprof(NULL)
+## blah<-summaryRprof(lines="both")
+
+##clusterEvalQ(cl,setwd("/Users/jwbowers/Documents/PROJECTS/FisherSUTVA/"))
[…]

### 2014-09-11 simulation/models.R
Utility files for simulations

@@ -1,0 +1,34 @@
+## The UniformityModel is a datatype defined in the R package on CRAN. 
+## The first argument is a function to go from observed data to the uniformity trial 
+## The second argument is the inverse of the first model (it generates
+## observed data from the uniformity trial)
+
+## y0 is the uniformity trial
+## y is the observed outcome
+## z is 0 or 1 denoting treatment assignment
+
+## Since this model depends on S, we need to generate for a specific S
+## The growthCurve function specifies how spillover happens.
+
+growthCurve <- function(beta, tau, x) {
+  (beta + (1 - beta) * exp(-(tau^2) * x))
+}
+
+interference.model.maker <- function(S) {
+  ## just to be safe about lazy argument evaluation,
+  ## probably not needed
+  force(S)
+  
+  ## we will return a UniformityModel object  
+  UniformityModel(
+    function(y, z, beta, tau) {
+      zS <- as.vector(z %*% S)
+      z * (1 / beta) * y +
+        (1 - z) * y / growthCurve(beta, tau, zS)
+    },
+    function(y0, z, beta, tau) {
+      zS <- as.vector(z %*% S)
+      z * beta * y0 +
+        (1 - z) * growthCurve(beta, tau, zS) * y0  
+    })
+}

### 2014-09-11 simulation/setup.R

@@ -1,0 +1,103 @@
+require(RItools, lib.loc = ".libraries")
+
+set.seed(20120614)
+
+source("simulation/models.R")
+# values we will use for creating data most often
+TRUTH <- list(beta = 2, tau = 0.5)
+
+set.seed(20120618)
+
+MAXI <- 32
+MAXN <- MAXI^2
+ORDER <- sample.int(MAXN, MAXN)
+UNIFORMITY <- runif(MAXN, min = 30, max = 70)
+X1 <- sample.int(5, MAXN, replace = T)
+
+# blocks: determines sample size
+simulationData <- function(n, edges, makey0 = function(base, S) {
+  base * growthCurve(TRUTH$beta, TRUTH$tau, rowSums(S))}) {
+
+  # make sure both are integers
+  edges <- floor(edges)
+  n <- floor(n)
+
+  stopifnot(n <= MAXN)
+  ##stopifnot(edges <= choose(n, 2))
+  if(edges > choose(n,2)){ warning("You may have an overly dense network.") }
+  sampler <- simpleRandomSampler(total = n, treated = n / 2)
+
+
+  nodes <- ORDER[1:n]
+  cx <- nodes %% MAXI
+  cy <- floor(nodes / MAXI)
+
+  data <- data.frame( cx, cy)
+  rownames(data) <- nodes
+
+  dists <- dist(data[,c("cx", "cy")])
+
+  tmp <- matrix(Inf, ncol = n, nrow = n)
+  tmp[lower.tri(tmp, diag = F)] <- dists
+  torder <- order(tmp)[1:edges]
+
+  S <- matrix(0, ncol = n, nrow = n)
+  if(edges!=0) { S[torder] <- 1 }
+  S[upper.tri(S, diag = F)] <- t(S)[upper.tri(S, diag = F)]
+
+  data$base <- UNIFORMITY[1:n]
+  data$y0 <-  with(data, makey0(base, S))
+
+  return(list(n = n,
+              sampler = sampler,
+              data = data,
+              S = S))
+}
+
+DENSITY <- 2
+
+# the environemental variable SEARCH_POINTS controls how fine grained a search
+# to perform. More points means a more fine grained search, but takes more
+# processing power.
+
+if (Sys.getenv("SEARCH_POINTS") == "") {
+  SEARCH_POINTS <- 20
+} else {
[…]

### 2014-10-06 simulation/teststat-parallel-results.R
Updating results creation

@@ -6,13 +6,17 @@ source("simulation/setup.R")
 
 load("simulation/teststat-parallel.rda")
 
+library(abind)
+tzResultsArrSSR<-abind(testStatResults[["ssrTestStat"]],along=-1)
+
+
 simPow2<-function(lst){
-		   m <- sapply(lst,function(l){ l[-1,"p.value"]})
-           m <- m <= ALPHA
-		   rowMeans(m)
+	m <- sapply(lst,function(l){ l$p})
+	m <- m <= ALPHA
+	rowMeans(m)
 }
 
-## powResults<-lapply(testStatResults,function(obj){ simPow2(obj) })
+powResults<-lapply(testStatResults,function(obj){ simPow2(obj) })
 
 
 

### 2014-10-06 paper/introduction.tex

--- 2014-01-27 notes/notes.tex
@@
-\documentclass[12pt]{article}
-\usepackage{parskip}
-\usepackage[margin=1in]{geometry}
-\usepackage{natbib}
+ \citet{bowers2013sutva} showed that one could use Fisher's
+ randomization-based hypothesis testing framework to assess models of
+ treatment propagation and spillover across social networks. Working in a
+ frequentist framework, they showed that one could make statements such as:
+ "Our design provides strong evidence against the idea that the treatment
+ spilled over by more than 5 points given this model of treatment
+ propagation." This research note shows that the test statistic that they used
+ --- the Kolmogorov-Smirnov test statistic  \citep[\S
+ 5.4]{MylesHollander1999a} --- could be improved substantially. We replicate
+ their analyses to demonstrate that the sum of squared residuals from a linear
+ model in which aspects of the network itself play a role.
 
-\title{Test statistic selection for randomization inference under interference}
-\author{Peter Aronow and Jake Bowers}
-\date{\today}
+ They also showed that the strength of evidence against the
+ specific hypotheses implied by a given model varied with different features
+ of the design as well as the extent to which the true process by which
+ treatment given to one unit changed the outcomes of another unit diverged
+ from the model. Throughout their paper they used the  Kolmogorov-Smirnov (KS)
+ test statistic so that their tests would be sensitive differences in the
+ treatment and control distributions implied by different hypotheses and not
+ merely sensitive to differences in one aspect of those distributions (such as
+ the differences in the mean).\footnote{If the empirical cumulative distribution function (ECDF) of
+   the treated units is $F_1$ and the ECDF of the control units is $F_0$ then
+   the KS test statistic is
+   $\T(\yu,\bz) = \underset{i = 1,\ldots,n}{\text{max}}\left[F_1(y_{i,\bzero})
+     - F_0(y_{i,\bzero})\right]$, where $F(x)=(1/n)\sum_{i=1}^n I(x_i \le x)$
+   records the proportion of the distribution of $x$ at or below $x_i$
+   \citep[\S 5.4]{MylesHollander1999a}.}
 
-\begin{document}
-\maketitle
-
-%% 4 pages
-
- \citep{bowers2013sutva} followed Rosenbaum by recommending effect increasing
- statistic.
-
- controls Type I error, but power can be very poor for these types of models.
- We explain why and propose a refinement. Multivariate dependence of
- parameters and outcome space. A weighted E-statistic.
-
-Fisher's sharp null hypothesis of no effects combined with his
-randomization-based hypothesis testing framework provided a very clear and
-clean vision of what it means to do statistical inference \citep[Chap
-2]{fisher:1935}. In recent years, a number of scholars have realized that this
-clarity also implied great flexibility in regard to the scientific questions
+ provides both clarity for
+ non-parametric statistical inference  also implied great flexibility in regard to the scientific questions
 that one might engage: for example, one could use the sharp null itself to
-detect interference among units \citep{aronow-general}, and sharp hypotheses
+detect spillover  among units \citep{aronow-general} or to characterize the
+benefits from treatment on treated units even if  and sharp hypotheses
 about effects (in contrast to hypotheses about no effects) have been developed
@@
 
+Fisher's sharp null hypothesis of no effects combined with his
+randomization-based hypothesis testing framework provides a very clear and
+clean vision of what it means to test a hypothesis \citep[Chap
+2]{fisher:1935} and \citep{rosenbaum2010design}.
 One problem that has become apparent, as scholars have extended Fisher's basic
@@
 Under the model, model consistent.
-
-\bibliographystyle{apsr}
-\bibliography{/Users/jwbowers/Documents/PROJECTS/BIB/big}
-
-\end{document}

### 2014-10-06 paper/paper.tex

\documentclass[12pt]{article} 
\RequirePackage{snapshot}

\author{Peter Aronow \thanks{Assistant Professor, Dept of Political Science,
    Yale University} \and Jake Bowers \thanks{Associate Professor, Dept of Political
    Science and Statistics, University of Illinois @ Urbana-Champaign
    (\texttt{jwbowers@illinois.edu}). \textit{Acknowledgements:} Part
    of this work funded by NSF Grants SES-0753168 and SES-0753164
    (joint with Ben Hansen, Dept of Statistics, University of
    Michigan).} \and Mark Fredrickson \thanks{Ph.D. Student in Statistics and Political Science, University of Illinois @ Urbana-Champaign.}}

  \date{\today}

\usepackage{../styles/common}

%%% Paper specific packages
\usepackage{setspace}
\usepackage[letterpaper,bottom=1in,top=1in,right=1in,left=1in,includemp=FALSE]{geometry}
\usepackage{microtype} %only useful with pdflatex

%%% Document Structure Formatting
\usepackage[compact,bottomtitles]{titlesec}
%\titleformat{ ⟨command⟩}[⟨shape⟩]{⟨format⟩}{⟨label⟩}{⟨sep⟩}{⟨before⟩}[⟨after⟩]
\titleformat{\section}[hang]{\large\bfseries}{\thesection}{.5em}{\hspace{0in}}[\vspace{-.2\baselineskip}]
\titleformat{\subsection}[hang]{\bfseries}{\thesubsection}{.5em}{\hspace{0in}}[\vspace{-.2\baselineskip}]
\titleformat{\subsubsection}[hang]{\itshape}{\thesubsubsection}{.5em}{\hspace{0in}}[\vspace{-.2\baselineskip}]
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1ex}{}{\vspace{-.2\baselineskip}}

%\titlespacing*{ ⟨command⟩}{⟨left⟩}{⟨beforesep⟩}{⟨aftersep⟩}[⟨right⟩]
\titlespacing{\section}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}
\titlespacing{\subsection}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}
\titlespacing{\subsubsection}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}

\newif\ifanon

% \ifanon
% \footnote{We are awesome!}
% \else
% \fi

\begin{document}
\VerbatimFootnotes
\input{titlepage}
%\setstretch{1.6} %1.6 or 1.37 which I think is "true double spacing"
%\onehalfspacing
%%I don't think I need to double space footnotes for AJPS. Check this. 
\abovedisplayshortskip=0pt
\belowdisplayshortskip=0pt
\abovedisplayskip=0pt
\belowdisplayskip=0pt
%\tableofcontents

% include stuff here!
% include stuff here!

\input{introduction.tex}

\cleardoublepage
%\appendix % is redefined by hyperref and a mess, do the appendix stuff by hand instead
% problems with pdf bookmarks and appendix caused by resetting section counter.
% solution here: http://www.latex-community.org/forum/viewtopic.php?f=45&t=4950
\begin{singlespacing}
%\renewcommand{\thesection}{Appendix \thesection}
%\setcounter{section}{0}
\newcounter{mycounter}
\setcounter{mycounter}{0}
%\renewcommand{\thesection}{Appendix \Alph{mycounter}}
\let\osection\section
\renewcommand{\section}{\stepcounter{mycounter}\osection}
\renewcommand\thesection{Appendix \Alph{mycounter}}
%\titleformat{\section}[hang]{\large\bfseries}{\thesection}{.5em}{\hspace{0in}}[]
%\input{codeexample}
%%\clearpage
%%\input{generalinterference}

\clearpage
\bibliographystyle{apsr} 
\bibliography{../BIB/big}

% this appendix is the response to reviewers. TODO: remove for released
% version
% \clearpage
% \setlength{\parindent}{0pt}
% \addtolength{\parskip}{.5\baselineskip}
% \pagenumbering{arabic} %\setcounter{page}{1}
% \setcounter{mycounter}{0}
% \setcounter{secnumdepth}{0}
% %\let\osection\section
% %\renewcommand{\section}{\stepcounter{mycounter}\osection}
% \renewcommand\thesection{\arabic{mycounter}}
% \input{response}
% 
\end{singlespacing}
%\bibliography{/Users/jwbowers/Documents/BIB/trunk/big}
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-10-06 paper/titlepage.tex

\begin{titlepage}
\maketitle

%\SweaveOpts{echo=true}

\begin{abstract}

\end{abstract}

Key Words: Causal effect; Interference; Randomized experiment;
Randomization inference; Fisher's Sharp Null Hypothesis; SUTVA

\end{titlepage}

### 2014-10-06 code/teststatistics.R
Starting to add some test statistics to code/ to make merging into the RItools master easier

@@ -1,0 +1,68 @@
+# These are some test statistics for permutation based randomization distributions.
+
+require(RItools)
+
+baseKSTest<-function(y,z){
+	# next borrowed from KS test
+	# next borrowed from KS test
+	# first, set up using the KS test var names
+	x <- y[z == 1]
+	y <- y[z == 0]
+
+	x <- x[!is.na(x)]
+	n <- length(x)
+
+	y <- y[!is.na(y)]
+	n.x <- as.double(n)
+	n.y <- length(y)
+
+	n <- n.x * n.y/(n.x + n.y)
+	w <- c(x, y)
+	z <- cumsum(ifelse(order(w) <= n.x, 1/n.x, -1/n.y))
+	if (length(unique(w)) < (n.x + n.y)) {
+		z <- z[c(which(diff(sort(w)) != 0), n.x + n.y)]
+	}
+
+	return(max(abs(z)))
+}
+
+ksNetResidTestStatisticMaker <- function(S){
+	## This function represents a bit of the true model of propagation in addition to knowledge about the network
+	force(S)
+	total1HopPeers<-colSums(S) ## S is undirected and symmetric
+	function(y, z) {
+		e<-lm.fit(cbind(1,z,as.vector(z %*% S),total1HopPeers),y)$residuals
+		baseKSTest(y=e,z=z)
+	}
+}
+
+
+ksNetTestStatisticMaker <- function(S){
+	## This function represents a bit of the true model of propagation in addition to knowledge about the network
+	force(S)
+	total1HopPeers<-colSums(S) ## S is undirected and symmetric
+	function(y, z) {
+		e<-lm.fit(cbind(1,z,total1HopPeers),y)$residuals
+		baseKSTest(y=e,z=z)
+	}
+}
+
+
+
+ssrSimpleTestStatistic<- function(y,z){
+	sum(lm.fit(cbind(1,z),y)$residuals^2) #resid(lm(y~z+d+total1HopPeers))^2)
+}
+
+
+
+ssrNetTestStatistic<-function(S){
+	## This statistic only takes into account the degree of the nodes.
+	force(S)
+	total1HopPeers<-colSums(S) ## degree, S is undirected and symmetric so colSums=rowSums
+	function(y,z){
+		sum(lm.fit(cbind(1,z,total1HopPeers),y)$residuals^2) #resid(lm(y~z+d+total1HopPeers))^2)
+	}
+}
+
+
+

### 2014-10-06 simulation/teststat-parallel.R
New version comparing only KS and SSR but with different amounts of network information represented in the test statistic

@@ -1,11 +1,14 @@
 #.libPaths("../.libraries")
 masterdir<-getwd()
+require(compiler)
+enableJIT(3)
 
 library(RItools,lib.loc=".libraries")
 options("RItools-lapply"=lapply)
 options("RItools-sapply"=sapply)
 
 source("simulation/setup.R")
+source("code/teststatistics.R")
 
 ## for each of REPETITIONS draws from the randomization distribution, do simsamples tests.
 
@@ -17,6 +20,7 @@ uniformityData <- simulationData(n,e)
 ##num1hopPeers<-colSums(uniformityData$S)
 themodel <- interference.model.maker(uniformityData$S)
 ssrTestStat<-SSRTmaker(uniformityData$S)
+ksResidTestStat<-ksNetResidTestStatisticMaker(uniformityData$S)
 
 dotestMaker<-function(model,y0,truth,TZ,thegrid,simsamples){
 	force(model);force(y0);force(truth);force(TZ);force(thegrid);force(simsamples)
@@ -51,6 +55,9 @@ source("code/setup-clusters.R")
 ##clusterEvalQ(cl,setwd("/Users/jwbowers/Documents/PROJECTS/FisherSUTVA/"))
 clusterEvalQ(cl,.libPaths(".libraries"))
 clusterEvalQ(cl,library(RItools)) ##,lib.loc=".libraries"))
+clusterEvalQ(cl,library(compiler))
+clusterEvalQ(cl,enableJIT(3))
+clusterEvalQ(cl,source("code/teststatistics.R"))
 clusterEvalQ(cl,options("RItools-lapply"=lapply))
 clusterEvalQ(cl,options("RItools-sapply"=sapply))
 ## clusterExport(cl,"dotest")
@@ -58,61 +65,37 @@ clusterExport(cl,"growthCurve")
 
 
 ##clusterEvalQ(cl,if(length(grep("FisherSUTVA$",getwd()))==0){setwd("$HOME/Documents/PROJECTS/FisherSUTVA")})
-##res<
-testStats <- list("ssrTestStat" = ssrTestStat,
-		  "Mean Diff" = mean.difference,
-		  "KS Test" = ksTestStatistic,
-		  "Mann-Whitney U" = mann.whitney.u)
 
+testStats <- list("SSR Test Net Full" = ssrTestStat, ## test stat including a bit of the true model
+				  "SSR Test Net Degree" = ssrNetTestStatistic, ## test stat including only network degree
[…]

### 2014-10-06 simulation/teststat-parallel.R
Fixing names of functions.

@@ -20,7 +20,9 @@ uniformityData <- simulationData(n,e)
 ##num1hopPeers<-colSums(uniformityData$S)
 themodel <- interference.model.maker(uniformityData$S)
 ssrTestStat<-SSRTmaker(uniformityData$S)
-ksResidTestStat<-ksNetResidTestStatisticMaker(uniformityData$S)
+ksNetResidTestStat<-ksNetResidTestStatisticMaker(uniformityData$S)
+ksNetTestStatistic<-ksNetTestStatisticMaker(uniformityData$S)
+
 
 dotestMaker<-function(model,y0,truth,TZ,thegrid,simsamples){
 	force(model);force(y0);force(truth);force(TZ);force(thegrid);force(simsamples)
@@ -41,6 +43,7 @@ dotestMaker<-function(model,y0,truth,TZ,thegrid,simsamples){
 ## 		    simsamples=simsamples)
 
 source("code/setup-clusters.R")
+require(snow)
 ##library(parallel)
 ##thecluster<-rep("localhost",8) ##rep(c("localhost","jwbowers.pol.illinois.edu"),c(8,8))
 ##thecluster<-rep(c("localhost","jwbowers.pol.illinois.edu"),c(12,8))
@@ -69,7 +72,7 @@ clusterExport(cl,"growthCurve")
 testStats <- list("SSR Test Net Full" = ssrTestStat, ## test stat including a bit of the true model
 				  "SSR Test Net Degree" = ssrNetTestStatistic, ## test stat including only network degree
 				  "SSR Test" = ssrSimpleTestStatistic, ## test stat ignoring network
-				  "KS Test Net Full" = ksResidTestStat, ## test stat including some of the true model
+				  "KS Test Net Full" = ksNetResidTestStat, ## test stat including some of the true model
 				  "KS Test Net Degree" = ksNetTestStatistic, ## test stat including only network degree
 				  "KS Test" = ksTestStatistic  ## test stat ignoring network
 				  )

### 2014-10-08 simulation/teststat-parallel.R
Fixing test statististics

@@ -70,12 +70,12 @@ clusterExport(cl,"growthCurve")
 ##clusterEvalQ(cl,if(length(grep("FisherSUTVA$",getwd()))==0){setwd("$HOME/Documents/PROJECTS/FisherSUTVA")})
 
 testStats <- list("SSR Test Net Full" = ssrTestStat, ## test stat including a bit of the true model
-				  "SSR Test Net Degree" = ssrNetTestStatistic, ## test stat including only network degree
-				  "SSR Test" = ssrSimpleTestStatistic, ## test stat ignoring network
-				  "KS Test Net Full" = ksNetResidTestStat, ## test stat including some of the true model
-				  "KS Test Net Degree" = ksNetTestStatistic, ## test stat including only network degree
-				  "KS Test" = ksTestStatistic  ## test stat ignoring network
-				  )
+		  "SSR Test Net Degree" = ssrNetTestStatistic, ## test stat including only network degree
+		  "SSR Test" = ssrSimpleTestStatistic, ## test stat ignoring network
+		  "KS Test Net Full" = ksNetResidTestStat, ## test stat including some of the true model
+		  "KS Test Net Degree" = ksNetTestStatistic, ## test stat including only network degree
+		  "KS Test" = ksTestStatistic  ## test stat ignoring network
+		  )
 ##		  "Mann-Whitney U" = mann.whitney.u)
 ##		  "Mean Diff" = mean.difference,
 
@@ -88,11 +88,11 @@ names(testStatResults)<-names(testStats)
 for(i in 1:length(testStats)){
 	TZ<-testStats[[i]]
 	dotest<-dotestMaker(model=themodel,
-						y0=uniformityData$data$y0,
-						truth=TRUTH,
-						TZ=TZ,
-						thegrid=SEARCH,
-						simsamples=simsamples)
+			    y0=uniformityData$data$y0,
+			    truth=TRUTH,
+			    TZ=TZ,
+			    thegrid=SEARCH,
+			    simsamples=simsamples)
 	clusterExport(cl,"dotest")
 	clusterSetupRNG(cl,seed=rep(1,6)) ## use same stream of randomness for each test statistic
 	testStatResults[[i]]<-parCapply(cl,Zs,function(z){ dotest(z)})

### 2014-10-08 code/teststatistics.R
Fixing a test statistic

@@ -55,7 +55,7 @@ ssrSimpleTestStatistic<- function(y,z){
 
 
 
-ssrNetTestStatistic<-function(S){
+ssrNetTestStatisticMaker<-function(S){
 	## This statistic only takes into account the degree of the nodes.
 	force(S)
 	total1HopPeers<-colSums(S) ## degree, S is undirected and symmetric so colSums=rowSums

### 2014-10-08 simulation/teststat-parallel.R

@@ -1,5 +1,6 @@
 #.libPaths("../.libraries")
-masterdir<-getwd()
+##masterdir<-getwd()
+if(length(grep("TestStatRIInterference$",getwd()))==0){setwd("..")}
 require(compiler)
 enableJIT(3)
 
@@ -20,6 +21,7 @@ uniformityData <- simulationData(n,e)
 ##num1hopPeers<-colSums(uniformityData$S)
 themodel <- interference.model.maker(uniformityData$S)
 ssrTestStat<-SSRTmaker(uniformityData$S)
+ssrNetTestStatistic<-ssrNetTestStatisticMaker(uniformityData$S)
 ksNetResidTestStat<-ksNetResidTestStatisticMaker(uniformityData$S)
 ksNetTestStatistic<-ksNetTestStatisticMaker(uniformityData$S)
 
@@ -43,7 +45,7 @@ dotestMaker<-function(model,y0,truth,TZ,thegrid,simsamples){
 ## 		    simsamples=simsamples)
 
 source("code/setup-clusters.R")
-require(snow)
+##require(snow)
 ##library(parallel)
 ##thecluster<-rep("localhost",8) ##rep(c("localhost","jwbowers.pol.illinois.edu"),c(8,8))
 ##thecluster<-rep(c("localhost","jwbowers.pol.illinois.edu"),c(12,8))
@@ -81,21 +83,24 @@ testStats <- list("SSR Test Net Full" = ssrTestStat, ## test stat including a bi
 
 
 ## For debugging on the keeling cluster
-##simsamples<-100
-##Zs<-Zs[,1:48]
+##simsamples<-10
+##Zs<-Zs[,1:10]
 testStatResults<-vector("list",length=length(testStats))
 names(testStatResults)<-names(testStats)
 for(i in 1:length(testStats)){
 	TZ<-testStats[[i]]
+	message(names(testStats)[i])
 	dotest<-dotestMaker(model=themodel,
 			    y0=uniformityData$data$y0,
 			    truth=TRUTH,
 			    TZ=TZ,
 			    thegrid=SEARCH,
 			    simsamples=simsamples)
+	##dotestcmp<-cmpfun(dotest,options=list(optimize=3))
 	clusterExport(cl,"dotest")
-	clusterSetupRNG(cl,seed=rep(1,6)) ## use same stream of randomness for each test statistic
[…]

### 2014-10-16 figures/teststat-parallel-summary.R
Fixing names of things

@@ -37,19 +37,19 @@ oneDplot<-function(x,f,...){
 
 
 ## Power comparison holding tau fixed
-par(mfrow=c(1,3),oma=rep(0,4),mgp=c(1.5,.5,0),pty="s")
+par(mfrow=c(1,2),oma=rep(0,4),mgp=c(1.5,.5,0),pty="s")
 
 oneDplot(x="beta",f=res$tau==TRUTH$tau,xlab="Beta")
-oneDplot(x="beta",f=res$tau==min(SEARCH$tau),xlab="Beta")
-oneDplot(x="beta",f=res$tau==max(SEARCH$tau),xlab="Beta")
+##oneDplot(x="beta",f=res$tau==min(SEARCH$tau),xlab="Beta")
+##oneDplot(x="beta",f=res$tau==max(SEARCH$tau),xlab="Beta")
 
 
 ## Power comparison holding beta fixed
-par(mfrow=c(1,3),oma=rep(0,4),mgp=c(1.5,.5,0),pty="s")
+##par(mfrow=c(1,3),oma=rep(0,4),mgp=c(1.5,.5,0),pty="s")
 
 oneDplot(x="tau",f=res$beta==TRUTH$beta,xlab="Tau")
-oneDplot(x="tau",f=res$beta==min(SEARCH$beta),xlab="Tau")
-oneDplot(x="tau",f=res$beta==max(SEARCH$beta),xlab="Tau")
+##oneDplot(x="tau",f=res$beta==min(SEARCH$beta),xlab="Tau")
+##oneDplot(x="tau",f=res$beta==max(SEARCH$beta),xlab="Tau")
 
 
 
@@ -62,7 +62,7 @@ matlist<-lapply(res[,1:(ncol(res)-2)],function(x){
 names(matlist)<-names(res)[1:(ncol(res)-2)]
 
 for(i in 1:length(matlist)){
-   pdf(file=paste("figures/",names(matlist)[[i]],".pdf",sep=""))
+   pdf(file=paste("figures/",names(matlist)[i],".pdf",sep=""))
    plot2DPower(matlist[[i]],TRUTH,main=names(matlist)[i])
    dev.off()
 }

### 2014-10-16 simulation/teststat-parallel-results.R

@@ -5,9 +5,16 @@ library(RItools,lib.loc=".libraries")
 source("simulation/setup.R")
 
 load("simulation/teststat-parallel.rda")
+names(testStatResults)<- c("SSR Test Net Full" , ## test stat including a bit of the true model
+		  "SSR Test Net Degree" , ## test stat including only network degree
+		  "SSR Test" , ## test stat ignoring network
+		  "KS Test Net Full" , ## test stat including some of the true model
+		  "KS Test Net Degree" , ## test stat including only network degree
+		  "KS Test" ## test stat ignoring network
+		  )
 
 library(abind)
-tzResultsArrSSR<-abind(testStatResults[["ssrTestStat"]],along=-1)
+tzResultsArrSSR<-abind(testStatResults[[1]],along=-1)
 
 
 simPow2<-function(lst){
@@ -20,7 +27,7 @@ powResults<-lapply(testStatResults,function(obj){ simPow2(obj) })
 
 
 
-searchParams<-expand.grid(SEARCH)
+##searchParams<-expand.grid(SEARCH)
 
 
 ##searchParams2<-testStatResults[[1]][[1]]@params

### 2014-10-16 paper/introduction.tex
Starting writing

@@
 \citet{bowers2013sutva} showed that one could use Fisher's
- randomization-based hypothesis testing framework to assess models of
- treatment propagation and spillover across social networks. Working in a
- frequentist framework, they showed that one could make statements such as:
- "Our design provides strong evidence against the idea that the treatment
- spilled over by more than 5 points given this model of treatment
- propagation." This research note shows that the test statistic that they used
- --- the Kolmogorov-Smirnov test statistic  \citep[\S
- 5.4]{MylesHollander1999a} --- could be improved substantially. We replicate
- their analyses to demonstrate that the sum of squared residuals from a linear
- model in which aspects of the network itself play a role.
+randomization-based hypothesis testing framework to assess counterfactual
+causal models of treatment propagation and spillover across social networks.
+That is, they showed that one could make
+statements such as: ``Our design provides strong evidence against the idea
+that the treatment spilled over by more than 5 points given this model of
+treatment propagation.'' This research note shows that the test statistic that
+they used --- the Kolmogorov-Smirnov test statistic  \citep[\S
+5.4]{MylesHollander1999a} --- could be improved substantially. We demonstrate that the sum of squared residuals from a linear
+model in which aspects of the network play a role is a preferable test
+statistic for the model and network and design analyzed by
+\citet{bowers2013sutva} (henceforth BFP). The point of this short article is to improve the
+use of the ``reasoning about interference'' approach and also to highlight
+the methodological and statistical questions raised by that article. We do
+not aim for general answers to those questions here and now, but we do hope
+that this piece stimulates further, and  deeper, work on test statistics and
+sharp hypotheses.
 
- They also showed that the strength of evidence against the
- specific hypotheses implied by a given model varied with different features
- of the design as well as the extent to which the true process by which
- treatment given to one unit changed the outcomes of another unit diverged
- from the model. Throughout their paper they used the  Kolmogorov-Smirnov (KS)
- test statistic so that their tests would be sensitive differences in the
- treatment and control distributions implied by different hypotheses and not
- merely sensitive to differences in one aspect of those distributions (such as
- the differences in the mean).\footnote{If the empirical cumulative distribution function (ECDF) of
-   the treated units is $F_1$ and the ECDF of the control units is $F_0$ then
-   the KS test statistic is
-   $\T(\yu,\bz) = \underset{i = 1,\ldots,n}{\text{max}}\left[F_1(y_{i,\bzero})
-     - F_0(y_{i,\bzero})\right]$, where $F(x)=(1/n)\sum_{i=1}^n I(x_i \le x)$
+\section{Background}
+
+We can define a counterfactual causal effect for a person $i$ as a difference
+between the outcome, $y_i$, person $i$ would have expressed
+after a set of treatments, collected in the vector $\bz$, had been assigned to some subset of a social network,
+$y_{i,\bz}$ compared to the outcome that person $i$ would have expressed
+under a different treatment vector. If the experiment had an effect on person
+$i$, then her outcome would differ under different realizations of the
+experimental treatment $y_{i,\bz} \ne y_{i,\bz'}$. The fundamental problem of
+causal inference (a phrase famously coined in \cite{holland:1986a}) reminds
+us that we can never see both states of the world: we only observe the
+outcome from person $i$ under one treatment assignment, either $\bz$ or
+$\bz'$ not both.
+
+R.A. Fisher's approach to design-based statistical inference  \citep[Chap
+2]{fisher:1935} begins with the idea that we do not learn directly about the
+unobserved counterfactuals. Rather, he suggests, we can learn about models of
+unobserved counterfactuals --- we can learn how much information we have to
+dispel certain claims about those counterfactuals as implausible. This
+conceptual move --- sidestepping the fundamental problem of causal inference
+via learning about claims made by scientists --- drives hypothesis testing in
+general.  BFP build on this insight as developed and
+linked to counterfactual causal inference by Paul Rosenbaum
+\citep{rosenbaum2010design}. They also say: we do not know how the
+counterfactual conditions might have turned out, they are unobserved, but we
+can make claims about how treatment might have come to have an effect and we
+can assess those claims. For example, the claim used by
+BFP is a model in which the effects of treatment
+propagating from one person to another die off as the network distance between
+nodes increases.\footnote{See the paper itself for details of the example
+  model.} They further showed that the strength of evidence against the
+specific hypotheses implied by a given model varied with different features of
+the design as well as the extent to which the true process by which treatment
+given to one unit changed the outcomes of another unit diverged from the
+model. Throughout their paper they used the Kolmogorov-Smirnov (KS) test
+statistic so that their tests would be sensitive differences in the treatment
+and control distributions implied by different hypotheses and not merely
+sensitive to differences in one aspect of those distributions (such as the
+differences in the mean).\footnote{If the empirical cumulative distribution
+  function (ECDF) of the treated units is $F_1$ and the ECDF of the control
+  units is $F_0$ then the KS test statistic is $\T(\yu,\bz) = \underset{i =
+    1,\ldots,n}{\text{max}}\left[F_1(y_{i,\bzero}) -
+    F_0(y_{i,\bzero})\right]$, where $F(x)=(1/n)\sum_{i=1}^n I(x_i \le x)$
   records the proportion of the distribution of $x$ at or below $x_i$
@@
 
- provides both clarity for
- non-parametric statistical inference  also implied great flexibility in regard to the scientific questions
-that one might engage: for example, one could use the sharp null itself to
-detect spillover  among units \citep{aronow-general} or to characterize the
-benefits from treatment on treated units even if  and sharp hypotheses
-about effects (in contrast to hypotheses about no effects) have been developed
-to assess models of interference \citep{bowers2013sutva}, [ include Rosenbaum
-2007, Bowers and Hansen 2009, and a bunch of other stuff by Rosenbaum ].
-
-Fisher's sharp null hypothesis of no effects combined with his
-randomization-based hypothesis testing framework provides a very clear and
-clean vision of what it means to test a hypothesis \citep[Chap
-2]{fisher:1935} and \citep{rosenbaum2010design}.
-One problem that has become apparent, as scholars have extended Fisher's basic
-ideas, is that as the models of effects have become more complex, the power of
-the testing procedure itself has become more important. A simple test of the
+Notice that the test statistic choice matters: the engine of statistical
+inference involves summarizing information against the hypothesized claim, yet
+different ways to summarize information might be more or less sensitive to
+substantively meaningful differences. A simple test of the
 sharp null can be executed with power that varies as a function of the design
@@
 of the outcome, and the way that a test statistic summarizes the outcome. Yet,
-when the models are complex,  \citet{bowers2013sutva} showed that sometimes a
-test will have (1) no power to address the model at all (a function, in part
-of the model itself --- for example ...) such that all hypothesized parameters
+when the models are complex,  BFP showed that sometimes a
+test will have (1) no power to address the model at all such that all hypothesized parameters
 would receive the same implausibility assessment; (2) or might reject all such
@@
 
-In this paper we use a very simple causal model and a simple test statistic to
-first show how the combination of model and test statistic is a crucial choice
---- and how more attention ought to be paid to both the model and the test
-statistic.
+BFP recommends choosing ``effect increasing" test statistics ``that will be
+small when the treated and control distributions in the adjusted data ... are
+similar, and large when the distributions diverge." However, test statistics
+that only compare distributional features of hypothesized potential outcomes
+between individuals assigned to treatment and those assigned to control fail
+to fully exploit the data at hand. The problem is that these test statistics
+are  not necessarily increasing in spillover effects.
 
-Then we suggest one test statistic that we think should have good properties
-when the number of parameters in a model is low (i.e. $k << n$). And we should
-this to be the case --- that the robust F-stat from a linear model is a better
-way to go than the KS-test and certainly better than any single measure of
-distributional difference.
+I demonstrate two results, each showing how effect increasing test statistics,
+as defined by BFP, may provide little to no information about the spillover
+parameter $\tau_2$.  Let us first consider the case when the test statistic is
+the absolute mean difference between hypothesized uniformity trial outcomes
+for treated and control individuals: $\T(\widetilde y_{ij}(\0,\Z), \Z') =
+\left\vert \mu_1 ( \widetilde y_{ij}(\0,\Z), \Z' ) - \mu_0 ( \widetilde
+  y_{ij}(\0,\Z), \Z' ) \right\vert.  $ Fix a realization of $\Z = \z$, and
+further assume noncollinearity: $ \z \notin \{\z^I, (1-\z^I)\}$. We can see
+that, for any hypothesized $\widetilde\tau_2$, there always exists a
+$\widetilde\tau_1$ such that the $p$-value is 1. Expanding the test statistic,
+$\T(\widetilde y_{ij}(\0,\z), \z) = \left\vert \mu_1 (y_{ij}(\z) -
+  \widetilde\tau_1 - \tau_2 z_{i(-j)}, \z) - \mu_0(\widetilde y_{ij}(\z) -
+  \widetilde\tau_2 z_{i(-j)} , \z) \right\vert = \left\vert \mu_1 (y_{ij}(\z)
+  - \widetilde\tau_2 z_{i(-j)}, \z)- \mu_0 (\widetilde y_{ij}(\z) -
+  \widetilde\tau_2 z_{i(-j)}, \z)- \widetilde\tau_1 \right\vert.  $ Thus,
+regardless of the value of $\widetilde\tau_2$, we can always find a
+$\widetilde\tau_1$ that sets $\T(\widetilde y_{ij}(\0,\z), \z) = 0$. Given
+that $\T(\widetilde y_{ij}(\0,\z), \Z')$ can only take values in
+$\mathbb{R}^+$, then $p=1$, as any other possible randomization $\Z'$ will
+yield a value of $\T(\widetilde y_{ij}(\0,\z), \Z') \geq 0$. It follows that,
+given this test, we cannot rule out any value of $\tau_2$ as being
+implausible.\footnote{In fact, this result does not depend on example in the
+  setting, and holds for any randomization scheme coupled with a linear causal
+  model of the form \eqref{causal}.}
 
-Finally, we highlight the similarities between testing the sharp hypothesis
-and the weak hypothesis --- which in turn allows even more flexibility and
-perhaps better performing testing procedures.
+Second, I show that, given the running example, a test based on any effect
+increasing statistic is inconsistent against a radically incorrect hypothesis
+about $\tau_2$. Let us assume that $\tau_1$ is properly hypothesized at its
+true value, and that the hypothesized value of $\tau_2$ ($\widetilde \tau_2$)
+is radically incorrect: $\vert \widetilde\tau_2 - \tau_2 \vert \gg
+\max(y_{ij}(\0)) - \min(y_{ij}(\0))$. Denote $\beta \equiv \tau_2 -
+\widetilde\tau_2$, and assume $\vert \beta \vert < \infty$.  Under the
+assumption that $\tau_2 = \widetilde \tau_2$: $\widetilde y_{ij}(\0,\Z) =
+y_{ij}(\0) + \beta Z_{i(-j)}$.  Assume a sequence of nested finite populations
+$\{U_k\} = U_1, U_2,...$, growing in $k$ households, so that $U_k$ consists of
+$N_k = 2k$ individuals, each with an associated random assignment $\Z_k$
+\citep{isaki1982survey}. (For notational parsimony, I drop explicit dependence
+on the subscript $k$.)  Given an incorrect hypothesis, a necessary, but not
+sufficient, condition for a test's consistency is that $\lim_{k \rightarrow
+\infty} \Pr(p > \alpha) < \epsilon$, for any $\alpha,\epsilon > 0$.
 
-Then we suggest a few alternative avenues to pursue.
+Let an effect increasing test statistic $\T(\widetilde y_{ij}(\0,\Z),\Z')$ be
+some measure of the absolute divergence in empirical distributions of
+$\widetilde y_{ij}(\0,\Z)$ for treated and control individuals under the
+treatment vector $\Z'$.  Since distributional differences in $y_{ij}(\0)$ are
+small relative to shifts induced by differences in proportions of treated
+housemates,  $\T(\widetilde y_{ij}(\0,\Z),\Z')$ will be increasing in $
+\left\vert \mu_1(Z_{i(-j)},\Z') -   \mu_0(Z_{i(-j)},\Z') \right\vert$. , For
+a realization $\Z = \z$, the hypothesized randomization distribution of
+$\mu_1(z_{i(-j)},\Z') - \mu_0(z_{i(-j)},\Z')$ is asymptotically normal with
+mean $0$ and variance $V_{x(\z)} = 4 x(\z) \left[1-x(\z)\right] / N$, where
+$x(\z) \equiv \sum_{ij\in U}z _{ij} /N$, if $0 < x(\z) <  1$.  (by the
+Lindeberg CLT, Taylor linearization and Slutsky's theorem).  Thus the
+$p$-value associated with randomization $\z$, $p = 2
+\Phi(-|\mu_1(z_{i(-j)},\z) - \mu_0(z_{i(-j)},\z)|/V_{x(\z)}^{1/2})$. It
+follows that $|\mu_1(z_{i(-j)},\z) - \mu_0(z_{i(-j)},\z)\vert <
+-\Phi^{-1}(\alpha/2) V_{x(\z)}^{1/2} \Leftrightarrow p > \alpha$.  By similar
+calculations at the household level, the (actual, not hypothesized)
+randomization distribution of $\mu_1(Z_{i(-j)},\Z) - \mu_0(Z_{i(-j)},\Z)$ is
+asymptotically mean $0$ with variance $2 V_{1/2} =  2/N$.\footnote{This result
+  follows from reframing the problem as randomly assigning households into 4
+  conditions \{00,01,10,11\} with equal probability, then linearizing
+  $\mu_1(Z_{i(-j)},\Z) - \mu_0(Z_{i(-j)},\Z)$.} By the LLN ,  $\plim_{k
+\rightarrow \infty} x = 1/2$, and so, by the and continuous mapping theorem,
+$\lim_{k \rightarrow \infty} \Pr(\vert NV_{x(\Z)} - NV_{1/2}\vert >
+\epsilon_3) = 0$.  Then, by Chebyshev's inequality, $\lim_{k \rightarrow
+\infty}  \Pr(p > \alpha) > 1- 2\left[\Phi^{-1}(\alpha/2)\right]^{-2}$,
+$\epsilon = 1- 2\left[\Phi^{-1}(\alpha/2)\right]^{-2}$ proportion of possible
+$\Z \in \Omega$ have $p > \alpha$,
+for any $0 < \alpha < 2\Phi(-\sqrt{2})$.  Tests based on effect increasing
+statistics are therefore not generally consistent.
 
-Some models are well powered for some designs and not other designs.
+\subsection*{Alternative test statistics}
 
-Using same simulation study using same data from PA study. Better than KS
-test.
+There are usually better ways to evaluate the fit of a model than comparing
+two marginal distributions. One simple test statistic would be the
+sum-of-squared-residuals (SSR) from a least squares regression of $\widetilde
+y_{ij}(\0,\Z)$ on $Z_{ij}$ and $Z_{i(-j)}$. Since the model in \eqref{causal}
+is linear, $\T(\widetilde y_{ij}(\0,\Z), \Z) = \sum_{ij \in U} \left[
+  \widetilde \Y(\0) - (\X^T\X)^{-1}\X^T \widetilde \Y(\0) \right]^2$, where $$
+\widetilde \Y(\0) \equiv \left( \begin{array}{c} \widetilde y_{11}(0) \\
+    \vdots \\ y_{(N/2)(-1)}(0) \end{array} \right), \X \equiv  \left(
+  \begin{array}{ccc} 1 & Z_{11} & Z_{1(-1)} \\ \vdots & \vdots & \vdots  \\ 1
+    & Z_{(N/2)(-1)} & Z_{(N/2)1} \end{array} \right) = ({\bf 1} \; \Z \;
+\Z^I).  $$ Assume now that either the hypothesized $\tau_1$ or $\tau_2$ is
+radically incorrect: i.e., $\vert \widetilde\tau_1 - \tau_1 \vert$ or $\vert
+\widetilde\tau_2 - \tau_2 \vert \gg \max(y_{ij}(\0)) - \min(y_{ij}(\0))$.
+Fixing $\Z = \z$, for any randomization $\Z' \notin \{\z, (1-\z), \z^I,
+(1-\z^I), \0, \one \}$, $\T(\widetilde y_{ij}(\0,\z),\Z') > \T(\widetilde
+y_{ij}(\0,\z),\z)$. Assuming noncollinearity,
+%$\Z \notin \{\Z^I, (1-\Z^I)\}$, , (0,...,0), (1,...,1)\}$)
+$p \leq 6  \left( 2^{-N} \right)$, rejecting the hypothesis with $\alpha =
+0.05$ for any $N \geq 8$. %Consistency follows straightforwardly.
 
-The F-test is model consistent. By design it controls Type I error, but if the
-model is correct, it has some optimality criterion. If you had control
-variables, for example, then it would be consistent against mean differences.
+In the example above, use of SSR as a test statistic has maximal power to
+reject radically incorrect hypotheses, whereas the test statistics proposed by
+BFP are inconsistent. But SSR from an ordinary least squares regression is not
+always appropriate: for example, when the probability of exposure to spillover
+is heterogeneous across individuals, we may wish to apply inverse probability
+weights so as to ensure representative samples of potential outcomes. This
+suggests a conjecture: that the $SSR$ from an {\it
+  inverse-probability-weighted} least squares regression is more generally a
+sensible test statistic for models that include interference.  (The nature of
+these weights will be discussed when discussing alternative modes of
+inference.) Additionally, when nonlinear deviations from model predictions are
+of concern, a weighted variant of the Brownian distance covariance
+\cite{szekely2009brownian} or other {\it E}-statistic may be more sensible
+than SSR.
 
-Under the model, model consistent.
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.9\textwidth]{twoDplots.pdf}
+\caption{Proportion of $p$-values less than .05 for tests of joint hypotheses
+about $\tau$ and $\beta$. Darker values mean rare rejection. White means
+rejection always. Truth is shown at the intersection of the straight
+lines  $\tau=.5, \beta=2$. Each panel shows a different test statistic. The
+SSR Tests refer to eq, the KS tests refer to eq.  }\label{fig:ks}
+\end{figure}
+
+Test statistic selection is not trivial and BFP's contribution has effectively
+opened a new line of research. Via counterexample, we can see that effect
+increasing test statistics are not always appropriate for assessing hypotheses
+about interference. BFP is correct in the assertion that, regardless of the
+choice of test statistic selection, a set of implausible hypotheses is
+identified by the procedure. But we should not be led to believe that, for any
+given test statistic, that some hypotheses are more plausible than others.
+Such inferences -- comparing hypotheses -- may depend on the test statistic
+used, and not necessarily reflect the plausibility of the model at hand. This
+point will gain importance as we now consider the limitations of sharp null
+hypotheses.

### 2014-10-16 code/plotting.R

@@ -7,13 +7,13 @@ library(igraph)
 library(colorspace)
 
 plotexperiment <- function(s, z, mode = "undirected",
-       vertex.shape = ifelse(z, "circle", "square"),
-       layout = layout.fruchterman.reingold,
-       vertex.size = 6,
-       vertex.color = ifelse(z, "#252525", "#A0A0A0"),
-       vertex.label = NA,
-       edge.color = "black",
-       ...) {
+                           vertex.shape = ifelse(z, "circle", "square"),
+                           layout = layout.fruchterman.reingold,
+                           vertex.size = 6,
+                           vertex.color = ifelse(z, "#252525", "#A0A0A0"),
+                           vertex.label = NA,
+                           edge.color = "black",
+                           ...) {
 
   tmp <- graph.adjacency(s, mode = mode)
 
@@ -42,8 +42,8 @@ make.filled.contour.plot<-function(mat,main,levs=10,col=rev(terrain_hcl(levs)),
                  nrow=length(unique(mat[,params[1]])),
                  ncol=length(unique(mat[,params[2]])),
                  dimnames=list(sort(unique(mat[,params[1]])),
-                   sort(unique(mat[,params[2]]))
-                   )
+                               sort(unique(mat[,params[2]]))
+                               )
                  )
   theps<-sort(unique(as.numeric(mat.p)))
   p.line.levels<-quantile(theps,seq(0,1,length=levs) )
@@ -76,16 +76,16 @@ make.filled.contour.plot<-function(mat,main,levs=10,col=rev(terrain_hcl(levs)),
         #expression(paste(main,"( Type I Error Rate=",round(p.truth,3), " for ", alpha = (1/20))),
         col=col,...)
   if(length(theps)>1){
-  contour(list(x=parm1.values,y=parm2.values,z=mat.p),
-          ##nlevels=levs,
-          levels=thelevels,
-          labels=thelabels,
-          col="gray",
-          method="flattest",
-          labcex=.7,
-          vfont=c("sans serif","bold"),
-          add=TRUE)
-}
+    contour(list(x=parm1.values,y=parm2.values,z=mat.p),
+            ##nlevels=levs,
+            levels=thelevels,
+            labels=thelabels,
+            col="gray",
+            method="flattest",
+            labcex=.7,
[…]

### 2014-10-16 figures/teststat-parallel-summary.R

@@ -61,11 +61,28 @@ matlist<-lapply(res[,1:(ncol(res)-2)],function(x){
 		   })
 names(matlist)<-names(res)[1:(ncol(res)-2)]
 
+thetitles<-c("SSR Test w/ Treated Neighbors and Degree",
+	     "SSR Test w/ Degree",
+	     "SSR Test w/o Network Information",
+	     "KS Test w/ Treated Neighbors and Degree",
+	     "KS Test w/ Degree",
+	     "KS Test w/o Network Information")
+names(thetitles)<-names(matlist)
+
 for(i in 1:length(matlist)){
    pdf(file=paste("figures/",names(matlist)[i],".pdf",sep=""))
-   plot2DPower(matlist[[i]],TRUTH,main=names(matlist)[i])
+   par(mfrow=c(1,1),mgp=c(1.5,.5,0),oma=rep(0,4))
+   plot2DPower(matlist[[i]],TRUTH,main=thetitles[names(matlist)[i]])
    dev.off()
 }
+
+
+pdf(file="figures/twoDplots.pdf",width=8,height=6)
+par(mfrow=c(2,3),mgp=c(1.5,.5,0),oma=rep(0,4),pty="s",mar=c(3,3,1,.5))
+for(i in 1:length(matlist)){
+   plot2DPower(matlist[[i]],TRUTH,main=thetitles[names(matlist)[i]],cex.lab=1,cex.main=.9)
+}
+dev.off()
 
 system("touch figures/teststat2dfigs.txt")
 

### 2014-10-17 paper/introduction.tex
More work on drafting the piece

@@
 
-BFP recommends choosing ``effect increasing" test statistics ``that will be
-small when the treated and control distributions in the adjusted data ... are
-similar, and large when the distributions diverge." However, test statistics
-that only compare distributional features of hypothesized potential outcomes
-between individuals assigned to treatment and those assigned to control fail
-to fully exploit the data at hand. The problem is that these test statistics
-are  not necessarily increasing in spillover effects.
+How should one choose a test statistic to assess relationships between complex
+models of treatment propagation and data? General advice for test statistics
+is that they should be powerful against relevant alternatives (find canonical
+cite). \citet[\S 2.4.4]{rosenbaum:2002} provides more specific advice about
+large sample performance of certain classes of test statistics and BFP repeat
+his general advice: `` Select a test statistic T that will be small when the
+treated and control distributions in the adjusted data \ldots are
+similar, and large when the distributions diverge.'' \citep[p.
+107]{bowers2013sutva}. They should have noted, instead, that when models
+involve more than one parameter, such as the models of treatment propagation across
+networks, test statistics should be increasing in \emph{all} of the parameters
+(for example, in both parameters capturing spillover and parameters capturing
+direct experience with the treatment). Here, we also show a benefit from
+taking into account the network structure in the test statistic as well as in
+the model.
 
-I demonstrate two results, each showing how effect increasing test statistics,
+We demonstrate two results, each showing how effect increasing test statistics,
 as defined by BFP, may provide little to no information about the spillover
@@
 $\widetilde\tau_1$ such that the $p$-value is 1. Expanding the test statistic,
-$\T(\widetilde y_{ij}(\0,\z), \z) = \left\vert \mu_1 (y_{ij}(\z) -
-  \widetilde\tau_1 - \tau_2 z_{i(-j)}, \z) - \mu_0(\widetilde y_{ij}(\z) -
-  \widetilde\tau_2 z_{i(-j)} , \z) \right\vert = \left\vert \mu_1 (y_{ij}(\z)
+
+\begin{flalign}
+  \T(\widetilde y_{ij}(\0,\z), \z) = & \left\vert \mu_1 (y_{ij}(\z) - \widetilde\tau_1 - \tau_2 z_{i(-j)}, \z) - \mu_0(\widetilde y_{ij}(\z) -
+  \widetilde\tau_2 z_{i(-j)} , \z) \right\vert \\
+= & \left\vert \mu_1 (y_{ij}(\z)
   - \widetilde\tau_2 z_{i(-j)}, \z)- \mu_0 (\widetilde y_{ij}(\z) -
-  \widetilde\tau_2 z_{i(-j)}, \z)- \widetilde\tau_1 \right\vert.  $ Thus,
+  \widetilde\tau_2 z_{i(-j)}, \z)- \widetilde\tau_1 \right\vert.
+\end{flalign}
+
+Thus,
 regardless of the value of $\widetilde\tau_2$, we can always find a
@@
 
+As an example of the performance of these new statistics, we re-analyze the
+model and design from BFP. Recall that their model of treatment propagation
+was:
+
+\begin{equation}
+\HH(\by_\bz, \bw, \beta, \tau) =
+ \frac{\beta + (1 - w_i) (1 - \beta) \exp(- \tau^2 \bw^{T} \bS)}
+      {\beta + (1 - z_i) (1 - \beta) \exp(- \tau^2 \bz^{T} \bS)} \by_\bz
+\label{eq:spillovermodelA}
+\end{equation}
+
+Briefly, the model posits that treatment effects can depend on either direct
+assignment to treatment ($\bz$) governed by $\beta$ or spillover as an
+increasing (but flattening) function of the number of directly connected
+treated neighbors ($\bz^{T} \bS$) and is governed by $\tau$. So, we have a
+model with two parameters. The network used by BFP involves 256 nodes
+connected in an undirected, random graph with node degree ranging from 0 to 10
+(mean degree 4, 95\% of nodes with degree between 1 and 8, five nodes with
+degree 0 [i.e. unconnected]).  Treatment is assigned to 50\% of the nodes at
+completely at random in the BFP example.
+
+The top row of figure~\ref{fig:twoD} compares the power of the SSR+Design test
+statistic (upper left panel) to versions of this statistic that either only
+include fixed node degree (SSR+Degree) , or no information about the network
+at all (SSR). For each test statistic, we tested the hypothesis
+$\tau=\tau_0,\beta=\tau_0$ by using a simulated permutation test (i.e. we used
+1000 permutations instead of all of them). We executed that test 10,000 times
+for each pair of parameters.  The proportion of $p$-values from that test less
+than .05 is plotted in Figure~\ref{fig:twoD}: darker values are fewer
+rejections, lighter values record more rejections.  All of these test
+statistics are valid --- they reject the true null of $\tau=.5,\beta=2$ no
+more than 5\% of the time at $\alpha=.05$ --- the plots are darkest in the
+area where the two lines showing the true parameters intersect. All of the
+plots have some power to reject non-true alternatives --- as we can see with
+the large white areas in all of the plots. However, only when we add
+information about the number of treated neighbors to the SSR+Degree
+statistic, do we see high power against all alternatives in the plane.
+
 \begin{figure}[h!]
 \centering
-\includegraphics[width=.9\textwidth]{twoDplots.pdf}
+\includegraphics[width=.99\textwidth]{twoDplots.pdf}
 \caption{Proportion of $p$-values less than .05 for tests of joint hypotheses
@@
 lines  $\tau=.5, \beta=2$. Each panel shows a different test statistic. The
-SSR Tests refer to eq, the KS tests refer to eq.  }\label{fig:ks}
+SSR Tests refer to eq, the KS tests refer to eq.  }\label{fig:twoD}
 \end{figure}
 
-Test statistic selection is not trivial and BFP's contribution has effectively
-opened a new line of research. Via counterexample, we can see that effect
+The bottom row of Figure~\ref{fig:twoD} demonstrates the power of the KS test.
+The bottom right hand panel shows the test used in BFP. Again all of the tests
+are valid in the sense of rejecting the truth no more than 5\% of the time
+when $\alpha=.05$ although all of these tests are conservative: the SSR based
+tests rejected the truth roughly 4\% of the 10,000 simulations but the KS
+tests rejected the truth roughly 2\% of the time.  The KS+Design and
+KS+Degree panels show the power of applying the KS test to residuals from linear
+models including network degree only (the +Degree version) or degree and also
+the number of treated neighbors (a quantity that changes for each simulation
+of the experiment). That is, where as the SSR panels used the sum of squared
+residuals after accounting for network degree and/or number of treated
+neighbors, the KS+Design and KS+Degree panels apply the KS test to the raw
+residuals. These panels show (1) that inclusion of a quantity from the true
+model (number of treated neighbors) is not enough to increase power against
+all alternatives to the level shown by the SSR+Design test statistic and (2)
+that the KS tests and the SSR tests have different patterns of power --- the
+KS tests might be less powerful in general (more darker areas on the plots),
+but our general sense from the analysis accords with the general sense about
+the KS test in the scholarly literature --- that it is relatively low powered
+(cites).
+
+We cannot say here whether the SSR+Design test will provide the best power
+against relevant alternatives for all possible models of treatment effect
+propagation and designs. However, we hope that this research note both
+improves the application of the BFP approach and raises new questions for
+research.  Via counterexample, we can see that effect
 increasing test statistics are not always appropriate for assessing hypotheses
@@
 Such inferences -- comparing hypotheses -- may depend on the test statistic
-used, and not necessarily reflect the plausibility of the model at hand. This
-point will gain importance as we now consider the limitations of sharp null
-hypotheses.
+used, and not necessarily reflect the plausibility of the model at hand.

### 2014-11-24 paper/introduction.tex
Editing

@@
-\citet{bowers2013sutva} showed that one could use Fisher's
-randomization-based hypothesis testing framework to assess counterfactual
-causal models of treatment propagation and spillover across social networks.
-That is, they showed that one could make
-statements such as: ``Our design provides strong evidence against the idea
-that the treatment spilled over by more than 5 points given this model of
-treatment propagation.'' This research note shows that the test statistic that
-they used --- the Kolmogorov-Smirnov test statistic  \citep[\S
-5.4]{MylesHollander1999a} --- could be improved substantially. We demonstrate that the sum of squared residuals from a linear
-model in which aspects of the network play a role is a preferable test
-statistic for the model and network and design analyzed by
-\citet{bowers2013sutva} (henceforth BFP). The point of this short article is to improve the
-use of the ``reasoning about interference'' approach and also to highlight
-the methodological and statistical questions raised by that article. We do
-not aim for general answers to those questions here and now, but we do hope
-that this piece stimulates further, and  deeper, work on test statistics and
-sharp hypotheses.
+\citet{bowers2013sutva} showed that one could use Fisher's randomization-based
+hypothesis testing framework to assess counterfactual causal models of
+treatment propagation and spillover across social networks.  This research
+note improves the statistical infernce presented in \citet{bowers2013sutva}
+(henceforth BFP) by substituting a test statistic based on a sum of squared
+residuals and incorporating information about the fixed network for the simple
+the Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
+that they used. The point of this short article is to incrementally improve
+the use of BFP's ``reasoning about interference'' approach. We do not offer
+results about test statistics for multi-parameter causal models on social
+networks here, but instead hope that this research note stimulates further,
+and deeper, work on test statistics and sharp hypotheses.
 
-\section{Background}
+\section{Background on randomization based statistical inference for questions
+about causal effects from social network experiments}
 
-We can define a counterfactual causal effect for a person $i$ as a difference
-between the outcome, $y_i$, person $i$ would have expressed
-after a set of treatments, collected in the vector $\bz$, had been assigned to some subset of a social network,
-$y_{i,\bz}$ compared to the outcome that person $i$ would have expressed
-under a different treatment vector. If the experiment had an effect on person
-$i$, then her outcome would differ under different realizations of the
-experimental treatment $y_{i,\bz} \ne y_{i,\bz'}$. The fundamental problem of
-causal inference (a phrase famously coined in \cite{holland:1986a}) reminds
-us that we can never see both states of the world: we only observe the
-outcome from person $i$ under one treatment assignment, either $\bz$ or
-$\bz'$ not both.
+In a randomized experiment with $n=4$ subjects connected via a fixed network,
+subject $i=1$ might respond differently to different ways that treatment is
+assigned to the whole network. When the treatment assignment vector provides
+treatment to persons 2 and 3, $\bz=\{0,1,1,0\}$ she might act one way,
+$y_{i=1,\bz=\{0,1,1,0\}}$ and when treatment is assigned to persons 3 and 4,
+$\bz=\{0,0,1,1\}$ she might act another way, $y_{i=1,\bz=\{0,0,1,1\}}$ That
+is, if the experiment had a causal effect on person $i$, then her outcome
+would differ under different realizations of the experimental treatment
+$y_{i,\bz} \ne y_{i,\bz'}$. The fundamental problem of causal inference
+reminds us that we can never see both states of the world: we only observe the
+outcome from person $i$ under one treatment assignment vector, either $\bz$ or
+some $\bz'$ not both  \cite{holland:1986a}.
 
 R.A. Fisher's approach to design-based statistical inference  \citep[Chap
-2]{fisher:1935} begins with the idea that we do not learn directly about the
-unobserved counterfactuals. Rather, he suggests, we can learn about models of
-unobserved counterfactuals --- we can learn how much information we have to
-dispel certain claims about those counterfactuals as implausible. This
-conceptual move --- sidestepping the fundamental problem of causal inference
-via learning about claims made by scientists --- drives hypothesis testing in
-general.  BFP build on this insight as developed and
-linked to counterfactual causal inference by Paul Rosenbaum
-\citep{rosenbaum2010design}. They also say: we do not know how the
-counterfactual conditions might have turned out, they are unobserved, but we
-can make claims about how treatment might have come to have an effect and we
-can assess those claims. For example, the claim used by
-BFP is a model in which the effects of treatment
-propagating from one person to another die off as the network distance between
-nodes increases.\footnote{See the paper itself for details of the example
-  model.} They further showed that the strength of evidence against the
-specific hypotheses implied by a given model varied with different features of
-the design as well as the extent to which the true process by which treatment
-given to one unit changed the outcomes of another unit diverged from the
-model. Throughout their paper they used the Kolmogorov-Smirnov (KS) test
+2]{fisher:1935} as developed by Paul Rosenbaum  \citep{rosenbaum2010design}
+begins with the premise of the fundamental problem of causal inference. Rather
+than imagine that we could learn directly about the unobservable
+counterfactual comparisons of outcomes under different treatment vectors,
+Fisher would suggest that we focus on learning about how \emph{models} of
+those counterfactual outcomes relate to what we observe. Although we do not
+know how person $i$ would have acted under all possible experimental
+scenarios, we can learn how much information we have to dispel certain claims
+or hypotheses.  This conceptual move --- sidestepping the fundamental problem
+of causal inference via learning about claims made by scientists --- drives
+hypothesis testing in general.  BFP build on this insight by showing that
+models of counterfactual effects can involve statements about how treatment
+given to one node in a social network can come to have an effect on other
+nodes. Their example model allows the effects of treatment to die off as the
+network distance between nodes increases.\footnote{See the paper itself for
+  details of the example model.} They further showed that the strength of
+evidence against the specific hypotheses implied by a given model varied with
+different features of the research design as well as the extent to which the
+true causal process diverged from the model. Since their simulated experiment
+involved only two treatments, the only observations available to evaluate the
+model were comparisons of the assigned to treatment group and the assigned to
+control group. Since their model could imply not only shifts in the mean of
+the observed treatment versus control outcome distributions, but changes in
+shape of those distributions, they used the Kolmogorov-Smirnov (KS) test
 statistic so that their tests would be sensitive differences in the treatment
@@
   records the proportion of the distribution of $x$ at or below $x_i$
-  \citep[\S 5.4]{MylesHollander1999a}.}
+  \citep[\S 5.4]{MylesHollander1999a}.} So, in broad outline, the BFP approach
+involves (1) the articulation of a model for how a treatment assignment vector
+can change outcomes for all subjects in the experiment (holding the network
+fixed) and (2) using function of comparing actually treated and control
+observations to tell us whether such a model is implausible (codified as a low
+$p$-value)  or whether we have too little information available from the data
+and design about a the model (codified as a high $p$-value). This is classic
+hypothesis testing.
 
-Notice that the test statistic choice matters: the engine of statistical
-inference involves summarizing information against the hypothesized claim, yet
-different ways to summarize information might be more or less sensitive to
-substantively meaningful differences. A simple test of the
-sharp null can be executed with power that varies as a function of the design
-of the study (proportion treated, blocking structure, etc..), characteristics
-of the outcome, and the way that a test statistic summarizes the outcome. Yet,
-when the models are complex,  BFP showed that sometimes a
-test will have (1) no power to address the model at all such that all hypothesized parameters
-would receive the same implausibility assessment; (2) or might reject all such
-hypothesized parameters. Thus, although in theory one may assess sharp
-hypotheses about multiparameter settings, in practice one may not learn much
-from such tests --- or, in the absence of simulation studies assessing the
-model and test statistic, one may be mislead.
-
-How should one choose a test statistic to assess relationships between complex
-models of treatment propagation and data? General advice for test statistics
-is that they should be powerful against relevant alternatives (find canonical
-cite). \citet[\S 2.4.4]{rosenbaum:2002} provides more specific advice about
-large sample performance of certain classes of test statistics and BFP repeat
-his general advice: `` Select a test statistic T that will be small when the
-treated and control distributions in the adjusted data \ldots are
-similar, and large when the distributions diverge.'' \citep[p.
-107]{bowers2013sutva}. They should have noted, instead, that when models
-involve more than one parameter, such as the models of treatment propagation across
-networks, test statistics should be increasing in \emph{all} of the parameters
-(for example, in both parameters capturing spillover and parameters capturing
-direct experience with the treatment). Here, we also show a benefit from
-taking into account the network structure in the test statistic as well as in
-the model.
-
-We demonstrate two results, each showing how effect increasing test statistics,
-as defined by BFP, may provide little to no information about the spillover
-parameter $\tau_2$.  Let us first consider the case when the test statistic is
-the absolute mean difference between hypothesized uniformity trial outcomes
-for treated and control individuals: $\T(\widetilde y_{ij}(\0,\Z), \Z') =
-\left\vert \mu_1 ( \widetilde y_{ij}(\0,\Z), \Z' ) - \mu_0 ( \widetilde
-  y_{ij}(\0,\Z), \Z' ) \right\vert.  $ Fix a realization of $\Z = \z$, and
-further assume noncollinearity: $ \z \notin \{\z^I, (1-\z^I)\}$. We can see
-that, for any hypothesized $\widetilde\tau_2$, there always exists a
-$\widetilde\tau_1$ such that the $p$-value is 1. Expanding the test statistic,
-
-\begin{flalign}
-  \T(\widetilde y_{ij}(\0,\z), \z) = & \left\vert \mu_1 (y_{ij}(\z) - \widetilde\tau_1 - \tau_2 z_{i(-j)}, \z) - \mu_0(\widetilde y_{ij}(\z) -
-  \widetilde\tau_2 z_{i(-j)} , \z) \right\vert \\
-= & \left\vert \mu_1 (y_{ij}(\z)
-  - \widetilde\tau_2 z_{i(-j)}, \z)- \mu_0 (\widetilde y_{ij}(\z) -
-  \widetilde\tau_2 z_{i(-j)}, \z)- \widetilde\tau_1 \right\vert.
-\end{flalign}
-
-Thus,
-regardless of the value of $\widetilde\tau_2$, we can always find a
-$\widetilde\tau_1$ that sets $\T(\widetilde y_{ij}(\0,\z), \z) = 0$. Given
-that $\T(\widetilde y_{ij}(\0,\z), \Z')$ can only take values in
-$\mathbb{R}^+$, then $p=1$, as any other possible randomization $\Z'$ will
-yield a value of $\T(\widetilde y_{ij}(\0,\z), \Z') \geq 0$. It follows that,
-given this test, we cannot rule out any value of $\tau_2$ as being
-implausible.\footnote{In fact, this result does not depend on example in the
-  setting, and holds for any randomization scheme coupled with a linear causal
-  model of the form \eqref{causal}.}
-
-Second, I show that, given the running example, a test based on any effect
-increasing statistic is inconsistent against a radically incorrect hypothesis
-about $\tau_2$. Let us assume that $\tau_1$ is properly hypothesized at its
-true value, and that the hypothesized value of $\tau_2$ ($\widetilde \tau_2$)
-is radically incorrect: $\vert \widetilde\tau_2 - \tau_2 \vert \gg
-\max(y_{ij}(\0)) - \min(y_{ij}(\0))$. Denote $\beta \equiv \tau_2 -
-\widetilde\tau_2$, and assume $\vert \beta \vert < \infty$.  Under the
-assumption that $\tau_2 = \widetilde \tau_2$: $\widetilde y_{ij}(\0,\Z) =
-y_{ij}(\0) + \beta Z_{i(-j)}$.  Assume a sequence of nested finite populations
-$\{U_k\} = U_1, U_2,...$, growing in $k$ households, so that $U_k$ consists of
-$N_k = 2k$ individuals, each with an associated random assignment $\Z_k$
-\citep{isaki1982survey}. (For notational parsimony, I drop explicit dependence
-on the subscript $k$.)  Given an incorrect hypothesis, a necessary, but not
-sufficient, condition for a test's consistency is that $\lim_{k \rightarrow
-\infty} \Pr(p > \alpha) < \epsilon$, for any $\alpha,\epsilon > 0$.
-
-Let an effect increasing test statistic $\T(\widetilde y_{ij}(\0,\Z),\Z')$ be
-some measure of the absolute divergence in empirical distributions of
-$\widetilde y_{ij}(\0,\Z)$ for treated and control individuals under the
-treatment vector $\Z'$.  Since distributional differences in $y_{ij}(\0)$ are
-small relative to shifts induced by differences in proportions of treated
-housemates,  $\T(\widetilde y_{ij}(\0,\Z),\Z')$ will be increasing in $
-\left\vert \mu_1(Z_{i(-j)},\Z') -   \mu_0(Z_{i(-j)},\Z') \right\vert$. , For
-a realization $\Z = \z$, the hypothesized randomization distribution of
-$\mu_1(z_{i(-j)},\Z') - \mu_0(z_{i(-j)},\Z')$ is asymptotically normal with
-mean $0$ and variance $V_{x(\z)} = 4 x(\z) \left[1-x(\z)\right] / N$, where
-$x(\z) \equiv \sum_{ij\in U}z _{ij} /N$, if $0 < x(\z) <  1$.  (by the
-Lindeberg CLT, Taylor linearization and Slutsky's theorem).  Thus the
-$p$-value associated with randomization $\z$, $p = 2
-\Phi(-|\mu_1(z_{i(-j)},\z) - \mu_0(z_{i(-j)},\z)|/V_{x(\z)}^{1/2})$. It
-follows that $|\mu_1(z_{i(-j)},\z) - \mu_0(z_{i(-j)},\z)\vert <
--\Phi^{-1}(\alpha/2) V_{x(\z)}^{1/2} \Leftrightarrow p > \alpha$.  By similar
-calculations at the household level, the (actual, not hypothesized)
-randomization distribution of $\mu_1(Z_{i(-j)},\Z) - \mu_0(Z_{i(-j)},\Z)$ is
-asymptotically mean $0$ with variance $2 V_{1/2} =  2/N$.\footnote{This result
-  follows from reframing the problem as randomly assigning households into 4
-  conditions \{00,01,10,11\} with equal probability, then linearizing
-  $\mu_1(Z_{i(-j)},\Z) - \mu_0(Z_{i(-j)},\Z)$.} By the LLN ,  $\plim_{k
-\rightarrow \infty} x = 1/2$, and so, by the and continuous mapping theorem,
-$\lim_{k \rightarrow \infty} \Pr(\vert NV_{x(\Z)} - NV_{1/2}\vert >
-\epsilon_3) = 0$.  Then, by Chebyshev's inequality, $\lim_{k \rightarrow
-\infty}  \Pr(p > \alpha) > 1- 2\left[\Phi^{-1}(\alpha/2)\right]^{-2}$,
-$\epsilon = 1- 2\left[\Phi^{-1}(\alpha/2)\right]^{-2}$ proportion of possible
-$\Z \in \Omega$ have $p > \alpha$,
-for any $0 < \alpha < 2\Phi(-\sqrt{2})$.  Tests based on effect increasing
-statistics are therefore not generally consistent.
+Notice that the test statistic choice matters in this process: the engine of
+statistical inference involves summarizing information against the
+hypothesized claim, yet different ways to summarize information might be more
+or less sensitive to substantively meaningful differences. The statistical
+power of a simple test of the sharp null hypothesis of no effects will vary as
+a function of the design of the study (proportion treated, blocking structure,
+etc.), characteristics of the outcome (continuous, binary, skewed, extreme
+points, etc.), and the way that a test statistic summarizes the outcome (does
+it compare means, standard deviations, medians, qqplots, etc.). In general,
+test statistics should be powerful against relevant alternatives (find
+canonical cite). \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
+advice about large sample performance of certain classes of test statistics
+and BFP repeat his general advice: `` Select a test statistic T that will be
+small when the treated and control distributions in the adjusted data \ldots
+are similar, and large when the distributions diverge.'' \citet[Proposition 4
+and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
+with this property (``effect increasing'' test statistics), produces an
+unbiased test of the hypothesis of no effects or positive effects when the
+positive effects involve one parameter.  Yet, when the models are complex and
+may involve increasing effects in the direction of one parameter and
+non-linear effects in the direction of another parameter, BFP showed that
+sometimes a KS-test will have (1) no power to address the model at all such
+that all hypothesized parameters would receive the same implausibility
+assessment; (2) or might reject all such hypothesized parameters. Thus,
+although in theory one may assess sharp hypotheses about multiparameter
+settings, in practice one may not learn much from such tests --- or, in the
+absence of simulation studies assessing the model and test statistic, one may
+be misled. BFP thus recommended simulation studies of the operating
+characteristics of tests as a piece of their workflow --- because  the theory
+justifying simple one-dimensional effect increasing test statistics clearly
+did not cover multi-parameter situations like those easily arising from social
+network experiments.
 

### 2014-11-24 paper/paper.tex
Starting writing
Editing

@@
 \author{Peter Aronow \thanks{Assistant Professor, Dept of Political Science,
-    Yale University} \and Jake Bowers \thanks{Associate Professor, Dept of Political
+    Yale University \textit{Acknowledgements:} } \and Jake Bowers \thanks{Associate Professor, Dept of Political
     Science and Statistics, University of Illinois @ Urbana-Champaign
-    (\texttt{jwbowers@illinois.edu}). \textit{Acknowledgements:} Part
-    of this work funded by NSF Grants SES-0753168 and SES-0753164
-    (joint with Ben Hansen, Dept of Statistics, University of
-    Michigan).} \and Mark Fredrickson \thanks{Ph.D. Student in Statistics and Political Science, University of Illinois @ Urbana-Champaign.}}
+    (\texttt{jwbowers@illinois.edu}).} \and Mark Fredrickson \thanks{Ph.D. Student in Statistics and Political Science, University of Illinois @ Urbana-Champaign.}}
 
@@
 
-% this appendix is the response to reviewers. TODO: remove for released
-% version
+% appendix 
 % \clearpage

### 2014-12-02 paper/introduction.tex
Draft to Peter and Mark

@@
 
+So, say $Y_i$ is the observed outcome and we hypothesize that units do not
+interfere with each other and also that $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$, we can
+assess which (if any) hypothesized values of $\tau$ appear implausible from
+the perspective of the data by: (1) Mapping the hypothesis about unobserved
+quantities using the identity $Y_i=Z_i y_{i,Z_i=1} + (1-Z_i) y_{i,Z_i=0}$ ---
+noticing that if $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ then $y_{i,Z_i=0}=Y_i - Z_i
+\tau$ (by substituting from the hypothesized relationship into the observed
+data identity); (2) Using this result to adjust the observed outcome to
+represent what would be implied by the hypothesis for given $\tau_0$ such that
+$\widetilde y_{i,Z_i=0}=Y_i - Z_i \tau_0$; and (3) Under the hypothesis,
+$\widetilde y_{i,Z_i=0}$ should have no systematic relationship with treatment
+assignment, so we can summarize this relationship with a test statistic,
+$\mathcal T(\widetilde y_{i,Z_i=0},Z_i)$ and the distribution of values for
+this test statistic arising from repetitions of treatment assignment (new
+draws of $\bz$ from all of the ways that such treatment assignment vectors
+could have been produced); and finally (4) a $p$-value arises by comparing the
+observed test statistic, $T(Y_i,Z_i)$ against the distribution of that test
+statistic that characterizes the hypothesis.
+
 Notice that the test statistic choice matters in this process: the engine of
@@
 
-\subsection*{Alternative test statistics}
+\section{Hypothesis testing as model fit assessment: The SSR test statistic}
 
-There are usually better ways to evaluate the fit of a model than comparing
-two marginal distributions. One simple test statistic would be the
-sum-of-squared-residuals (SSR) from a least squares regression of $\widetilde
-y_{ij}(\0,\Z)$ on $Z_{ij}$ and $Z_{i(-j)}$. Since the model in \eqref{causal}
-is linear, $\T(\widetilde y_{ij}(\0,\Z), \Z) = \sum_{ij \in U} \left[
-  \widetilde \Y(\0) - (\X^T\X)^{-1}\X^T \widetilde \Y(\0) \right]^2$, where $$
-\widetilde \Y(\0) \equiv \left( \begin{array}{c} \widetilde y_{11}(0) \\
-    \vdots \\ y_{(N/2)(-1)}(0) \end{array} \right), \X \equiv  \left(
-  \begin{array}{ccc} 1 & Z_{11} & Z_{1(-1)} \\ \vdots & \vdots & \vdots  \\ 1
-    & Z_{(N/2)(-1)} & Z_{(N/2)1} \end{array} \right) = ({\bf 1} \; \Z \;
-\Z^I).  $$ Assume now that either the hypothesized $\tau_1$ or $\tau_2$ is
-radically incorrect: i.e., $\vert \widetilde\tau_1 - \tau_1 \vert$ or $\vert
-\widetilde\tau_2 - \tau_2 \vert \gg \max(y_{ij}(\0)) - \min(y_{ij}(\0))$.
-Fixing $\Z = \z$, for any randomization $\Z' \notin \{\z, (1-\z), \z^I,
-(1-\z^I), \0, \one \}$, $\T(\widetilde y_{ij}(\0,\z),\Z') > \T(\widetilde
-y_{ij}(\0,\z),\z)$. Assuming noncollinearity,
-%$\Z \notin \{\Z^I, (1-\Z^I)\}$, , (0,...,0), (1,...,1)\}$)
-$p \leq 6  \left( 2^{-N} \right)$, rejecting the hypothesis with $\alpha =
-0.05$ for any $N \geq 8$. %Consistency follows straightforwardly.
+Rosenbaum style randomization inference tends to use test statistics that
+compare two distributions. Simple models imply that the distribution of the
+outcome in the control remains fixed. For example, $\widetilde
+y_{i,Z_i=0}=Y_i-Z_i \tau$ only changes the distribution of outcomes for units
+in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$
+to the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$ makes intuitive sense in this
+case, and, if $Y_i$ is Normal or at least unimodal without major outliers,
+then this test might have optimal power. The complex model used as an example
+by BFP involved adjustments to both control and treated outcomes --- some
+hypothesized parameters would cause shifts in variance, others in location.
+So, BFP proposed to use the KS-test statistic to assess the relationship
+between $\widetilde y_{i,Z_i=0,Z_{-i}=0}$ and $Z_i$ (where $Z_{-i}=0$ means
+"when all units other than $i$ are also not treated".
 
-In the example above, use of SSR as a test statistic has maximal power to
-reject radically incorrect hypotheses, whereas the test statistics proposed by
-BFP are inconsistent. But SSR from an ordinary least squares regression is not
-always appropriate: for example, when the probability of exposure to spillover
-is heterogeneous across individuals, we may wish to apply inverse probability
-weights so as to ensure representative samples of potential outcomes. This
-suggests a conjecture: that the $SSR$ from an {\it
-  inverse-probability-weighted} least squares regression is more generally a
-sensible test statistic for models that include interference.  (The nature of
-these weights will be discussed when discussing alternative modes of
-inference.) Additionally, when nonlinear deviations from model predictions are
-of concern, a weighted variant of the Brownian distance covariance
-\cite{szekely2009brownian} or other {\it E}-statistic may be more sensible
-than SSR.
+Yet, one can also think about the process of hypothesis testing as a process
+of assessing model fit, and there are usually better ways to evaluate the fit
+of a model than comparing two marginal distributions. In the case were we know
+the fixed adjacency matrix of the network, $\bS$, and where we imagine that
+network attributes (like degree) of a node play a role in the mechanism by
+which treatment propagates, the idea of assessing model fit rather than
+distribution comparison leads naturally to the the sum-of-squared-residuals
+(SSR) from a least squares regression of $ y_{i,Z_i=0}$ on $Z_{i}$ and
+$\bz^{T} \bS$ (i.e. the number of directly connected nodes assigned treatment)
+as well as the $\mathbf{1}^{T} \bS$ (i.e. the degree of the node).
 
 As an example of the performance of these new statistics, we re-analyze the
-model and design from BFP. Recall that their model of treatment propagation
+model and design from BFP. Their model of treatment propagation
 was:
@@
 
-Briefly, the model posits that treatment effects can depend on either direct
+Briefly, this model posits that treatment effects can depend on either direct
 assignment to treatment ($\bz$) governed by $\beta$ or spillover as an
@@
 
-The top row of figure~\ref{fig:twoD} compares the power of the SSR+Design test
-statistic (upper left panel) to versions of this statistic that either only
-include fixed node degree (SSR+Degree) , or no information about the network
-at all (SSR). For each test statistic, we tested the hypothesis
-$\tau=\tau_0,\beta=\tau_0$ by using a simulated permutation test (i.e. we used
-1000 permutations instead of all of them). We executed that test 10,000 times
-for each pair of parameters.  The proportion of $p$-values from that test less
-than .05 is plotted in Figure~\ref{fig:twoD}: darker values are fewer
-rejections, lighter values record more rejections.  All of these test
-statistics are valid --- they reject the true null of $\tau=.5,\beta=2$ no
-more than 5\% of the time at $\alpha=.05$ --- the plots are darkest in the
-area where the two lines showing the true parameters intersect. All of the
-plots have some power to reject non-true alternatives --- as we can see with
-the large white areas in all of the plots. However, only when we add
-information about the number of treated neighbors to the SSR+Degree
-statistic, do we see high power against all alternatives in the plane.
+We assess three versions of the SSR test statistic versus three versions of
+the KS test statistic. The first, describe above, we call the SSR+Design test
+statistic because it represents information about how treatment is assigned to
+the nodes (in $\bz^T \bS$). The second version of the SSR test statistic only
+includes network degree and excludes information about the treatment status of
+other nodes (only includes $\b1^T \bS$). And the third version includes only
+treatment assignment $\bz$. The top row of figure~\ref{fig:twoD} compares the
+power of the SSR+Design test statistic (upper left panel) to versions of this
+statistic that either only include fixed node degree (SSR+Degree), or no
+information about the network at all (SSR). For each test statistic, we tested
+the hypothesis $\tau=\tau_0,\beta=\tau_0$ by using a simulated permutation
+test (i.e. we sampled 1000 permutations instead of all of them). We executed that
+test 10,000 times for each pair of parameters.  The proportion of $p$-values
+from that test less than .05 is plotted in Figure~\ref{fig:twoD}: darker
+values show fewer rejections, lighter values record more rejections.  All of
+these test statistics are valid --- they reject the true null of
+$\tau=.5,\beta=2$ no more than 5\% of the time at $\alpha=.05$ --- the plots
+are darkest in the area where the two lines showing the true parameters
+intersect. All of the plots have some power to reject non-true alternatives
+--- as we can see with the large white areas in all of the plots. However,
+only when we add information about the number of treated neighbors to the
+SSR+Degree statistic, do we see high power against all alternatives in the
+plane.
 
@@
 tests rejected the truth roughly 4\% of the 10,000 simulations but the KS
-tests rejected the truth roughly 2\% of the time.  The KS+Design and
-KS+Degree panels show the power of applying the KS test to residuals from linear
-models including network degree only (the +Degree version) or degree and also
-the number of treated neighbors (a quantity that changes for each simulation
-of the experiment). That is, where as the SSR panels used the sum of squared
-residuals after accounting for network degree and/or number of treated
-neighbors, the KS+Design and KS+Degree panels apply the KS test to the raw
-residuals. These panels show (1) that inclusion of a quantity from the true
-model (number of treated neighbors) is not enough to increase power against
-all alternatives to the level shown by the SSR+Design test statistic and (2)
-that the KS tests and the SSR tests have different patterns of power --- the
-KS tests might be less powerful in general (more darker areas on the plots),
-but our general sense from the analysis accords with the general sense about
-the KS test in the scholarly literature --- that it is relatively low powered
-(cites).
+tests rejected the truth roughly 2\% of the time.  The KS+Design and KS+Degree
+panels show the power of applying the KS test to residuals from linear models
+including network degree only (the +Degree version) or degree and also the
+number of treated neighbors (the +Design version). That is, whereas the SSR
+panels used the sum of squared residuals after accounting for network degree
+and/or number of treated neighbors, the KS+Design and KS+Degree panels apply
+the KS test to the raw residuals after adjusting for information about the
+design and network (or with no adjustment). These panels show (1) that
+inclusion of a quantity from the true model (number of treated neighbors) is
+not enough to increase power against all alternatives to the level shown by
+the SSR+Design test statistic and (2) that the KS tests and the SSR tests have
+different patterns of power --- the KS tests might be less powerful in general
+(more darker areas on the plots), but our general sense from the analysis
+accords with the general sense about the KS test in the scholarly literature
+--- that it is relatively low powered.\footnote{For a short description and a
+few citations to the well-known low-power of the KS-test, see
+\url{https://asaip.psu.edu/Articles/beware-the-kolmogorov-smirnov-test}}
+
+\section{Discussion and Speculations}
 
@@
 against relevant alternatives for all possible models of treatment effect
-propagation and designs. However, we hope that this research note both
-improves the application of the BFP approach and raises new questions for
-research.  Via counterexample, we can see that effect
-increasing test statistics are not always appropriate for assessing hypotheses
-about interference. BFP is correct in the assertion that, regardless of the
-choice of test statistic selection, a set of implausible hypotheses is
+propagation, network topologies and designs. However, we hope that this
+research note both improves the application of the BFP approach and raises new
+questions for research.  BFP is correct in the assertion that, regardless of
+the choice of test statistic selection, a set of implausible hypotheses is
 identified by the procedure. But we should not be led to believe that, for any
@@
 Such inferences -- comparing hypotheses -- may depend on the test statistic
-used, and not necessarily reflect the plausibility of the model at hand.
+used, and not necessarily reflect the plausibility of the model at hand. That
+is, the results of any hypothesis test (or confidence interval creation) tell
+us both about the test statistic and about the causal model under scrutiny.
+
+In the example above, the SSR+Design test statistic had much better power than
+any other test statistic. But SSR from an ordinary least squares regression is
+not always appropriate: for example, when the probability of exposure to
+spillover is heterogeneous across individuals and not well captured by the
+$\bz^T \bS$ term or some other analogous term, we may wish to apply inverse
+probability weights so as to ensure representative samples of potential
+outcomes. This suggests a conjecture: that the $SSR$ from an {\it
+  inverse-probability-weighted} least squares regression is more generally a
+sensible test statistic for models that include
+interference.\footnote{\citet{aronowsamii2012interfere} use such weights for
+  unbiased estimation of network-treatment-exposure probability weighted
+  average treatment effects.}  Additionally, when nonlinear deviations from
+model predictions are of concern, a weighted variant of the Brownian distance
+covariance \cite{szekely2009brownian} or other {\it E}-statistic may be more
+sensible than SSR.

### 2015-08-04 paper/titlepage.tex
Adding an abstract to titlepage. Now in introduction.tex itself. Later delete this file

@@
 \begin{titlepage}
+
 \maketitle
@@
 \begin{abstract} 
-
+\citet{bowers2013sutva} showed that one could use Fisher's randomization-based
+hypothesis testing framework to assess counterfactual causal models of
+treatment propagation and spillover across social networks.  This research
+note improves the statistical infernce presented in \citet{bowers2013sutva}
+(henceforth BFP) by substituting a test statistic based on a sum of squared
+residuals and incorporating information about the fixed network for the simple
+the Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
+that they used. The point of this short article is to incrementally improve
+the use of BFP's ``reasoning about interference'' approach. We do not offer
+results about test statistics for multi-parameter causal models on social
+networks here, but instead hope that this research note stimulates further,
+and deeper, work on test statistics and sharp hypotheses.
 \end{abstract}

### 2015-08-04 paper/paper.tex
Adding the abstract and titlepage to paper.tex. Also struggling with weeird errors related to the acknowledgements, I think, to no avail.

@@
 \RequirePackage{snapshot}
+\usepackage{../styles/common}
 
-\author{Peter Aronow \thanks{Assistant Professor, Dept of Political Science,
-    Yale University \textit{Acknowledgements:} } \and Jake Bowers \thanks{Associate Professor, Dept of Political
-    Science and Statistics, University of Illinois @ Urbana-Champaign
-    (\texttt{jwbowers@illinois.edu}).} \and Mark Fredrickson \thanks{Ph.D. Student in Statistics and Political Science, University of Illinois @ Urbana-Champaign.}}
-
-  \date{\today}
-
-\usepackage{../styles/common}
+\author{Jake Bowers \thanks{Associate Professor, Dept of Political Science and Statistics, University of Illinois @ Urbana-Champaign (\texttt{jwbowers@illinois.edu}).} \and Mark Fredrickson \thanks{Ph.D.
+Student in Statistics and Political Science, University of Illinois @ Urbana-Champaign.} \and Peter M.
+Aronow \thanks{Assistant Professor, Dept of Political Science and Biostatistics, Yale University}}
 
@@
 \usepackage{microtype} %only useful with pdflatex
+\usepackage{fancyvrb} % for verbatim text in footnotes
 
@@
 % \fi
-
+%% \abovedisplayshortskip=0pt
+%% \belowdisplayshortskip=0pt
+%% \abovedisplayskip=0pt
+%% \belowdisplayskip=0pt
+%% 
 \begin{document}
 \VerbatimFootnotes
-\input{titlepage}
+
+\begin{titlepage}
+
+\maketitle
+
+\begin{abstract} 
+\citet{bowers2013sutva} showed that one could use Fisher's randomization-based
+hypothesis testing framework to assess counterfactual causal models of
+treatment propagation and spillover across social networks.  This research
+note improves the statistical infernce presented in \citet{bowers2013sutva}
+(henceforth BFP) by substituting a test statistic based on a sum of squared
+residuals and incorporating information about the fixed network for the simple
+the Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
+that they used. The point of this short article is to incrementally improve
+the use of BFP's ``reasoning about interference'' approach. We do not offer
+results about test statistics for multi-parameter causal models on social
+networks here, but instead hope that this research note stimulates further,
+and deeper, work on test statistics and sharp hypotheses.
+\end{abstract}
+
+Key Words: Causal effect; Interference; Randomized experiment;
+Randomization inference; Fisher's Sharp Null Hypothesis; SUTVA
+
+\end{titlepage}
+
+%\input{titlepage.tex}
 %\setstretch{1.6} %1.6 or 1.37 which I think is "true double spacing"
 %\onehalfspacing
-%%I don't think I need to double space footnotes for AJPS. Check this.
-\abovedisplayshortskip=0pt
-\belowdisplayshortskip=0pt
-\abovedisplayskip=0pt
-\belowdisplayskip=0pt
-%\tableofcontents
-
-% include stuff here!
-% include stuff here!
 
@@
 %%% mode: latex
-%%% TeX-master: "main"
+%%% TeX-master: "paper"
 %%% End:

### 2015-08-04 paper/introduction.tex
Peters edits to introduction.tex

@@
-\citet{bowers2013sutva} showed that one could use Fisher's randomization-based
-hypothesis testing framework to assess counterfactual causal models of
-treatment propagation and spillover across social networks.  This research
-note improves the statistical infernce presented in \citet{bowers2013sutva}
-(henceforth BFP) by substituting a test statistic based on a sum of squared
-residuals and incorporating information about the fixed network for the simple
-the Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
-that they used. The point of this short article is to incrementally improve
-the use of BFP's ``reasoning about interference'' approach. We do not offer
-results about test statistics for multi-parameter causal models on social
-networks here, but instead hope that this research note stimulates further,
-and deeper, work on test statistics and sharp hypotheses.
-
 \section{Background on randomization based statistical inference for questions
@@
 although in theory one may assess sharp hypotheses about multiparameter
-settings, in practice one may not learn much from such tests --- or, in the
-absence of simulation studies assessing the model and test statistic, one may
-be misled. BFP thus recommended simulation studies of the operating
+settings, in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
 characteristics of tests as a piece of their workflow --- because  the theory
@@
 includes network degree and excludes information about the treatment status of
-other nodes (only includes $\b1^T \bS$). And the third version includes only
+other nodes (only includes $\bOne^T \bS$). And the third version includes only
 treatment assignment $\bz$. The top row of figure~\ref{fig:twoD} compares the
@@
 different patterns of power --- the KS tests might be less powerful in general
-(more darker areas on the plots), but our general sense from the analysis
-accords with the general sense about the KS test in the scholarly literature
---- that it is relatively low powered.\footnote{For a short description and a
-few citations to the well-known low-power of the KS-test, see
-\url{https://asaip.psu.edu/Articles/beware-the-kolmogorov-smirnov-test}}
+(more darker areas on the plots).
 

### 2015-08-08 paper/introduction.tex
Working on final edit before submitting

@@
-\section{Background on randomization based statistical inference for questions
-about causal effects from social network experiments}
+\section{Background on randomization based statistical inference for causal effects from social network experiments}
 
@@
 outcome from person $i$ under one treatment assignment vector, either $\bz$ or
-some $\bz'$ not both  \cite{holland:1986a}.
+some $\bz'$ not both  \citep{holland:1986a,brady2008cae}.
 
@@
 2]{fisher:1935} as developed by Paul Rosenbaum  \citep{rosenbaum2010design}
-begins with the premise of the fundamental problem of causal inference. Rather
-than imagine that we could learn directly about the unobservable
-counterfactual comparisons of outcomes under different treatment vectors,
-Fisher would suggest that we focus on learning about how \emph{models} of
-those counterfactual outcomes relate to what we observe. Although we do not
-know how person $i$ would have acted under all possible experimental
-scenarios, we can learn how much information we have to dispel certain claims
-or hypotheses.  This conceptual move --- sidestepping the fundamental problem
-of causal inference via learning about claims made by scientists --- drives
-hypothesis testing in general.  BFP build on this insight by showing that
-models of counterfactual effects can involve statements about how treatment
-given to one node in a social network can come to have an effect on other
-nodes. Their example model allows the effects of treatment to die off as the
-network distance between nodes increases.\footnote{See the paper itself for
-  details of the example model.} They further showed that the strength of
-evidence against the specific hypotheses implied by a given model varied with
-different features of the research design as well as the extent to which the
-true causal process diverged from the model. Since their simulated experiment
-involved only two treatments, the only observations available to evaluate the
-model were comparisons of the assigned to treatment group and the assigned to
+begins with the premise of the fundamental problem of causal inference. Given
+the unobservability of potential outcomes, the Fisher and Rosenbaum approach
+suggests that we focus on learning about how \emph{models} of unobservable
+counterfactual outcomes relate to what we can observe. Although we do not know
+how person $i$ would have acted under all possible experimental scenarios, we
+can learn how much information we have to dispel certain claims or hypotheses.
+This conceptual move --- sidestepping the fundamental problem of causal
+inference via learning about claims made by scientists --- drives hypothesis
+testing in general.  BFP build on this insight by showing that models of
+counterfactual effects can involve statements about how treatment given to one
+node in a social network can influence other nodes. Their
+example model allows the effects of treatment to die off as the network
+distance between nodes increases.\footnote{See the paper itself for details of
+  the example model.} They further show that the strength of evidence
+against the specific hypotheses implied by a given model varies with different
+features of the research design as well as the extent to which the true causal
+process diverged from the model. Since their simulated experiment involved
+two treatments, the only observations available to evaluate the model
+were comparisons of the assigned to treatment group and the assigned to
 control group. Since their model could imply not only shifts in the mean of
@@
 shape of those distributions, they used the Kolmogorov-Smirnov (KS) test
-statistic so that their tests would be sensitive differences in the treatment
+statistic so that their tests would be sensitive to differences in the treatment
 and control distributions implied by different hypotheses and not merely

### 2015-08-08 paper/paper.tex
Working on final edit before submitting

@@
 \usepackage[letterpaper,bottom=1in,top=1in,right=1in,left=1in,includemp=FALSE]{geometry}
-\usepackage{microtype} %only useful with pdflatex
-\usepackage{fancyvrb} % for verbatim text in footnotes
+%\usepackage{microtype} %only useful with pdflatex
+%\usepackage{fancyvrb} % for verbatim text in footnotes
 
@@
 \begin{document}
-\VerbatimFootnotes
+%\VerbatimFootnotes
 
@@
 residuals and incorporating information about the fixed network for the simple
-the Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
-that they used. The point of this short article is to incrementally improve
-the use of BFP's ``reasoning about interference'' approach. We do not offer
-results about test statistics for multi-parameter causal models on social
-networks here, but instead hope that this research note stimulates further,
-and deeper, work on test statistics and sharp hypotheses.
+Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
+they used. This note incrementally improves
+the application of BFP's ``reasoning about interference'' approach. We do not
+offer general results about test statistics for multi-parameter causal models on social
+networks here, but instead hope to stimulate further,
+and deeper, work on test statistics and sharp hypothesis testing.
 \end{abstract}
@@
 %\input{titlepage.tex}
-%\setstretch{1.6} %1.6 or 1.37 which I think is "true double spacing"
+\setstretch{1.6} %1.6 or 1.37 which I think is "true double spacing"
 %\onehalfspacing

### 2015-08-08 paper/introduction.tex
Done editing the paper

@@
 counterfactual effects can involve statements about how treatment given to one
-node in a social network can influence other nodes. Their
-example model allows the effects of treatment to die off as the network
-distance between nodes increases.\footnote{See the paper itself for details of
-  the example model.} They further show that the strength of evidence
+node in a social network can influence other nodes. For example, they present
+a model that allows the effects of treatment to die off as the network
+distance between nodes increases.\footnote{We present this model later in this
+  paper in equation~\ref{eq:spillovermodelA}. See the original paper itself for more
+  details of the example model.} They also show that the strength of evidence
 against the specific hypotheses implied by a given model varies with different
 features of the research design as well as the extent to which the true causal
-process diverged from the model. Since their simulated experiment involved
-two treatments, the only observations available to evaluate the model
-were comparisons of the assigned to treatment group and the assigned to
-control group. Since their model could imply not only shifts in the mean of
-the observed treatment versus control outcome distributions, but changes in
-shape of those distributions, they used the Kolmogorov-Smirnov (KS) test
-statistic so that their tests would be sensitive to differences in the treatment
-and control distributions implied by different hypotheses and not merely
-sensitive to differences in one aspect of those distributions (such as the
-differences in the mean).\footnote{If the empirical cumulative distribution
-  function (ECDF) of the treated units is $F_1$ and the ECDF of the control
-  units is $F_0$ then the KS test statistic is $\T(\yu,\bz) = \underset{i =
+process diverged from the model. Since their simulated experiment involved two
+treatments, the only observations available to evaluate the model were
+comparisons of the assigned to treatment group and the assigned to control
+group. Since their model could imply not only shifts in the mean of the
+observed treatment versus control outcome distributions, but changes in shape
+of those distributions, they used the Kolmogorov-Smirnov (KS) test statistic
+so that their tests would be sensitive to differences in the treatment and
+control distributions implied by different hypotheses and not merely sensitive
+to differences in one aspect of those distributions (such as the differences
+in the mean).\footnote{If the empirical cumulative distribution function
+  (ECDF) of the treated units is $F_1$ and the ECDF of the control units is
+  $F_0$ then the KS test statistic is $\T(\yu,\bz)_{\text{KS}} = \underset{i =
     1,\ldots,n}{\text{max}}\left[F_1(y_{i,\bzero}) -
@@
   records the proportion of the distribution of $x$ at or below $x_i$
-  \citep[\S 5.4]{MylesHollander1999a}.} So, in broad outline, the BFP approach
-involves (1) the articulation of a model for how a treatment assignment vector
-can change outcomes for all subjects in the experiment (holding the network
-fixed) and (2) using function of comparing actually treated and control
-observations to tell us whether such a model is implausible (codified as a low
-$p$-value)  or whether we have too little information available from the data
-and design about a the model (codified as a high $p$-value). This is classic
-hypothesis testing.
+  \citep[\S 5.4]{MylesHollander1999a}. \label{fn:kstest}} So, in broad
+outline, the BFP approach involves (1) the articulation of a model for how a
+treatment assignment vector can change outcomes for all subjects in the
+experiment (holding the network fixed) and (2) use a function to comparing
+actually treated and control observations summarize whether such a model is
+implausible (codified as a low $p$-value)  or whether we have too little
+information available from the data and design about the model (codified as a
+high $p$-value). This is classic hypothesis testing applied to an experiment
+on a social network.
 
 So, say $Y_i$ is the observed outcome and we hypothesize that units do not
-interfere with each other and also that $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$, we can
+interfere and also that $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$. We can
 assess which (if any) hypothesized values of $\tau$ appear implausible from
 the perspective of the data by: (1) Mapping the hypothesis about unobserved
-quantities using the identity $Y_i=Z_i y_{i,Z_i=1} + (1-Z_i) y_{i,Z_i=0}$ ---
+quantities to observed data using the identity $Y_i=Z_i y_{i,Z_i=1} + (1-Z_i) y_{i,Z_i=0}$ ---
 noticing that if $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ then $y_{i,Z_i=0}=Y_i - Z_i
@@
 could have been produced); and finally (4) a $p$-value arises by comparing the
-observed test statistic, $T(Y_i,Z_i)$ against the distribution of that test
+observed test statistic, $\mathcal T(Y_i,Z_i)$ against the distribution of that test
 statistic that characterizes the hypothesis.
@@
 a function of the design of the study (proportion treated, blocking structure,
-etc.), characteristics of the outcome (continuous, binary, skewed, extreme
-points, etc.), and the way that a test statistic summarizes the outcome (does
-it compare means, standard deviations, medians, qqplots, etc.). In general,
-test statistics should be powerful against relevant alternatives (find
-canonical cite). \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
-advice about large sample performance of certain classes of test statistics
-and BFP repeat his general advice: `` Select a test statistic T that will be
+etc), characteristics of the outcome (continuous, binary, skewed, extreme
+points, etc), and the way that a test statistic summarizes the outcome (does
+it compare means, standard deviations, medians, qqplots, etc). In general,
+test statistics should be powerful against relevant alternatives. \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
+advice about the large sample performance of certain classes of test statistics
+and BFP repeat his general advice: ``Select a test statistic [$\mathcal{T}$] that will be
 small when the treated and control distributions in the adjusted data \ldots
-are similar, and large when the distributions diverge.'' \citet[Proposition 4
+are similar, and large when the distributions diverge.'' \cite[Proposition 4
 and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
@@
 y_{i,Z_i=0}=Y_i-Z_i \tau$ only changes the distribution of outcomes for units
-in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$
+in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=1}|Z_i=1$
 to the mean of $\widetilde y_{i,Z_i=0}|Z_i=0$ makes intuitive sense in this
@@
 between $\widetilde y_{i,Z_i=0,Z_{-i}=0}$ and $Z_i$ (where $Z_{-i}=0$ means
-"when all units other than $i$ are also not treated".
+``when all units other than $i$ are also not treated".)
 
@@
 of assessing model fit, and there are usually better ways to evaluate the fit
-of a model than comparing two marginal distributions. In the case were we know
+of a model than comparing two marginal distributions. In the case where we know
 the fixed adjacency matrix of the network, $\bS$, and where we imagine that
@@
 which treatment propagates, the idea of assessing model fit rather than
-distribution comparison leads naturally to the the sum-of-squared-residuals
-(SSR) from a least squares regression of $ y_{i,Z_i=0}$ on $Z_{i}$ and
-$\bz^{T} \bS$ (i.e. the number of directly connected nodes assigned treatment)
-as well as the $\mathbf{1}^{T} \bS$ (i.e. the degree of the node).
+closeness of distributions leads naturally to the the sum-of-squared-residuals
+(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
+$\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned treatment)
+as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node). If we
+collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix $\bX$,
+and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
+coefficients $\bbeta$ when we could define the test statistic as:
+
+\begin{equation}
+ \T(\yu,\bz)_{\text{SSR}} \equiv \sum_{i} ( \widetilde y_{i,Z_i=0} -
+ \bX\hat{\bbeta}) ^2 \label{eq:ssr}
+\end{equation}
 
@@
 (mean degree 4, 95\% of nodes with degree between 1 and 8, five nodes with
-degree 0 [i.e. unconnected]).  Treatment is assigned to 50\% of the nodes at
+degree 0 [i.e.\ unconnected]).  Treatment is assigned to 50\% of the nodes 
 completely at random in the BFP example.
@@
 We assess three versions of the SSR test statistic versus three versions of
-the KS test statistic. The first, describe above, we call the SSR+Design test
+the KS test statistic. The first, described above, we call the SSR+Design test
 statistic because it represents information about how treatment is assigned to
-the nodes (in $\bz^T \bS$). The second version of the SSR test statistic only
-includes network degree and excludes information about the treatment status of
-other nodes (only includes $\bOne^T \bS$). And the third version includes only
-treatment assignment $\bz$. The top row of figure~\ref{fig:twoD} compares the
-power of the SSR+Design test statistic (upper left panel) to versions of this
-statistic that either only include fixed node degree (SSR+Degree), or no
-information about the network at all (SSR). For each test statistic, we tested
-the hypothesis $\tau=\tau_0,\beta=\tau_0$ by using a simulated permutation
-test (i.e. we sampled 1000 permutations instead of all of them). We executed that
-test 10,000 times for each pair of parameters.  The proportion of $p$-values
-from that test less than .05 is plotted in Figure~\ref{fig:twoD}: darker
-values show fewer rejections, lighter values record more rejections.  All of
-these test statistics are valid --- they reject the true null of
-$\tau=.5,\beta=2$ no more than 5\% of the time at $\alpha=.05$ --- the plots
-are darkest in the area where the two lines showing the true parameters
-intersect. All of the plots have some power to reject non-true alternatives
---- as we can see with the large white areas in all of the plots. However,
-only when we add information about the number of treated neighbors to the
-SSR+Degree statistic, do we see high power against all alternatives in the
-plane.
+the nodes, $\bz^T \bS$. The second version of
+the SSR test statistic (SSR+Degree) only includes network degree,  $\bOne^T \bS$, and
+excludes information about the treatment status of other nodes. And the
+third version (SSR) includes only treatment assignment $\bz$. The top row of
+figure~\ref{fig:twoD} compares the power of the SSR+Design test statistic
+(upper left panel) to versions of this statistic that either only include
+fixed node degree (SSR+Degree), or no information about the network at all
+(SSR). For each test statistic, we tested the hypothesis
+$\tau=\tau_0,\beta=\tau_0$ by using a simulated permutation test: we
+sampled 1000 permutations. We executed that test
+10,000 times for each pair of parameters.  The proportion of $p$-values from
+that test less than .05 is plotted in Figure~\ref{fig:twoD}: darker values
+show fewer rejections, lighter values record more rejections.  All of these
+test statistics are valid --- they reject the true null of $\tau=.5,\beta=2$
+no more than 5\% of the time at $\alpha=.05$ --- the plots are darkest in the
+area where the two lines showing the true parameters intersect. All of the
+plots have some power to reject non-true alternatives --- as we can see with
+the large white areas in all of the plots. However, only when we add
+information about the number of treated neighbors to the SSR+Degree statistic,
+do we see high power against all alternatives in the plane.
 
-\begin{figure}[h!]
-\centering
-\includegraphics[width=.99\textwidth]{twoDplots.pdf}
-\caption{Proportion of $p$-values less than .05 for tests of joint hypotheses
-about $\tau$ and $\beta$. Darker values mean rare rejection. White means
-rejection always. Truth is shown at the intersection of the straight
-lines  $\tau=.5, \beta=2$. Each panel shows a different test statistic. The
-SSR Tests refer to eq, the KS tests refer to eq.  }\label{fig:twoD}
+\begin{figure}[h!] \centering
+  \includegraphics[width=.99\textwidth]{twoDplots.pdf} \caption{Proportion of
+    $p$-values less than .05 for tests of joint hypotheses about $\tau$ and
+    $\beta$ for the model in equation~\ref{eq:spillovermodelA}. Darker values
+    mean rare rejection. White means rejection always. Truth is shown at the
+    intersection of the straight lines  $\tau=.5, \beta=2$. Each panel shows a
+    different test statistic. The SSR Tests refer to equation~\ref{eq:ssr},
+    the KS tests refer to the expression in
+    footnote~\ref{fn:kstest}.}\label{fig:twoD}
 \end{figure}
@@
 The bottom row of Figure~\ref{fig:twoD} demonstrates the power of the KS test.
-The bottom right hand panel shows the test used in BFP. Again all of the tests
+The bottom right hand panel shows the test used in the BFP paper. Again all of the tests
 are valid in the sense of rejecting the truth no more than 5\% of the time
@@
 the SSR+Design test statistic and (2) that the KS tests and the SSR tests have
-different patterns of power --- the KS tests might be less powerful in general
+different patterns of power --- the KS tests appear be less powerful in general
 (more darker areas on the plots).
@@
 research note both improves the application of the BFP approach and raises new
-questions for research.  BFP is correct in the assertion that, regardless of
+questions for research.  BFP are correct in the assertion that, regardless of
 the choice of test statistic selection, a set of implausible hypotheses is
 identified by the procedure. But we should not be led to believe that, for any
-given test statistic, that some hypotheses are more plausible than others.
-Such inferences -- comparing hypotheses -- may depend on the test statistic
+given test statistic, that some hypotheses are universally more plausible than others.
+Such inferences --- comparing hypotheses --- may depend on the test statistic
 used, and not necessarily reflect the plausibility of the model at hand. That
 is, the results of any hypothesis test (or confidence interval creation) tell
-us both about the test statistic and about the causal model under scrutiny.
+us \emph{both} about the test statistic \emph{and} about the causal model under scrutiny.
 
@@
 not always appropriate: for example, when the probability of exposure to
-spillover is heterogeneous across individuals and not well captured by the
+spillover is heterogeneous across individuals in a way not well captured by the
 $\bz^T \bS$ term or some other analogous term, we may wish to apply inverse

### 2015-08-08 paper/paper.tex
The anonymized version for submission

@@
 
-\author{Jake Bowers \thanks{Associate Professor, Dept of Political Science and Statistics, University of Illinois @ Urbana-Champaign (\texttt{jwbowers@illinois.edu}).} \and Mark Fredrickson \thanks{Ph.D.
+%\author{Jake Bowers \thanks{Associate Professor, Dept of Political Science and Statistics, University of Illinois @ Urbana-Champaign (\texttt{jwbowers@illinois.edu}).} \and Mark Fredrickson \thanks{Ph.D.
 Student in Statistics and Political Science, University of Illinois @ Urbana-Champaign.} \and Peter M.
@@
 \begin{titlepage}
-
 \maketitle
+\thispagestyle{empty}
 

### 2015-11-10 paper/paper.tex
Revealing our names
Fixing typos

--- 2015-08-08 paper/paper.tex
@@
 treatment propagation and spillover across social networks.  This research
-note improves the statistical infernce presented in \citet{bowers2013sutva}
+note improves the statistical inference presented in \citet{bowers2013sutva}
 (henceforth BFP) by substituting a test statistic based on a sum of squared

### 2015-11-10 paper/introduction.tex
Sketched section for replication of Coppocks analysis

@@
 and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
-with this property (``effect increasing'' test statistics), produces an
+with this property (``effect increasing'' test statistics), produce an
 unbiased test of the hypothesis of no effects or positive effects when the
@@
 plots have some power to reject non-true alternatives --- as we can see with
-the large white areas in all of the plots. However, only when we add
+the large white areas in all of the plots. Only when we add
 information about the number of treated neighbors to the SSR+Degree statistic,
-do we see high power against all alternatives in the plane.
+does the plot show high power against all alternatives in the plane.
 
@@
 
+\section{Application: The Impact of Constituency Opinion on Legislative
+  Preferences}
+
+REWRITE. FIX.
+
+\citep{coppock2014information} used a version of the SSR+Design test statistic
+to show the effect of constituency opinion on legislators preferences within
+the BFP framework.  We here show how his results would have been different had
+he used the simple KS test.\footnote{In this case we also found that only
+  the SSR+Design test statistic had power in all directions and that the KS
+  statistics (+Design, +Degree, and plain) performed basically as they do in
+  figure~\ref{fig:twoD}.}
+
 \section{Discussion and Speculations}

### 2015-11-10 response.md
Sketch/structure for response to reviewers

Dear Professor Grimmer,

We really appreciate the constructive and timely feedback from you and the reviewers.

1) As Reviewer 1 suggests, how does the model perform as the assumptions
about spill overs deviate?  How sensitive are the inferences to the
particular assumptions you can make?

from peter I think we reiterate that this is about testing sharp null
hypotheses, and this will be sensitive to mean deviations w.r.t.
exposure. Any test statistic will yield valid inferences, but SSR is
sensitive to "main effects."

2) Reviewer 2 asks you to carry forward your toy example with four units
through the entire analysis.  If this isn't possible, please provide the
simplest case that illuminate the core intuition of your method.

We have done this.

3) Reviewer 2 also asks you to provide intuition about when the KS-test
would outperform the Sum of Squared Residuals.  As part of a simulation or
analysis of a simple case, demonstrating when/why the KS test outperforms
will be useful for gaining intuition about why the SSR is a generally
superior method.

We have added some discussion about the idea that the SSR is particular sensitive to differences in means.

4) In letters we have an emphasis that new methods matter when applied to
real examples.  Can you demonstrate that a case where the SSR will change
the inferences we make when applied to a real world example?

We have added a section in which we replicate the analysis in Coppock's 2014 Journal of Experimental Political Science article "Information Spillovers: Another Look at Experimental Estimates of Legislator Responsiveness".

### 2015-11-27 coppock-replication/CoppockJEPS_figure2code.R
Revert "the p-value calculation had the inequality in the wrong direction"
The test statistic is *decreasing* in evidence against the null (i.e.,
we want to look in the left tail, not the right tail of the reference
distribution).

This reverts commit 2894569d445bb979871b6b66dbaed88d03541167.

@@ -67,7 +67,7 @@ for(j in 1:length(directs)){
       ssr.sims[i] <- test.stat(pure.Y0, Z.sim)
     }
     
-    pmat.ssr[j,k] <- mean(ssr.sims >= t.obs) # probability of seeing a statistic greater than observed
+    pmat.ssr[j,k] <- mean(t.obs >= ssr.sims)
     
   }
 }

### 2015-11-27 coppock-replication/CoppockJEPS_figure2code.R
adding in KS test along with SSR for Coppock replication

@@ -27,8 +27,9 @@ exposure.obs <- CoppockJEPS$s.difs.dw
 directs <-seq(from=-.7, to=0.2, by=.025)
 indirects <-seq(from=-.7, to=0.2, by=.025)
 
-sims <- 500
+sims <- 50
 pmat.ssr <- matrix(NA, length(directs), length(indirects))
+pmat.ks  <- matrix(NA, length(directs), length(indirects))
 
 # write a function to compute exposure as a function of a random treatment variable Z
 exposure <- function(Z) {
@@ -38,10 +39,13 @@ exposure <- function(Z) {
 }
 
 # likewise, here's the test statistic
-test.stat <- function(y, z) {
+ssr.test.stat <- function(y, z) {
   e <- exposure(z)
   sum(resid(lm(y ~ z + e))^2)
 }
+
+# we get the KS test statistic from the our implemention
+source("../code/teststatistics.R")
 
 set.seed(343)
 for(j in 1:length(directs)){
@@ -59,15 +63,22 @@ for(j in 1:length(directs)){
     pure.Y0 <- Y.obs + (-1*exposure.obs*indirect.sim)
     pure.Y0[Z.obs==1] <- pure.Y0[Z.obs==1] - direct.sim
 
-    t.obs <- test.stat(pure.Y0, Z.obs)
+    t.ssr.obs <- ssr.test.stat(pure.Y0, Z.obs)
+    t.ks.obs  <- baseKSTest(pure.Y0, Z.obs)
+    
+    ssr.sims <- rep(NA, sims)
+    ks.sims  <- rep(NA, sims)
 
-    ssr.sims <- rep(NA,sims)
     for(i in 1:sims){
       Z.sim <- Z_block[,sample(1:10000, 1)]
-      ssr.sims[i] <- test.stat(pure.Y0, Z.sim)
+      ssr.sims[i] <- ssr.test.stat(pure.Y0, Z.sim)
+      ks.sims[i]  <- baseKSTest(pure.Y0, Z.sim)
     }
     
-    pmat.ssr[j,k] <- mean(t.obs >= ssr.sims)
+    # the SSR statistic is *decreasing* in evidence against the null
+    # the KS statistic is *increasing* in evidence against the null 
+    pmat.ssr[j,k] <- mean(t.ssr.obs >= ssr.sims)
+    pmat.ks[j,k]  <- mean(t.ks.obs <= ks.sims)
     
   }
[…]

### 2015-11-27 coppock-replication/CoppockJEPS_figure2code.R
Splitting out Coppock fig 2 into code and figure files, to make it easier to tweak graphics

@@ -27,7 +27,7 @@ exposure.obs <- CoppockJEPS$s.difs.dw
 directs <-seq(from=-.7, to=0.2, by=.025)
 indirects <-seq(from=-.7, to=0.2, by=.025)
 
-sims <- 50
+sims <- 500
 pmat.ssr <- matrix(NA, length(directs), length(indirects))
 pmat.ks  <- matrix(NA, length(directs), length(indirects))
 
@@ -86,61 +86,17 @@ for(j in 1:length(directs)){
 direct_breaks <- find_breaks(apply(pmat.ssr, MARGIN=1, FUN=max) >.05)
 indirect_breaks <- find_breaks(apply(pmat.ssr, MARGIN=2, FUN=max) >.05)
 
-plotPvalues <- function(pmat) {
-  graph.frame <- expand.grid(x=directs, y=indirects)
-  graph.frame$z <- as.vector(pmat)
-  col.l <- colorRampPalette(c('white', 'black'))(1000)
-  depth.breaks <- do.breaks(c(0,1), 20)
-  fig2 <- levelplot(z~x*y, graph.frame, cuts=20, col.regions=col.l,
-                    colorkey=FALSE,
-                    at=depth.breaks,
-                          ylab = "Hypothesized indirect effect",
-                          xlab = "Hypothesized direct effect",
-                          scales=list(x=list(at=round(seq(-.7, .2, by=.1), digits=1), labels=round(seq(-.7, .2, by=.1), digits=1)),
-                                      y=list(at=round(seq(-.7, .2, by=.1), digits=1), labels=round(seq(-.7, .2, by=.1), digits=1))),
-                          panel = function(...) {
-                            panel.levelplot(...)
-                            panel.abline(h = 0, lty=2)
-                            panel.abline(v = 0, lty=2)
-                            larrows(y0=-.5, y1= -.5, x0=directs[direct_breaks[1]], x1=directs[direct_breaks[2]], angle=90,code=3)
-                            larrows(x0=-.6, x1= -.6, y0=indirects[indirect_breaks[1]], y1=indirects[indirect_breaks[2]], angle=90,code=3)
-                          },
-                    legend = 
-                      list(right = 
-                             list(fun = draw.colorkey,
[…]

### 2015-11-27 coppock-replication/CoppockJEPS_figure2graphic.R

@@ -1,0 +1,41 @@
+library(lattice)
+load("fig2.rdata")
+
+n <- prod(dim(pmat.ssr))
+
+tmp <- expand.grid(x=directs, y=indirects)
+graph.frame <- rbind(tmp, tmp)
+graph.frame$z <- c(as.vector(pmat.ssr), as.vector(pmat.ks))
+graph.frame$t <- factor(c(rep("SSR", n), rep("KS", n)))
+
+col.l <- colorRampPalette(c('white', 'black'))(1000)
+
+depth.breaks <- do.breaks(c(0,1), 20)
+fig2 <- levelplot(z ~ x*y | t, graph.frame, cuts=20, col.regions=col.l,
+                  colorkey=FALSE,
+                  at=depth.breaks,
+                        ylab = "Hypothesized indirect effect",
+                        xlab = "Hypothesized direct effect",
+                        scales=list(x=list(at=round(seq(-.7, .2, by=.1), digits=1), labels=round(seq(-.7, .2, by=.1), digits=1)),
+                                    y=list(at=round(seq(-.7, .2, by=.1), digits=1), labels=round(seq(-.7, .2, by=.1), digits=1))),
+                        panel = function(...) {
+                          panel.levelplot(...)
+                          panel.abline(h = 0, lty=2)
+                          panel.abline(v = 0, lty=2)
+                          larrows(y0=-.5, y1= -.5, x0=directs[direct_breaks[1]], x1=directs[direct_breaks[2]], angle=90,code=3)
+                          larrows(x0=-.6, x1= -.6, y0=indirects[indirect_breaks[1]], y1=indirects[indirect_breaks[2]], angle=90,code=3)
+                        },
+                  legend = 
+                    list(right = 
+                           list(fun = draw.colorkey,
+                                args = list(key = list(col = col.l, at = depth.breaks),
+                                            draw = FALSE))),
+)
+  
+pdf("CoppockJEPS_figure2.pdf", width = 8, height = 4)
+print(fig2)
+dev.off()
+
+png("CoppockJEPS_figure2.png", width = 8, height = 4)
+print(fig2)
+dev.off()

### 2015-11-28 coppock-replication/CoppockJEPS_figure2code.R
using a .rdata file supplied by Alex for the replication

@@ -27,7 +27,7 @@ exposure.obs <- CoppockJEPS$s.difs.dw
 directs <-seq(from=-.7, to=0.2, by=.025)
 indirects <-seq(from=-.7, to=0.2, by=.025)
 
-sims <- 500
+sims <- 50
 pmat.ssr <- matrix(NA, length(directs), length(indirects))
 pmat.ks  <- matrix(NA, length(directs), length(indirects))
 

### 2015-11-28 coppock-replication/CoppockJEPS_figure2code.R
rolling back to an earlier version of the figure file, minimal addition of a ks test

@@ -18,67 +18,58 @@ find_breaks <- function(x){
 }
 
 load("CoppockJEPS.rdata")
+source("../code/teststatistics.R")
 
 ## Pick exposure model -- main analysis uses dw nominate
 Z.obs <- CoppockJEPS$treatment
 Y.obs <- CoppockJEPS$sb24
 exposure.obs <- CoppockJEPS$s.difs.dw
 
+fit.obs <- lm(Y.obs ~ Z.obs + exposure.obs)
+ssr.obs <- sum(residuals(fit.obs)^2)
+direct.obs <- fit.obs$coefficients[2]
+indirect.obs <- fit.obs$coefficients[3]
+
+exposure.expected.1 <- CoppockJEPS$exposure.expected.1.dw
+exposure.expected.0 <- CoppockJEPS$exposure.expected.0.dw
+
 directs <-seq(from=-.7, to=0.2, by=.025)
 indirects <-seq(from=-.7, to=0.2, by=.025)
 
 sims <- 50
 pmat.ssr <- matrix(NA, length(directs), length(indirects))
-pmat.ks  <- matrix(NA, length(directs), length(indirects))
-
-# write a function to compute exposure as a function of a random treatment variable Z
-exposure <- function(Z) {
-  exposure.observed.dw <- similarity.matrix.dw %*% Z
-  difs.dw <- exposure.observed.dw - (Z * CoppockJEPS$exposure.expected.1.dw + (1 - Z) * CoppockJEPS$exposure.expected.0.dw)
-  (difs.dw - mean(difs.dw)) / sd(difs.dw)
-}
-
-# likewise, here's the test statistic
-ssr.test.stat <- function(y, z) {
-  e <- exposure(z)
-  sum(resid(lm(y ~ z + e))^2)
-}
-
-# we get the KS test statistic from the our implemention
-source("../code/teststatistics.R")
+pmat.ks  <- matrix(1, length(directs), length(indirects))
 
 set.seed(343)
 for(j in 1:length(directs)){
   for(k in 1:length(indirects)){
-    # the proper inference algorithm is a loop over each
-    # parameter combination (directs, indirects):
-    # 1. adjust the outcome using the observed Z and the params
-    # 2. compute the test statistic using the observed Z and adjusted data
[…]

### 2015-11-28 coppock-replication/CoppockJEPS_figure2code.R
creating KS tests, also computing exposure in the figure file, instead of relying on the precomputed value for the observed exposure

@@ -33,12 +33,25 @@ indirect.obs <- fit.obs$coefficients[3]
 exposure.expected.1 <- CoppockJEPS$exposure.expected.1.dw
 exposure.expected.0 <- CoppockJEPS$exposure.expected.0.dw
 
+exposure <- function(Z) {
+  exposure.sim <- (similarity.matrix.dw %*% Z)
+  exposure.sim.corrected <- exposure.sim - (Z * exposure.expected.1 + (1-Z)*exposure.expected.0)
+  s.exposure.sim <- (exposure.sim.corrected - mean(exposure.sim.corrected))/sd(exposure.sim.corrected)
+}
+
+exposure.computed <- exposure(Z.obs)
+
+t.ssr.comp <- sum(residuals(lm(Y.obs ~ Z.obs + exposure.computed))^2)
+
 directs <-seq(from=-.7, to=0.2, by=.025)
 indirects <-seq(from=-.7, to=0.2, by=.025)
 
 sims <- 50
 pmat.ssr <- matrix(NA, length(directs), length(indirects))
-pmat.ks  <- matrix(1, length(directs), length(indirects))
+pmat.ks  <- matrix(NA, length(directs), length(indirects))
+
+pmat.ssr.comp <- matrix(NA, length(directs), length(indirects))
+pmat.ks.comp  <- matrix(NA, length(directs), length(indirects))
 
 set.seed(343)
 for(j in 1:length(directs)){
@@ -49,27 +62,39 @@ for(j in 1:length(directs)){
     pure.Y0 <- Y.obs - Z.obs * direct.sim - exposure.obs * indirect.sim
     t.ks.obs <- baseKSTest(pure.Y0, Z.obs)
 
+    Y0.comp <- Y.obs - Z.obs * direct.sim - exposure.computed * indirect.sim
+    t.ks.comp  <- baseKSTest(Y0.comp, Z.obs)
+
     ssr.sims <- rep(NA, sims)
     ks.sims  <- rep(NA, sims)
+    ssr.comp.sims <- rep(NA, sims)
+    ks.comp.sims  <- rep(NA, sims)
 
     for(i in 1:sims){
       Z.sim <- Z_block[,sample(1:10000, 1)]
 
-      exposure.sim <- (similarity.matrix.dw %*% Z.sim)
-      exposure.sim.corrected <- exposure.sim - (Z.sim * exposure.expected.1 + (1-Z.sim)*exposure.expected.0)
[…]

### 2015-11-28 coppock-replication/CoppockJEPS_figure2graphic.R

@@ -4,9 +4,17 @@ load("fig2.rdata")
 n <- prod(dim(pmat.ssr))
 
 tmp <- expand.grid(x=directs, y=indirects)
-graph.frame <- rbind(tmp, tmp)
-graph.frame$z <- c(as.vector(pmat.ssr), as.vector(pmat.ks))
-graph.frame$t <- factor(c(rep("SSR", n), rep("KS", n)))
+graph.frame <- rbind(tmp, tmp, tmp, tmp)
+
+graph.frame$z <- c(as.vector(pmat.ssr),
+                   as.vector(pmat.ks),
+                   as.vector(pmat.ssr.comp),
+                   as.vector(pmat.ks.comp))
+
+graph.frame$t <- factor(c(rep("SSR (Coppock)", n),
+                          rep("KS (Coppock)", n),
+                          rep("SSR (Computed)", n),
+                          rep("KS (Computed)", n)))
 
 col.l <- colorRampPalette(c('white', 'black'))(1000)
 

### 2015-11-30 response.md
First draft of response letter

@@
 
-from peter I think we reiterate that this is about testing sharp null
-hypotheses, and this will be sensitive to mean deviations w.r.t.
-exposure. Any test statistic will yield valid inferences, but SSR is
-sensitive to "main effects."
+In the paper we clarified that we are focusing on testing sharp null
+hypotheses arising from one toy model of how an experimental treatment might
+propagate across a social network. We have added some discussion to the paper
+highlighting the fact that the power of a test of a sharp null hypothesis
+depends on *both* the model of causal effects *and* the test statistic (and
+the effective sample size which itself depends on sample size and network
+characteristics). We also note that, given randomization, any test statistic
+ought to produce valid statistical inferences (i.e. tests that would reject
+the truth no more than $100 \alpha \%$ of the time).
+
+ Since the point of the paper is to learn about test statistics, so, in this
+paper we do not explore alternative models of causal effects. The original
+paper (BFP) on this method compared the performance of tests when the data did
+and did not arise from the causal model and we point to that paper for those
+who would like to pursue those set of questions. We did add some discussion
+about the power of the SSR test statistic against different kinds of causal
+effects: the SSR test statistic ought to be particularly sensitive to changes
+in mean differences (after all, recall that the coefficient on a dichotomous
+variable in an least squares model with a continuous outcome that minimizes
+the SSR is the difference of means) and ought to be less sensitive to other
+differences.
 
@@
 
-We have added some discussion about the idea that the SSR is particular sensitive to differences in means.
+In the text we note that the SSR is particularly sensitive to differences in
+means (or perhaps symmetric differences in tails that also produce a lot of
+movement in the mean of one group relative to another group) while the KS-test
+ought to have more power to detect asymmetric differences in the distributions
+of the treated versus control groups --- for example, differences that change
+variance without moving the means by stretching one tail of the distribution.
 

### 2015-11-30 simulation/ksvsssrpow.R
Do a small simulation to show SSR versus KS power

@@ -1,0 +1,36 @@
+## A file to do a quick demonstration of the differences in power between SSR and KS for simple constant effects models, simple experiment with no interference
+
+if(length(grep("TestStatRIInterference$",getwd()))==0){setwd("..")}
+
+library(RItools)
+source("simulation/setup.R")
+source("code/teststatistics.R")
+
+sampler <- simpleRandomSampler(total = n, treated = n/2)
+Zs <- sampler(REPETITIONS)$samples # N x nsims
+
+## Make two outcomes in control that both have 0 mean and 1 sd but with different distributions
+set.seed(20151130)
+tmpzif<-rgeom(n,prob=.7)
+y0zif<-scale(tmpzif)
+y0norm<-rnorm(n)
+
+summary(y0zif)
+summary(y0norm)
+
+## Truth is no effects
+y1norm<-y0norm
+y1zif<-y0zif
+
+##Alternatives
+ALTS<-seq(-10,10,length=100)
+simsamples<-10
+coamodel<-constant.additive.model
+
+rit<-RItest(y,z, ksTestStatistic, coamodel, ALTS,samples=simsamples)
+return(cbind(rit@params, p = rit[-1, "p.value"])) ## inefficient but avoids errors later
+
+##ksTestStatistic
+##ssrSimpleTestStatistic,
+
+

### 2015-11-30 simulation/ksvsssrpow.R
Working on the SSR versus KS simple power sim for different models of effects and different distributions of outcomes

@@ -1,4 +1,4 @@
-## A file to do a quick demonstration of the differences in power between SSR and KS for simple constant effects models, simple experiment with no interference
+## A file to do a quick demonstration of the differences in power between SSR and KS  simple experiment with no interference
 
 if(length(grep("TestStatRIInterference$",getwd()))==0){setwd("..")}
 
@@ -18,17 +18,65 @@ y0norm<-rnorm(n)
 summary(y0zif)
 summary(y0norm)
 
-## Truth is no effects
-y1norm<-y0norm
-y1zif<-y0zif
-
-##Alternatives
-ALTS<-seq(-10,10,length=100)
-simsamples<-10
+## Two models:
+### One model is like a model of mean shift
 coamodel<-constant.additive.model
+### One model focuses changes on the shape of the dist
+constant.multiplicative.model <- UniformityModel(function(y, z, tau) { y / ( 1 + z * ( tau-1 ) )  }, 
+                                           function(y_0, z, tau) { y_0 * ( 1 + z * ( tau-1 ) )})
 
-rit<-RItest(y,z, ksTestStatistic, coamodel, ALTS,samples=simsamples)
-return(cbind(rit@params, p = rit[-1, "p.value"])) ## inefficient but avoids errors later
+## quantchangemodel <- UniformityModel(function(y, z, tau) { y - z * tau }, 
+##                                            function(y_0, z, tau) { y_0 + z * tau})
+## 
+
+
+
+
+## Simulation settings
+ALTS<-list(tau=sort(unique(c(0,seq(-5,5,length=100)))))
+TRUTH<-list(tau=0)
+simsamples<-1000
+
+dotestMaker<-function(model,y0,truth,TZ,thegrid,simsamples){
+	force(model);force(y0);force(truth);force(TZ);force(thegrid);force(simsamples)
+	function(z){
+		y <- invertModel(model, y0, z, truth$tau)
+		rit<-RItest(y,z, TZ, model, thegrid,samples=simsamples)
+		return(cbind(rit@params, p = rit[-1, "p.value"])) ## inefficient but avoids errors later
+		##return(rit[-1, "p.value"]) ## efficient but depends on RItools not changing
+	}
+}
+
+
+
+testStats <- list( "SSR Test" = ssrSimpleTestStatistic, 
[…]

### 2015-11-30 coppock-replication/CoppockJEPS_figure2code.R
adding some more test stats, correcting bug in KS test stats

@@ -52,6 +52,7 @@ pmat.ks  <- matrix(NA, length(directs), length(indirects))
 
 pmat.ssr.comp <- matrix(NA, length(directs), length(indirects))
 pmat.ks.comp  <- matrix(NA, length(directs), length(indirects))
+pmat.ssr.comp.y0 <- matrix(NA, length(directs), length(indirects))
 
 set.seed(343)
 for(j in 1:length(directs)){
@@ -64,11 +65,13 @@ for(j in 1:length(directs)){
 
     Y0.comp <- Y.obs - Z.obs * direct.sim - exposure.computed * indirect.sim
     t.ks.comp  <- baseKSTest(Y0.comp, Z.obs)
+    t.ssr.comp.y0 <- sum(residuals(lm(Y0.comp ~ Z.obs + exposure.computed))^2)
 
     ssr.sims <- rep(NA, sims)
     ks.sims  <- rep(NA, sims)
     ssr.comp.sims <- rep(NA, sims)
     ks.comp.sims  <- rep(NA, sims)
+    ssr.comp.y0.sims <- rep(NA, sims)
 
     for(i in 1:sims){
       Z.sim <- Z_block[,sample(1:10000, 1)]
@@ -85,17 +88,21 @@ for(j in 1:length(directs)){
 
       ## KS tests
       # first, use Y0 based on "exposure.obs"
-      ks.sims[1] <- baseKSTest(pure.Y0, Z.sim)
+      ks.sims[i] <- baseKSTest(pure.Y0, Z.sim)
       # second, use Y0 based on the computed version of observed exposure
-      ks.comp.sims  <- baseKSTest(Y0.comp, Z.sim)
+      ks.comp.sims[i]  <- baseKSTest(Y0.comp, Z.sim)
+
+      # finally, compute the SSR using Y0.comp
+      ssr.comp.y0.sims[i] <- sum(residuals(lm(Y0.comp ~ Z.sim + s.exposure.sim))^2)
     }
     
     pmat.ssr[j,k] <- mean(ssr.obs > ssr.sims)
     pmat.ks[j, k] <- mean(ks.sims >= t.ks.obs)
 
     pmat.ssr.comp[j,k] <- mean(ssr.comp.sims <= t.ssr.comp)
-    pmat.ks.comp[j,k]  <= mean(ks.comp.sims >= t.ks.comp)
-    
[…]

### 2015-11-30 coppock-replication/CoppockJEPS_figure2graphic.R

@@ -4,17 +4,19 @@ load("fig2.rdata")
 n <- prod(dim(pmat.ssr))
 
 tmp <- expand.grid(x=directs, y=indirects)
-graph.frame <- rbind(tmp, tmp, tmp, tmp)
+graph.frame <- rbind(tmp, tmp, tmp, tmp, tmp)
 
 graph.frame$z <- c(as.vector(pmat.ssr),
                    as.vector(pmat.ks),
                    as.vector(pmat.ssr.comp),
-                   as.vector(pmat.ks.comp))
+                   as.vector(pmat.ks.comp),
+                   as.vector(pmat.ssr.comp.y0))
 
 graph.frame$t <- factor(c(rep("SSR (Coppock)", n),
                           rep("KS (Coppock)", n),
                           rep("SSR (Computed)", n),
-                          rep("KS (Computed)", n)))
+                          rep("KS (Computed)", n),
+                          rep("SSR (Y0, Comp)", n)))
 
 col.l <- colorRampPalette(c('white', 'black'))(1000)
 
@@ -40,10 +42,7 @@ fig2 <- levelplot(z ~ x*y | t, graph.frame, cuts=20, col.regions=col.l,
                                             draw = FALSE))),
 )
   
-pdf("CoppockJEPS_figure2.pdf", width = 8, height = 4)
+pdf("CoppockJEPS_figure2.pdf", width = 8, height = 8)
 print(fig2)
 dev.off()
 
-png("CoppockJEPS_figure2.png", width = 8, height = 4)
-print(fig2)
-dev.off()

### 2015-12-03 code/fastTestStatistics.cpp
Working on cpp versions of test statistics

@@ -1,0 +1,38 @@
+#include <RcppArmadillo.h>
+
+// [[Rcpp::depends(RcppArmadillo)]]
+
+using namespace Rcpp;
+using namespace arma;
+
+// [[Rcpp::export]]
+double ksFastTestStatistic(arma::colvec y, arma::colvec z){
+	uvec trt = find( z == 1 );
+	uvec ctrl = find( z == 0 );
+	vec yT = y.elem(trt);
+	vec	yC = y.elem(ctrl);
+	// assuming no missing vals	yT = yT[!is.na(yT)];
+	double nX = yT.n_rows; //length(yT);
+	double nY = yC.n_rows;
+	double n = nX * nY/(nX + nY);
+	colvec w = join_cols(yT, yC); // c(yT,yC)
+	uvec orderedw = sort_index(w); //order(w)
+	int totalN = nX + nY;
+	colvec d(totalN);
+	d.fill(1/nX);
+	d.elem( find( orderedw > nX ) ).fill(-1/nY);
+	// d = cumsum(Rcpp::ifelse(orderedw <= nX, 1/nX, -1/nY));
+	colvec F = cumsum(d);
+	// check for ties
+	vec uniqF = unique(w);
+	vec out = F;
+	if ( uniqF.n_elem < totalN ){
+		//uvec nonzerodiffs = find( diff(sort(F)) != 0 );
+		vec Fnonzerodiffs = F.elem( find( diff(sort(F)) != 0 ) );
+		vec out = Fnonzerodiffs; // I dont know why I cant use join_cols (Fnonzerodiffs,F(totalN-1))
+		out.insert_rows(1,F(totalN-1)); // z <- z[c(which(diff(sort(w)) != 0), n.x + n.y)]
+		}
+	// return(max(abs(out)));
+	return(max(abs(out)));
+}
+

### 2015-12-03 code/faststuff.cpp

@@ -1,0 +1,70 @@
+#include <RcppArmadillo.h>
+
+// [[Rcpp::depends(RcppArmadillo)]]
+
+using namespace Rcpp;
+using namespace arma;
+
+// from https://stackoverflow.com/questions/29851637/efficiently-perform-row-wise-distribution-test
+double KS(arma::colvec x, arma::colvec y) {
+	int n = x.n_rows;
+	arma::colvec w = join_cols(x, y);
+	arma::uvec z = arma::sort_index(w);
+	w.fill(-1); w.elem( find(z <= n-1) ).ones();
+	return max(abs(cumsum(w)))/n;
+}
+// [[Rcpp::export]]
+Rcpp::NumericVector K_S(arma::mat mt) {
+	int n = mt.n_cols;
+	Rcpp::NumericVector results(n);
+	for (int i=1; i<n;i++) {
+		arma::colvec x=mt.col(i-1);
+		arma::colvec y=mt.col(i);
+		results[i] = KS(x, y);
+	}
+	return results;
+}
+
+// [[Rcpp::export]]
+arma::rowvec fastcolMeans(const arma::mat & X){
+	arma::rowvec out = arma::mean(X,0); //Armadillo mean(X,dim)
+	return(out);
+}
+
+
+// [[Rcpp::export]]
+mat fastcolSums(const mat & X){
+	arma::mat out = arma::sum(X,0); //Armadillo mean(X,dim)
+	return(out);
+}
+
+// [[Rcpp::export]]
+mat fastcolVars(const mat & X){
+	arma::mat out = arma::var(X,0); //Armadillo var(X,dim),0=col, 1=row
+	return(out);
+}
+
+// [[Rcpp::export]]
+vec fastrowMeans(const arma::mat & X){
+	arma::vec out = arma::mean(X,1); //Armadillo mean(X,dim)
+	return(out);
+}
+
+
+// [[Rcpp::export]]
+vec fastrowSums(const mat & X){
+	arma::vec out = arma::sum(X,1); //Armadillo mean(X,dim)
+	return(out);
+}
+
+// [[Rcpp::export]]
+vec fastrowSds(const arma::mat & X){
+	arma::vec out = arma::stddev(X,0,1); //Armadillo stddev(X,norm_type=0,dim)
+	return(out);
+}
+
+// [[Rcpp::export]]
+mat fastcolSds(const arma::mat & X){
[…]

### 2015-12-03 simulation/ksvsssrpow.R
Trying to run the KS versus SSR power comparison for two simple models and no interference and two outcomes to respond to the question about KS versus SSR intuition

@@ -1,6 +1,7 @@
 ## A file to do a quick demonstration of the differences in power between SSR and KS  simple experiment with no interference
 
-if(length(grep("TestStatRIInterference$",getwd()))==0){setwd("..")}
+if(length(grep("TestStatRIInterference$",getwd(),ignore.case=TRUE))==0){setwd("..")}
+if(length(grep("TestStatRIInterference",.libPaths(),ignore.case=TRUE))==0){.libPaths(".libraries") }
 
 library(RItools)
 source("simulation/setup.R")
@@ -12,7 +13,7 @@ Zs <- sampler(REPETITIONS)$samples # N x nsims
 ## Make two outcomes in control that both have 0 mean and 1 sd but with different distributions
 set.seed(20151130)
 tmpzif<-rgeom(n,prob=.7)
-y0zif<-scale(tmpzif)
+y0zif<-as.vector(scale(tmpzif))
 y0norm<-rnorm(n)
 
 summary(y0zif)
@@ -22,20 +23,30 @@ summary(y0norm)
 ### One model is like a model of mean shift
 coamodel<-constant.additive.model
 ### One model focuses changes on the shape of the dist
-constant.multiplicative.model <- UniformityModel(function(y, z, tau) { y / ( 1 + z * ( tau-1 ) )  }, 
-                                           function(y_0, z, tau) { y_0 * ( 1 + z * ( tau-1 ) )})
-
-## quantchangemodel <- UniformityModel(function(y, z, tau) { y - z * tau }, 
-##                                            function(y_0, z, tau) { y_0 + z * tau})
-## 
-
-
-
+constant.multiplicative.model <- UniformityModel( function(y, z, tau) {
+													 if(tau==0){
+														 return(rep(0,length(y)))
+													 } else {
+														 y / ( 1 + z * ( tau-1 ) )
+													 }
+}, function(y_0, z, tau) { y_0 * ( 1 + z * ( tau-1 ) )})
 
 ## Simulation settings
 ALTS<-list(tau=sort(unique(c(0,seq(-5,5,length=100)))))
 TRUTH<-list(tau=0)
 simsamples<-1000
+
+## Trying to speed things up but not passing tests
+## library(Rcpp)
+## library(RcppArmadillo)
[…]

### 2015-12-04 simulation/ksvsssrpow.R
Finished setting up the KS versus SSR power without interference simulation comparison

@@ -115,5 +115,5 @@ for(i in 1:length(testStats)){
 	}
 }
 
-save(ls(patt="results"),file="simulation/ksvsssrpowresults.rda")
+save(list=ls(patt="resultsCO"),file="simulation/ksvsssrpowresults.rda")
 stopCluster(cl)

### 2015-12-04 simulation/ksvsssrpow-results.R
Adding a file to extract the power from the simulations of KS versus SSR

@@ -1,0 +1,29 @@
+## A file to extract power and error rate from the huge list of RItest objects
+
+library(RItools,lib.loc=".libraries")
+
+source("simulation/setup.R")
+
+load("simulation/ksvsssrpowresults.rda")
+ALTS<-list(tau=sort(unique(c(0,seq(-5,5,length=100)))))
+
+simPow2<-function(lst){
+	m <- sapply(lst,function(l){ l$p})
+	m <- m <= ALPHA
+	rowMeans(m)
+}
+
+powComparison<-sapply(ls(patt="resultsCO"),function(nm){
+simPow2(get(nm))
+})
+stopifnot(length(unique(signif(ALTS$tau,4)))==length(ALTS$tau))
+row.names(powComparison)<-signif(ALTS$tau,4)
+
+apply(powComparison,2,summary)
+
+powComparison[powComparison[,"resultsCOANormKS"]<1,]
+
+save(powComparison,file="simulation/ksvsssrpow-results.rda")
+
+
+

### 2015-12-04 figures/ksvsssrpow-summary.R
Started a file to summarize the results of the KS versus SSR no interference comparison

@@ -1,0 +1,5 @@
+# A file to summarize the results of the KS versus SSR power comparison for no interference but difference models of effects and outcomes
+
+load("simulations/ksvsssrpow-results.rda")
+
+

### 2015-12-04 figures/ksvsssrpow-summary.R
Small summary of the KS versus SSR no interference comparison

@@ -1,5 +1,13 @@
 # A file to summarize the results of the KS versus SSR power comparison for no interference but difference models of effects and outcomes
 
-load("simulations/ksvsssrpow-results.rda")
+if(length(grep("TestStatRIInterference$",getwd(),ignore.case=TRUE))==0){setwd("..")}
+if(length(grep("TestStatRIInterference",.libPaths(),ignore.case=TRUE))==0){.libPaths(".libraries") }
+source("simulation/setup.R")
+load("simulation/ksvsssrpow-results.rda")
+ALTS<-list(tau=sort(unique(c(0,seq(-5,5,length=100)))))
+
+apply(powComparison,2,summary)
+
+t(powComparison[powComparison[,"resultsCOANormKS"]<1,])
 
 

### 2015-12-14 simulation/ksvsssrpow.R
Fix up basic simulation of KS versus SSR

@@ -13,15 +13,16 @@ Zs <- sampler(REPETITIONS)$samples # N x nsims
 ## Make two outcomes in control that both have 0 mean and 1 sd but with different distributions
 set.seed(20151130)
 tmpzif<-rgeom(n,prob=.7)
-y0zif<-as.vector(scale(tmpzif))
-y0norm<-rnorm(n)
+y0zif<-as.vector(scale(tmpzif))+10
+y0norm<-rnorm(n,mean=10)
 
 summary(y0zif)
 summary(y0norm)
 
 ## Two models:
 ### One model is like a model of mean shift
-coamodel<-constant.additive.model
+constant.additive.model
+
 ### One model focuses changes on the shape of the dist
 constant.multiplicative.model <- UniformityModel( function(y, z, tau) {
 													 if(tau==0){
@@ -32,8 +33,41 @@ constant.multiplicative.model <- UniformityModel( function(y, z, tau) {
 }, function(y_0, z, tau) { y_0 * ( 1 + z * ( tau-1 ) )})
 
 ## Simulation settings
-ALTS<-list(tau=sort(unique(c(0,seq(-5,5,length=100)))))
-TRUTH<-list(tau=0)
+### Different parameter settings can mean the same thing in the two different models
+TRUTHAdditive<-list(tau=0)
+TRUTHMultiplicative<-list(tau=1)
+all.equal(constant.additive.model(y0norm,Zs[,1],0),constant.multiplicative.model(y0norm,Zs[,1],1))
+all.equal(constant.additive.model(y0zif,Zs[,1],0),constant.multiplicative.model(y0zif,Zs[,1],1))
+
+## Try to make the hypotheses in the multiplicative model similar in magnitude
+## to the hypotheses of the multiplicative model (we cannot make them all
+## exactly the same across all values of y0, but we can make them as close as
+## possible)
+
+ALTSAdditive<-list(tau=sort(unique(c(0,1,seq(-5,5,length=100)))))
+
+addvsmult<-function(x,tau){
+	addres<-invertModel(constant.additive.model,y0norm,Zs[,1],tau=tau)
+	mulres<-invertModel(constant.multiplicative.model,y0norm,Zs[,1],tau=x*tau)
+	median(abs(mulres-addres))
+}
+
[…]

### 2015-12-15 figures/ksvsssrpow-summary.R
Fixing up sims --- making sure that multiplicative and additive models both test the true null of no effects

@@ -4,10 +4,15 @@ if(length(grep("TestStatRIInterference$",getwd(),ignore.case=TRUE))==0){setwd(".
 if(length(grep("TestStatRIInterference",.libPaths(),ignore.case=TRUE))==0){.libPaths(".libraries") }
 source("simulation/setup.R")
 load("simulation/ksvsssrpow-results.rda")
-ALTS<-list(tau=sort(unique(c(0,seq(-5,5,length=100)))))
+load("simulation/simplealts.rda")
+load("simulation/simpletruth.rda")
 
-apply(powComparison,2,summary)
+apply(powCOA,2,summary)
+apply(powCOM,2,summary)
 
-t(powComparison[powComparison[,"resultsCOANormKS"]<1,])
+t(powCOA[powCOA[,"resultsCOANormKS"]<1,])
+t(powCOM[powCOM[,"resultsCOMNormKS"]<1,])
 
-
+## SSR is slightly more power than KS for COA with Normal outcomes
+## KS is more powerful than SSR for COA with Zif outcomes
+## SSR is more powerful that KS for COM with Normal outcomes and less power than KS for Zif outcomes

### 2015-12-15 simulation/ksvsssrpow-results.R

@@ -1,11 +1,16 @@
 ## A file to extract power and error rate from the huge list of RItest objects
 
-library(RItools,lib.loc=".libraries")
+if(length(grep("TestStatRIInterference$",getwd(),ignore.case=TRUE))==0){setwd("..")}
+if(length(grep("TestStatRIInterference",.libPaths(),ignore.case=TRUE))==0){.libPaths(".libraries") }
+
+library(RItools) ##,lib.loc=".libraries")
 
 source("simulation/setup.R")
 
 load("simulation/ksvsssrpowresults.rda")
-ALTS<-list(tau=sort(unique(c(0,seq(-5,5,length=100)))))
+load("simulation/simplealts.rda")
+load("simulation/simpletruth.rda")
+
 
 simPow2<-function(lst){
 	m <- sapply(lst,function(l){ l$p})
@@ -16,14 +21,24 @@ simPow2<-function(lst){
 powComparison<-sapply(ls(patt="resultsCO"),function(nm){
 simPow2(get(nm))
 })
-stopifnot(length(unique(signif(ALTS$tau,4)))==length(ALTS$tau))
-row.names(powComparison)<-signif(ALTS$tau,4)
 
-apply(powComparison,2,summary)
+powCOA<-powComparison[,grep("COA",colnames(powComparison))]
+stopifnot(length(unique(signif(ALTS$COA$tau,4)))==length(ALTS$COA$tau))
+row.names(powCOA)<-signif(ALTS$COA$tau,4)
 
-powComparison[powComparison[,"resultsCOANormKS"]<1,]
 
-save(powComparison,file="simulation/ksvsssrpow-results.rda")
+powCOM<-powComparison[,grep("COM",colnames(powComparison))]
+stopifnot(length(unique(signif(ALTS$COM$tau,5)))==length(ALTS$COM$tau))
+row.names(powCOM)<-signif(ALTS$COM$tau,5)
 
+
+
+apply(powCOA,2,summary)
+apply(powCOM,2,summary)
+
+powCOA[powCOA[,"resultsCOANormKS"]<1,]
+powCOM[powCOM[,"resultsCOMNormKS"]<1,]
+
+save(powCOA,powCOM,file="simulation/ksvsssrpow-results.rda")
 
 

### 2015-12-15 simulation/ksvsssrpow.R

@@ -25,50 +25,55 @@ constant.additive.model
 
 ### One model focuses changes on the shape of the dist
 constant.multiplicative.model <- UniformityModel( function(y, z, tau) {
-													 if(tau==0){
-														 return(rep(0,length(y)))
-													 } else {
-														 y / ( 1 + z * ( tau-1 ) )
-													 }
+						   if(tau==0){
+						     return(rep(0,length(y)))
+						   } else {
+						     y / ( 1 + z * ( tau-1 ) )
+						   }
 }, function(y_0, z, tau) { y_0 * ( 1 + z * ( tau-1 ) )})
 
 ## Simulation settings
 ### Different parameter settings can mean the same thing in the two different models
 TRUTHAdditive<-list(tau=0)
 TRUTHMultiplicative<-list(tau=1)
-all.equal(constant.additive.model(y0norm,Zs[,1],0),constant.multiplicative.model(y0norm,Zs[,1],1))
-all.equal(constant.additive.model(y0zif,Zs[,1],0),constant.multiplicative.model(y0zif,Zs[,1],1))
+stopifnot(all.equal(constant.additive.model(y0norm,Zs[,1],0),constant.multiplicative.model(y0norm,Zs[,1],1)))
+stopifnot(all.equal(constant.additive.model(y0zif,Zs[,1],0),constant.multiplicative.model(y0zif,Zs[,1],1)))
 
 ## Try to make the hypotheses in the multiplicative model similar in magnitude
 ## to the hypotheses of the multiplicative model (we cannot make them all
 ## exactly the same across all values of y0, but we can make them as close as
 ## possible)
 
-ALTSAdditive<-list(tau=sort(unique(c(0,1,seq(-5,5,length=100)))))
+ALTSAdditive<-list(tau=sort(unique(c(0,1,seq(-2,2,length=100)))))
 
 addvsmult<-function(x,tau){
-	addres<-invertModel(constant.additive.model,y0norm,Zs[,1],tau=tau)
-	mulres<-invertModel(constant.multiplicative.model,y0norm,Zs[,1],tau=x*tau)
-	median(abs(mulres-addres))
+  addres<-invertModel(constant.additive.model,y0norm,Zs[,1],tau=tau)
+  mulres<-invertModel(constant.multiplicative.model,y0norm,Zs[,1],tau=x*tau)
[…]

### 2015-12-15 paper/introduction.tex
CHanged section title
Edited the first section of the paper.

@@
-\section{Background on randomization based statistical inference for causal effects from social network experiments}
+\section{Overview: Randomization based statistical inference for causal effects from social network experiments}
 
 In a randomized experiment with $n=4$ subjects connected via a fixed network,
-subject $i=1$ might respond differently to different ways that treatment is
-assigned to the whole network. When the treatment assignment vector provides
-treatment to persons 2 and 3, $\bz=\{0,1,1,0\}$ she might act one way,
+the response of subject $i=1$ might depend on the different ways that
+treatment is assigned to the \emph{whole} network. When the treatment
+assignment vector,$\zb$, provides treatment to persons 2 and 3,
+$\bz=\{0,1,1,0\}$, person $i=1$ might respond one way,
 $y_{i=1,\bz=\{0,1,1,0\}}$ and when treatment is assigned to persons 3 and 4,
-$\bz=\{0,0,1,1\}$ she might act another way, $y_{i=1,\bz=\{0,0,1,1\}}$ That
-is, if the experiment had a causal effect on person $i$, then her outcome
-would differ under different realizations of the experimental treatment
-$y_{i,\bz} \ne y_{i,\bz'}$. The fundamental problem of causal inference
-reminds us that we can never see both states of the world: we only observe the
-outcome from person $i$ under one treatment assignment vector, either $\bz$ or
-some $\bz'$ not both  \citep{holland:1986a,brady2008cae}.
+$\bz=\{0,0,1,1\}$ person $i=1$ might act another way,
+$y_{i=1,\bz=\{0,0,1,1\}}$. More generally, we might say that if the experiment had a causal effect on
+person $i$, then her outcome would differ under different realizations of the
+experimental treatment as a whole $y_{i,\bz} \ne y_{i,\bz'}$. The fundamental problem of
+causal inference reminds us that we can never see both states of the world: we
+only observe the outcome from person $i$ under one treatment assignment
+vector, either $\bz$ or some $\bz'$ not both
+\citep{holland:1986a,brady2008cae}.\footnote{In simpler settings, where
+  treatment given to one individual has no effect on any other individual, we tend to
+  write $y_{i,Z_i=1} \ne y_{i,Z_i=0}$ to say that treatment had a causal
+  effect on person $i$.}
 
@@
 2]{fisher:1935} as developed by Paul Rosenbaum  \citep{rosenbaum2010design}
-begins with the premise of the fundamental problem of causal inference. Given
-the unobservability of potential outcomes, the Fisher and Rosenbaum approach
+begins with the premise of the fundamental problem of causal inference. Since
+we cannot observe all of the ways that a given person would respond to
+different treatments, the Fisher and Rosenbaum approach
 suggests that we focus on learning about how \emph{models} of unobservable
@@
 treatments, the only observations available to evaluate the model were
-comparisons of the assigned to treatment group and the assigned to control
+comparisons of the assigned-to-treatment group and the assigned-to-control
 group. Since their model could imply not only shifts in the mean of the
@@
 so that their tests would be sensitive to differences in the treatment and
-control distributions implied by different hypotheses and not merely sensitive
+control distributions implied by different hypotheses and not only sensitive
 to differences in one aspect of those distributions (such as the differences
@@
   records the proportion of the distribution of $x$ at or below $x_i$
-  \citep[\S 5.4]{MylesHollander1999a}. \label{fn:kstest}} So, in broad
+  \citep[\S 5.4]{MylesHollander1999a}.\label{fn:kstest}} So, in broad
 outline, the BFP approach involves (1) the articulation of a model for how a
 treatment assignment vector can change outcomes for all subjects in the
-experiment (holding the network fixed) and (2) use a function to comparing
-actually treated and control observations summarize whether such a model is
-implausible (codified as a low $p$-value)  or whether we have too little
+experiment (holding the network fixed) and (2) the use of a function to
+compare actually treated and control observations to summarize whether such a model is
+implausible (codified as a low $p$-value)  or to report that we have too little
 information available from the data and design about the model (codified as a
 high $p$-value). This is classic hypothesis testing applied to an experiment
-on a social network.
+on a social network with sharp null hypotheses.
 
-So, say $Y_i$ is the observed outcome and we hypothesize that units do not
-interfere and also that $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$. We can
-assess which (if any) hypothesized values of $\tau$ appear implausible from
-the perspective of the data by: (1) Mapping the hypothesis about unobserved
-quantities to observed data using the identity $Y_i=Z_i y_{i,Z_i=1} + (1-Z_i) y_{i,Z_i=0}$ ---
-noticing that if $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ then $y_{i,Z_i=0}=Y_i - Z_i
-\tau$ (by substituting from the hypothesized relationship into the observed
-data identity); (2) Using this result to adjust the observed outcome to
-represent what would be implied by the hypothesis for given $\tau_0$ such that
-$\widetilde y_{i,Z_i=0}=Y_i - Z_i \tau_0$; and (3) Under the hypothesis,
+Say $Y_i$ is the observed outcome and we hypothesize that units do not
+interfere and also that $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$. We can assess which
+(if any) hypothesized values of $\tau$ appear implausible from the perspective
+of the data by: (1) Mapping the hypothesis about unobserved quantities to
+observed data using the identity $Y_i=Z_i y_{i,Z_i=1} + (1-Z_i) y_{i,Z_i=0}$
+--- noticing that if $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ then $y_{i,Z_i=0}=Y_i -
+Z_i \tau$ (by substituting from the hypothesized relationship into the
+observed data identity); (2) Using this result to adjust the observed outcome
+to represent what would be implied by the hypothesis for a given $\tau_0$ such
+that $\widetilde y_{i,Z_i=0}=Y_i - Z_i \tau_0$; and (3) Under the hypothesis,
 $\widetilde y_{i,Z_i=0}$ should have no systematic relationship with treatment
 assignment, so we can summarize this relationship with a test statistic,
-$\mathcal T(\widetilde y_{i,Z_i=0},Z_i)$ and the distribution of values for
-this test statistic arising from repetitions of treatment assignment (new
-draws of $\bz$ from all of the ways that such treatment assignment vectors
-could have been produced); and finally (4) a $p$-value arises by comparing the
-observed test statistic, $\mathcal T(Y_i,Z_i)$ against the distribution of that test
-statistic that characterizes the hypothesis.
+$\mathcal T(\widetilde y_{i,Z_i=0},Z_i)$. A distribution of values for this
+test statistic arises from repetitions of the treatment assignment process
+(new draws of $\bz$ from all of the ways that such treatment assignment
+vectors could have been produced); and finally (4) a $p$-value arises by
+comparing the observed test statistic, $\mathcal T(Y_i,Z_i)$ against the
+distribution of that test statistic that characterizes the hypothesis.
 
@@
 assessment; (2) or might reject all such hypothesized parameters. Thus,
-although in theory one may assess sharp hypotheses about multiparameter
-settings, in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
+although in theory one may assess sharp multiparameter hypotheses,
+ in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
 characteristics of tests as a piece of their workflow --- because  the theory

### 2015-12-15 figures/ksvsssrpow-summary.R
Figures and summaries of the simple KS versus SSR simulation

@@ -16,3 +16,14 @@ t(powCOM[powCOM[,"resultsCOMNormKS"]<1,])
 ## SSR is slightly more power than KS for COA with Normal outcomes
 ## KS is more powerful than SSR for COA with Zif outcomes
 ## SSR is more powerful that KS for COM with Normal outcomes and less power than KS for Zif outcomes
+
+pdf(file="KSVSSSRSimplePowplot.pdf",width=8,height=4)
+par(mfrow=c(1,2),oma=rep(0,4),mgp=c(1.5,.5,0),pty="s",mar=c(3,3,2,0))
+matplot(x=ALTS$COA$tau,powCOA,type="l",lty=c(1,2,1,2),col=c(1,1,2,2),
+	main="Constant Additive Model",xlab="Hypotheses (Truth=0)",ylab="pr(p<=.05)")
+legend(x="bottomleft",legend=substr(colnames(powCOA),11,50),lty=c(1,2,1,2),col=c(1,1,2,2),bty="n")
+matplot(x=ALTS$COM$tau,powCOM,type="l",lty=c(1,2,1,2),col=c(1,1,2,2),
+	main="Constant Multiplicative Model", xlab="Hypotheses (Truth=1)",ylab="pr(p<=.05)")
+legend(x="bottomleft",legend=substr(colnames(powCOM),11,50),lty=c(1,2,1,2),col=c(1,1,2,2),bty="n")
+dev.off()
+

### 2015-12-15 coppock-replication/CoppockJEPS_figure2code.R
formatting KS vs. SSR plot using Coppock built data

@@ -23,12 +23,6 @@ source("../code/teststatistics.R")
 ## Pick exposure model -- main analysis uses dw nominate
 Z.obs <- CoppockJEPS$treatment
 Y.obs <- CoppockJEPS$sb24
-exposure.obs <- CoppockJEPS$s.difs.dw
-
-fit.obs <- lm(Y.obs ~ Z.obs + exposure.obs)
-ssr.obs <- sum(residuals(fit.obs)^2)
-direct.obs <- fit.obs$coefficients[2]
-indirect.obs <- fit.obs$coefficients[3]
 
 exposure.expected.1 <- CoppockJEPS$exposure.expected.1.dw
 exposure.expected.0 <- CoppockJEPS$exposure.expected.0.dw
@@ -39,20 +33,20 @@ exposure <- function(Z) {
   s.exposure.sim <- (exposure.sim.corrected - mean(exposure.sim.corrected))/sd(exposure.sim.corrected)
 }
 
-exposure.computed <- exposure(Z.obs)
+ssr <- function(y, z) {
+  e <- exposure(z)
+  sum(residuals(lm(y ~ z + e))^2)
+}
 
-t.ssr.comp <- sum(residuals(lm(Y.obs ~ Z.obs + exposure.computed))^2)
+exposure.obs <- exposure(Z.obs)
 
-directs <-seq(from=-.7, to=0.2, by=.025)
-indirects <-seq(from=-.7, to=0.2, by=.025)
+param.steps <- 50
+directs   <-seq(from = -1,  to = 1, length.out = param.steps)
+indirects <-seq(from = -1,  to = 1, length.out = param.steps)
 
 sims <- 50
 pmat.ssr <- matrix(NA, length(directs), length(indirects))
 pmat.ks  <- matrix(NA, length(directs), length(indirects))
-
-pmat.ssr.comp <- matrix(NA, length(directs), length(indirects))
-pmat.ks.comp  <- matrix(NA, length(directs), length(indirects))
-pmat.ssr.comp.y0 <- matrix(NA, length(directs), length(indirects))
 
 set.seed(343)
 for(j in 1:length(directs)){
@@ -61,48 +55,22 @@ for(j in 1:length(directs)){
     indirect.sim <- indirects[k]
 
[…]

### 2015-12-15 coppock-replication/CoppockJEPS_figure2graphic.R

@@ -3,46 +3,27 @@ load("fig2.rdata")
 
 n <- prod(dim(pmat.ssr))
 
-tmp <- expand.grid(x=directs, y=indirects)
-graph.frame <- rbind(tmp, tmp, tmp, tmp, tmp)
+tmp <- expand.grid(x = directs, y = indirects)
+graph.frame <- rbind(tmp, tmp)
 
 graph.frame$z <- c(as.vector(pmat.ssr),
-                   as.vector(pmat.ks),
-                   as.vector(pmat.ssr.comp),
-                   as.vector(pmat.ks.comp),
-                   as.vector(pmat.ssr.comp.y0))
+                   as.vector(pmat.ks))
 
-graph.frame$t <- factor(c(rep("SSR (Coppock)", n),
-                          rep("KS (Coppock)", n),
-                          rep("SSR (Computed)", n),
-                          rep("KS (Computed)", n),
-                          rep("SSR (Y0, Comp)", n)))
+graph.frame$t <- factor(c(rep("SSR", n),
+                          rep("KS", n)))
 
 col.l <- colorRampPalette(c('white', 'black'))(1000)
 
 depth.breaks <- do.breaks(c(0,1), 20)
-fig2 <- levelplot(z ~ x*y | t, graph.frame, cuts=20, col.regions=col.l,
+fig2 <- levelplot(z ~ x*y | t,
+                  graph.frame,
+                  cuts = 20,
+                  col.regions = col.l,
                   colorkey=FALSE,
-                  at=depth.breaks,
-                        ylab = "Hypothesized indirect effect",
-                        xlab = "Hypothesized direct effect",
-                        scales=list(x=list(at=round(seq(-.7, .2, by=.1), digits=1), labels=round(seq(-.7, .2, by=.1), digits=1)),
-                                    y=list(at=round(seq(-.7, .2, by=.1), digits=1), labels=round(seq(-.7, .2, by=.1), digits=1))),
-                        panel = function(...) {
-                          panel.levelplot(...)
-                          panel.abline(h = 0, lty=2)
-                          panel.abline(v = 0, lty=2)
-                          larrows(y0=-.5, y1= -.5, x0=directs[direct_breaks[1]], x1=directs[direct_breaks[2]], angle=90,code=3)
-                          larrows(x0=-.6, x1= -.6, y0=indirects[indirect_breaks[1]], y1=indirects[indirect_breaks[2]], angle=90,code=3)
-                        },
-                  legend = 
[…]

### 2015-12-16 coppock-replication/CoppockJEPS_figure2code.R
bumping up the number of iterations in p-value computation

@@ -44,7 +44,7 @@ param.steps <- 50
 directs   <-seq(from = -1,  to = 1, length.out = param.steps)
 indirects <-seq(from = -1,  to = 1, length.out = param.steps)
 
-sims <- 50
+sims <- 500
 pmat.ssr <- matrix(NA, length(directs), length(indirects))
 pmat.ks  <- matrix(NA, length(directs), length(indirects))
 

### 2015-12-16 paper/introduction.tex
Integrating Coppock relication into paper
Wrote up a few paragraphs about the Coppock replication and included the
figure.

@@
 treatment is assigned to the \emph{whole} network. When the treatment
-assignment vector,$\zb$, provides treatment to persons 2 and 3,
+assignment vector,$\bz$, provides treatment to persons 2 and 3,
 $\bz=\{0,1,1,0\}$, person $i=1$ might respond one way,
@@
 
-\section{Application: The Impact of Constituency Opinion on Legislative
-  Preferences}
+\section{Application: Legislative Information Spillovers}
 
-REWRITE. FIX.
+\citet{coppock2014information} presents an reanalysis of an experiment performed
+by \citet{butlernickerson2011}. Leading up to a key vote on a budget bill, SB24, in the
+New Mexico state senate, legislators were randomly assigned to be given
+constituent survey results, and their votes on SB24 recorded. The original
+analysis found that legislators assigned to receive constituent information were
+more likely to vote consistently with the preferences of their district.
+Given the easy nature of sharing information in
+field experiments \citep{Winters:2012fk}, Coppock reanalyzes the
+experimental results with an eye towards spillover, constructing a social
+network based on the similarity of W-NOMINATE ideology scores. Coppock hypothesized
+that ideologically similar legislators would be more likely to share
+information. If the constituent survey information does induce a change in behavior, this
+effect may be observed not only in those legislators that are directly treated
+in the experiment, but also by those of similar ideology to those that are
+treated.
 
-\citep{coppock2014information} used a version of the SSR+Design test statistic
-to show the effect of constituency opinion on legislators preferences within
-the BFP framework.  We here show how his results would have been different had
-he used the simple KS test.\footnote{In this case we also found that only
-  the SSR+Design test statistic had power in all directions and that the KS
-  statistics (+Design, +Degree, and plain) performed basically as they do in
-  figure~\ref{fig:twoD}.}
+With distance matrix $\Gamma$, Coppock evaluated the model:
+
+$$\yu = \yz - \beta_1 \bz - \beta_2 g(\Gamma \bz)$$
+
+Where $g$ is a function that normalizes the sums of distances to treated
+neighbors to have unit variance. This model has a direct
+effect ($\beta_1$) and an indirect effect ($\beta_2$) that is linear in the distances to
+treated neighbors. Coppock evaluated this spillover model using the SSR+Design statistic as
+presented in this paper. To demonstrate the value of this statistic,
+particularly for models that are linear in the spillover effects, we repeat
+Coppock's analysis using both the SSR+Design statistic and a KS statistic. The
+results in Figure~\ref{fig:coppock-replication} recreate Coppock's Figure
+2.\footnote{During the replication process, we noticed an inconsistency between
+  our results using the SSR+Design statistic and Coppock's published figure. In
+  coordination with the author, we discovered that the published figure suffered
+  from on a programming error. The figure in this document reflects a correction
+  to that error, so it will not precisely match Coppock's original publication.}
+The KS test statistic, while being more discerning in the neighborhood of (0,
+0), fails to bound the region of plausible hypotheses in the manner of the
+SSR+Design statistic. We find this quality of the test statistic to be
+particularly valuable. A powerful test statistic is a useful
+tool, but equally useful is the ability to succinctly describe the region in
+which we would fail to reject a null hypothesis. In this case, the SSR statistic
+provides such a region, making discussing and evaluating the result's
+of Coppock's analysis much more straightforward: it would be very unlikely that
+the data were generated from the model with parameters outside the rectangle
+with corners (-0.6, 0.2) and (0.3, -0.4).
+
+\begin{figure}[h!] \centering
+  \includegraphics[width=.99\textwidth]{../coppock-replication/CoppockJEPS_figure2.pdf}
+  \caption{Replication of Figure 2 from \citet{coppock2014information}: plotting
+    $p$-values for direct and indirect model parameters. The
+    original analysis used the SSR statistic, which nicely bounds the region of
+    plausible hypotheses. The KS statistic, while eliminating more hypotheses in
+  the neighborhood of (0, 0), fails to provide any such bounds.}
+  \label{fig:coppock-replication}
+\end{figure}
 

### 2015-12-16 coppock-replication/CoppockJEPS_figure2graphic.R

@@ -10,7 +10,8 @@ graph.frame$z <- c(as.vector(pmat.ssr),
                    as.vector(pmat.ks))
 
 graph.frame$t <- factor(c(rep("SSR", n),
-                          rep("KS", n)))
+                          rep("KS", n)),
+                        levels = c("SSR", "KS"))
 
 col.l <- colorRampPalette(c('white', 'black'))(1000)
 
@@ -19,7 +20,7 @@ fig2 <- levelplot(z ~ x*y | t,
                   graph.frame,
                   cuts = 20,
                   col.regions = col.l,
-                  colorkey=FALSE,
+                  colorkey = TRUE,
                   xlab = "Direct Effect",
                   ylab = "Indirect Effect")
   

### 2015-12-18 paper/introduction.tex
More work editing the introduction in response to reviews

@@
 unbiased test of the hypothesis of no effects or positive effects when the
-positive effects involve one parameter.  Yet, when the models are complex and
+positive effects involve one parameter. Notice that this means that, a test of
+the sharp null of no effects
+based on a known randomization using \emph{any} effect increasing
+test-statistic to be valid test (in that the test should produce $p$-values
+less than $\alpha$ no more than 100$\alpha$ of the time when the null is true) even though different
+test statistics may imply different power against false hypotheses.  Yet, when the models are complex and
 may involve increasing effects in the direction of one parameter and
@@
   \includegraphics[width=.99\textwidth]{twoDplots.pdf} \caption{Proportion of
-    $p$-values less than .05 for tests of joint hypotheses about $\tau$ and
+    $p$-values less than $.05$ for tests of joint hypotheses about $\tau$ and
     $\beta$ for the model in equation~\ref{eq:spillovermodelA}. Darker values
@@
 by \citet{butlernickerson2011}. Leading up to a key vote on a budget bill, SB24, in the
-New Mexico state senate, legislators were randomly assigned to be given
+New Mexico state senate, legislators were randomly assigned to see
 constituent survey results, and their votes on SB24 recorded. The original
@@
 probability weights so as to ensure representative samples of potential
-outcomes. This suggests a conjecture: that the $SSR$ from an {\it
-  inverse-probability-weighted} least squares regression is more generally a
+outcomes. This suggests a conjecture: that the $SSR$ from an \emph{inverse-probability-weighted} least squares regression is more generally a
 sensible test statistic for models that include

### 2015-12-19 response.md
Small edits to the response to reviewers and editors
More progress on editing

@@
 
- Since the point of the paper is to learn about test statistics, so, in this
-paper we do not explore alternative models of causal effects. The original
+Since the point of the paper is to learn about test statistics,
+we do not explore alternative models of causal effects. The original
 paper (BFP) on this method compared the performance of tests when the data did
@@
 
-We have done this.
+When we set about adding the four unit example to the different sections of the
+paper, we realized that we were expanding the length of the research note.
 
@@
 
-We have added a section in which we replicate the analysis in Coppock's 2014 Journal of Experimental Political Science article "Information Spillovers: Another Look at Experimental Estimates of Legislator Responsiveness".
+We have added a section in which we replicate the analysis in Coppock's 2014
+*Journal of Experimental Political Science* article "Information Spillovers:
+Another Look at Experimental Estimates of Legislator Responsiveness".

### 2015-12-19 paper/introduction.tex
More progress on editing

@@
 power of a simple test of the sharp null hypothesis of no effects will vary as
-a function of the design of the study (proportion treated, blocking structure,
+a function of the design of the study (total number of observations, proportion treated, blocking structure,
 etc), characteristics of the outcome (continuous, binary, skewed, extreme
@@
 which treatment propagates, the idea of assessing model fit rather than
-closeness of distributions leads naturally to the the sum-of-squared-residuals
+closeness of distributions leads naturally to the sum-of-squared-residuals
 (SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
@@
 \begin{equation}
- \T(\yu,\bz)_{\text{SSR}} \equiv \sum_{i} ( \widetilde y_{i,Z_i=0} -
- \bX\hat{\bbeta}) ^2 \label{eq:ssr}
+ \T(\yu,\bz)_{\text{SSR}} \equiv \sum_{i} ( \widetilde y_{i,Z_i=0} - \bX\hat{\bbeta} )^2 \label{eq:ssr}
 \end{equation}
+
+To provide a little intuition about the SSR test statistic as compared to the
+KS test statistic,
+
+we created a small simulation study of 256 units with half
+assigned to a binary treatment completely at random. We compared two outcomes
+--- a Normal outcome and a skewed outcome, both with mean of 10 and standard
+deviation of 1. And we compared two models of effects --- a constant additive
+effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant
+multiplicative effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set
+the truth to be the sharp null of no effects $\tau=0$ for the additive model
+and $\tau=1$ for the multiplicative model. Our intuition was that the SSR
+would have more power than the KS test when the outcome was Normal and when
+the observational implication of the model of effects would shift means (i.e.
+the additive model). We used direct simulation of the randomization
+distribution to generate $p$-values and repeated that process 1000 times to
+gauge the proportion of rejections of false hypotheses (i.e. the power of the
+tests). The results, not shown here but part of the reproduction archive for
+this paper, bear out this intuition: the SSR has slightly more power than KS
+for Normal outcomes in both the additive and multiplicative effects
+conditions. SSR has slightly less power than KS when the outcome is skewed for
+both models. In general, the SSR ought to be most powerful when the effect of
+the experiment involves a shift in the location of the distributions of the
+treated versus the controls. And it should be less powerful when the effect of
+the experiment is mostly to leave the center of the distribution alone, but,
+instead to stretch or compress the tails, or to concentrate the experimental
+effect on one quantile.
+
+\subsection{The SSR Test Statistic and the BFP Example Model}
 
@@
 \citet{coppock2014information} presents an reanalysis of an experiment performed
-by \citet{butlernickerson2011}. Leading up to a key vote on a budget bill, SB24, in the
-New Mexico state senate, legislators were randomly assigned to see
-constituent survey results, and their votes on SB24 recorded. The original
+by \citet{butler2011can}. Leading up to a key vote on a budget bill, SB24, in the
+New Mexico state senate, Butler and Nickerson randomly assigned legislators to see
+constituent survey results. The original
 analysis found that legislators assigned to receive constituent information were
@@
 Given the easy nature of sharing information in
-field experiments \citep{Winters:2012fk}, Coppock reanalyzes the
+field experiments \citep{winters2012using},  Coppock reanalyzes the
 experimental results with an eye towards spillover, constructing a social
-network based on the similarity of W-NOMINATE ideology scores. Coppock hypothesized
+network based on similarity on W-NOMINATE ideology scores. Coppock hypothesized
 that ideologically similar legislators would be more likely to share
@@
 treated neighbors. Coppock evaluated this spillover model using the SSR+Design statistic as
-presented in this paper. To demonstrate the value of this statistic,
-particularly for models that are linear in the spillover effects, we repeat
-Coppock's analysis using both the SSR+Design statistic and a KS statistic. The
+presented in this paper.
+
+\begin{figure}[H] \centering
+  \includegraphics[width=.75\textwidth]{../coppock-replication/CoppockJEPS_figure2.pdf}
+  \caption{Replication of Figure 2 from \citet{coppock2014information}: plotting
+    $p$-values for direct and indirect model parameters. The
+    original analysis used the SSR statistic, which nicely bounds the region of
+    plausible hypotheses. The KS statistic, while eliminating more hypotheses in
+  the neighborhood of (0, 0), fails to provide any such bounds.}\label{fig:coppock-replication}
+\end{figure}
+
+To demonstrate the performance of this statistic in
+an applied setting, such as this one in which  the model is linear in the spillover effects, we repeat
+Coppock's analysis using both the SSR+Design statistic and the simple KS statistic. The
 results in Figure~\ref{fig:coppock-replication} recreate Coppock's Figure
-2.\footnote{During the replication process, we noticed an inconsistency between
-  our results using the SSR+Design statistic and Coppock's published figure. In
-  coordination with the author, we discovered that the published figure suffered
-  from on a programming error. The figure in this document reflects a correction
-  to that error, so it will not precisely match Coppock's original publication.}
-The KS test statistic, while being more discerning in the neighborhood of (0,
+2.\footnote{We are grateful to Alex Coppock for his help with both code and
+  data. Figure~\ref{fig:coppock-replication} does not precisely match the
+  figure in the original publication because we corrected a small
+  programming error during replication.}
+ The KS test statistic, while being more powerful in the neighborhood of (0,
 0), fails to bound the region of plausible hypotheses in the manner of the
@@
 
-\begin{figure}[h!] \centering
-  \includegraphics[width=.99\textwidth]{../coppock-replication/CoppockJEPS_figure2.pdf}
-  \caption{Replication of Figure 2 from \citet{coppock2014information}: plotting
-    $p$-values for direct and indirect model parameters. The
-    original analysis used the SSR statistic, which nicely bounds the region of
-    plausible hypotheses. The KS statistic, while eliminating more hypotheses in
-  the neighborhood of (0, 0), fails to provide any such bounds.}
-  \label{fig:coppock-replication}
-\end{figure}
-
 \section{Discussion and Speculations}
@@
 against relevant alternatives for all possible models of treatment effect
-propagation, network topologies and designs. However, we hope that this
-research note both improves the application of the BFP approach and raises new
+propagation, network topologies and designs: this paper uses two models of
+effects each of which was applied to a different network topology and
+experimental design.\footnote{The BFP paper itself engages with some questions about
+the performance of this approach when the theoretical model is very different
+from the process generating the data, and we encourage readers to see that
+discussion in their \S~5.2.} However, we hope that the results from the 
+examples presented here improves the application of the BFP approach and raises new
 questions for research.  BFP are correct in the assertion that, regardless of
@@
 
-In the example above, the SSR+Design test statistic had much better power than
+In the example shown in Figure~\ref{fig:twoD}, the SSR+Design test statistic had much better power than
 any other test statistic. But SSR from an ordinary least squares regression is
@@
 model predictions are of concern, a weighted variant of the Brownian distance
-covariance \cite{szekely2009brownian} or other {\it E}-statistic may be more
-sensible than SSR.
+covariance \citep{szekely2009brownian} or other $E$-statistic may be more
+sensible than the sum of squared residuals.

### 2015-12-19 paper/paper.tex
More progress on editing

@@
 %\titlespacing*{ ⟨command⟩}{⟨left⟩}{⟨beforesep⟩}{⟨aftersep⟩}[⟨right⟩]
-\titlespacing{\section}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}
-\titlespacing{\subsection}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}
-\titlespacing{\subsubsection}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}
+\titlespacing{\section}{0pc}{1.5ex plus 0.1ex minus 0.2ex}{.5ex plus 0.1ex minus 0.1ex}
+\titlespacing{\subsection}{0pc}{1.5ex plus 0.1ex minus 0.2ex}{.5ex plus 0.1ex minus 0.1ex}
+\titlespacing{\subsubsection}{0pc}{1.5ex plus 0.1ex minus 0.2ex}{.5ex plus 0.1ex minus 0.1ex}
 
-\newif\ifanon
+%%% Links and Table of Contents
+\usepackage[dvipsnames]{xcolor}
+\usepackage[colorlinks=TRUE,linkcolor=BrickRed,citecolor=MidnightBlue]{hyperref}
 
+% \newif\ifanon
 % \ifanon

### 2015-12-21 paper/introduction.tex
Working on finishing the simplified discussion part

@@
 
-Rosenbaum style randomization inference tends to use test statistics that
+Fisher/Rosenbaum style randomization inference tends to use test statistics that
 compare two distributions. Simple models imply that the distribution of the
@@
 of assessing model fit, and there are usually better ways to evaluate the fit
-of a model than comparing two marginal distributions. In the case where we know
-the fixed adjacency matrix of the network, $\bS$, and where we imagine that
-network attributes (like degree) of a node play a role in the mechanism by
-which treatment propagates, the idea of assessing model fit rather than
-closeness of distributions leads naturally to the sum-of-squared-residuals
-(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
-$\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned treatment)
-as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node). If we
-collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix $\bX$,
-and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
-coefficients $\bbeta$ when we could define the test statistic as:
+of a model than comparing two marginal distributions: For example, the KS test
+  uses the maximum difference in the empirical cumulative distributions of
+  each treatment group calculated without regard for the relationship between
+the treated and control distributions, thereby ignoring information that could
+increase the precision of the test. The simplest version of the SSR test
+statistic merely sums the difference between the mean of the outcome implied
+by the hypothesis and individual outcomes:
 
 \begin{equation}
- \T(\yu,\bz)_{\text{SSR}} \equiv \sum_{i} ( \widetilde y_{i,Z_i=0} - \bX\hat{\bbeta} )^2 \label{eq:ssr}
+ {\T(\yu,\bz)}_{\text{SSR}} \equiv \sum_{i} {( \widetilde y_{i,Z_i=0} - \bar{\widetilde{y}}_{i,Z_i=0} )}^2 \label{eq:ssr1}
 \end{equation}
@@
 To provide a little intuition about the SSR test statistic as compared to the
-KS test statistic,
+KS test statistic, we created a small simulation study of 256 units with half
+assigned to a binary treatment completely at random. We compared the
+performance of the two test statistics on two outcomes with no interference
+--- a Normal outcome and a skewed outcome, both with mean of 10 and standard
+deviation of 1 as shown in Figure~\ref{fig:simpleoutcomes}.
 
-we created a small simulation study of 256 units with half
-assigned to a binary treatment completely at random. We compared two outcomes
---- a Normal outcome and a skewed outcome, both with mean of 10 and standard
-deviation of 1. And we compared two models of effects --- a constant additive
+\begin{figure}[H]\centering
+  \includegraphics[width=.5\textwidth]{figures/ksvsssroutcomes.pdf}
+  \caption{}\label{fig:simpleoutcomes}
+\end{figure}
+
+And we compared two models of effects --- a constant additive
 effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant
@@
 effect on one quantile.
+
+In the case where we know
+the fixed adjacency matrix of the network, $\bS$, and where we imagine that
+network attributes (like degree) of a node play a role in the mechanism by
+which treatment propagates, the idea of assessing model fit rather than
+closeness of distributions leads naturally to the sum-of-squared-residuals
+(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
+$\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned treatment)
+as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node). If we
+collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix $\bX$,
+and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
+coefficients $\bbeta$ when we could define the test statistic as:
+
+\begin{equation}
+ \T(\yu,\bz)_{\text{SSR}} \equiv \sum_{i} ( \widetilde y_{i,Z_i=0} - \bX\hat{\bbeta} )^2 \label{eq:ssr}
+\end{equation}
 

### 2015-12-21 simulation/ksvsssrpow.R
Now saving more parts of the simulation

@@ -19,6 +19,10 @@ y0norm<-rnorm(n,mean=10)
 summary(y0zif)
 summary(y0norm)
 
+simpledat<-data.frame(y0zif=y0zif,y0norm=y0norm,Z=Zs)
+
+save(simpledat,file="simulation/simpledat.rda")
+
 ## Two models:
 ### One model is like a model of mean shift
 constant.additive.model

### 2015-12-22 code/plotting.R
Working on the 'intuition' of SSR versus KS part

@@ -3,95 +3,95 @@
 ################################################################################
 par(oma=rep(0,4))
 
-library(igraph)
-library(colorspace)
-
-plotexperiment <- function(s, z, mode = "undirected",
-                           vertex.shape = ifelse(z, "circle", "square"),
-                           layout = layout.fruchterman.reingold,
-                           vertex.size = 6,
-                           vertex.color = ifelse(z, "#252525", "#A0A0A0"),
-                           vertex.label = NA,
-                           edge.color = "black",
-                           ...) {
-
-  tmp <- graph.adjacency(s, mode = mode)
-
-  plot(tmp,
-       vertex.shape = vertex.shape,
-       layout = layout,
-       vertex.size = vertex.size,
-       vertex.color = vertex.color,
-       vertex.label = vertex.label,
-       edge.color = edge.color,
-       margin=0,
-       ...)
-}
-
-
-make.filled.contour.plot<-function(mat,main,levs=10,col=rev(terrain_hcl(levs)),
-                                   addTruth=TRUE,autolab=TRUE,
-                                   thelevels=NULL,
-                                   true.params, ##names and values of the parameters
-                                   thez="p",
-                                   ...){
-
-  params<-names(true.params)
-
-  mat.p<- matrix(mat[,thez],
-                 nrow=length(unique(mat[,params[1]])),
-                 ncol=length(unique(mat[,params[2]])),
-                 dimnames=list(sort(unique(mat[,params[1]])),
-                               sort(unique(mat[,params[2]]))
-                               )
-                 )
-  theps<-sort(unique(as.numeric(mat.p)))
-  p.line.levels<-quantile(theps,seq(0,1,length=levs) )
-  if(addTruth){
-    p.truth<-mat[simp.eq.test(mat[,params[1]],true.params[[1]])&simp.eq.test(mat[,params[2]],true.params[[2]]),thez]
-    ##main<-paste(main,"(",thealpha,"test level=",round(p.truth,3),")")
-  }
-  ##p.line.levels<-sort(unique(c(p.truth,p.line.levels)))
-  if(!is.null(thelevels)){
-    thelabels<-thelevels
-  }else{
-    if(autolab){
-      thelabels <- NULL
-      thelevels <- pretty(range(mat.p,finite=TRUE),levs)
[…]

### 2015-12-22 figures/ksvsssrpow-summary.R

@@ -1,11 +1,14 @@
-# A file to summarize the results of the KS versus SSR power comparison for no interference but difference models of effects and outcomes
+# A file to summarize the results of the KS versus SSR power comparison for no
+# interference but difference models of effects and outcomes
 
 if(length(grep("TestStatRIInterference$",getwd(),ignore.case=TRUE))==0){setwd("..")}
 if(length(grep("TestStatRIInterference",.libPaths(),ignore.case=TRUE))==0){.libPaths(".libraries") }
+options(width=132,digits=4)
 source("simulation/setup.R")
 load("simulation/ksvsssrpow-results.rda")
 load("simulation/simplealts.rda")
 load("simulation/simpletruth.rda")
+load("simulation/simpledat.rda")
 
 apply(powCOA,2,summary)
 apply(powCOM,2,summary)
@@ -26,4 +29,48 @@ matplot(x=ALTS$COM$tau,powCOM,type="l",lty=c(1,2,1,2),col=c(1,1,2,2),
 	main="Constant Multiplicative Model", xlab="Hypotheses (Truth=1)",ylab="pr(p<=.05)")
 legend(x="bottomleft",legend=substr(colnames(powCOM),11,50),lty=c(1,2,1,2),col=c(1,1,2,2),bty="n")
 dev.off()
+
+
+pdf(file="figures/ksvsssroutcomes.pdf",width=8,height=4)
+par(mfrow=c(1,2),pty="s",mgp=c(1.5,.5,0),oma=rep(0,4),mar=c(3,3,2,0))
+plot(density(simpledat$y0norm),main="Normal Outcome")
+rug(simpledat$y0norm)
+plot(density(simpledat$y0zif),main="Geometric Outcome")
+rug(simpledat$y0zif)
+dev.off()
+
+
+## Two models:
+### One model is like a model of mean shift
+constant.additive.model
+### One model focuses changes on the shape of the dist
+constant.multiplicative.model <- UniformityModel( function(y, z, tau) {
+						   if(tau==0){
+						     return(rep(0,length(y)))
+						   } else {
+						     y / ( 1 + z * ( tau-1 ) )
+						   }
+	}, function(y_0, z, tau) { y_0 * ( 1 + z * ( tau-1 ) )})
+
+
+source("code/plotting.R")
+
+par(mfrow=c(1,2))
+plotCompareModels(simpledat$y0norm,
+		  Z=simpledat$Z.1,
[…]

### 2015-12-22 simulation/ksvsssrpow.R

@@ -13,7 +13,7 @@ Zs <- sampler(REPETITIONS)$samples # N x nsims
 ## Make two outcomes in control that both have 0 mean and 1 sd but with different distributions
 set.seed(20151130)
 tmpzif<-rgeom(n,prob=.7)
-y0zif<-as.vector(scale(tmpzif))+10
+y0zif<-jitter(as.vector(scale(tmpzif))+10)
 y0norm<-rnorm(n,mean=10)
 
 summary(y0zif)

### 2015-12-23 code/plotting.R
Changes to the plotCompareModels function

@@ -5,7 +5,7 @@ par(oma=rep(0,4))
 
 ## library(igraph)
 ## library(colorspace)
-## 
+##
 ## plotexperiment <- function(s, z, mode = "undirected",
 ##                            vertex.shape = ifelse(z, "circle", "square"),
 ##                            layout = layout.fruchterman.reingold,
@@ -14,9 +14,9 @@ par(oma=rep(0,4))
 ##                            vertex.label = NA,
 ##                            edge.color = "black",
 ##                            ...) {
-## 
+##
 ##   tmp <- graph.adjacency(s, mode = mode)
-## 
+##
 ##   plot(tmp,
 ##        vertex.shape = vertex.shape,
 ##        layout = layout,
@@ -27,17 +27,17 @@ par(oma=rep(0,4))
 ##        margin=0,
 ##        ...)
 ## }
-## 
-## 
+##
+##
 ## make.filled.contour.plot<-function(mat,main,levs=10,col=rev(terrain_hcl(levs)),
 ##                                    addTruth=TRUE,autolab=TRUE,
 ##                                    thelevels=NULL,
 ##                                    true.params, ##names and values of the parameters
 ##                                    thez="p",
 ##                                    ...){
-## 
+##
 ##   params<-names(true.params)
-## 
+##
 ##   mat.p<- matrix(mat[,thez],
 ##                  nrow=length(unique(mat[,params[1]])),
 ##                  ncol=length(unique(mat[,params[2]])),
@@ -63,12 +63,12 @@ par(oma=rep(0,4))
 ##       thelevels <- sort(unique(c(0,p.line.levels)))
 ##     }
 ##   }
-## 
+##
 ##   parm1.values<-sort(unique(mat[,params[1]]))
 ##   parm2.values<-sort(unique(mat[,params[2]]))
-## 
+##
 ##   if(length(theps)==1){col<-col[length(col)]}
-## 
+##
 ##   image(list(x=parm1.values,y=parm2.values,z=mat.p),
 ##         main=main,
 ##         ##main=expression ( paste(main, "(Type I Error Rate", == round(p.truth,3), " for ",  alpha==thealpha,")") ),
@@ -87,11 +87,11 @@ par(oma=rep(0,4))
 ##             add=TRUE)
 ##   }
 ##   if(addTruth) {points(true.params[1],true.params[2],pch=19,cex=2)}
-## 
+##
[…]

### 2015-12-23 figures/ksvsssrpow-boxplot.R
Created a boxplot showing, in more concrete detail, the operation of the KS and SSR statistics

@@ -30,16 +30,6 @@ matplot(x=ALTS$COM$tau,powCOM,type="l",lty=c(1,2,1,2),col=c(1,1,2,2),
 legend(x="bottomleft",legend=substr(colnames(powCOM),11,50),lty=c(1,2,1,2),col=c(1,1,2,2),bty="n")
 dev.off()
 
-
-pdf(file="figures/ksvsssroutcomes.pdf",width=8,height=4)
-par(mfrow=c(1,2),pty="s",mgp=c(1.5,.5,0),oma=rep(0,4),mar=c(3,3,2,0))
-plot(density(simpledat$y0norm),main="Normal Outcome")
-rug(simpledat$y0norm)
-plot(density(simpledat$y0zif),main="Geometric Outcome")
-rug(simpledat$y0zif)
-dev.off()
-
-
 ## Two models:
 ### One model is like a model of mean shift
 constant.additive.model
@@ -55,22 +45,47 @@ constant.multiplicative.model <- UniformityModel( function(y, z, tau) {
 
 source("code/plotting.R")
 
-par(mfrow=c(1,2))
+pdf(file="figures/ksvsssr-boxplots.pdf",width=8,height=4)
+par(mfrow=c(1,2),pty="s",mgp=c(1.5,.5,0),oma=rep(0,4),mar=c(3,3,2,0))
 plotCompareModels(simpledat$y0norm,
 		  Z=simpledat$Z.1,
 		  make.data=function(y,Z){y},
-		  models=list(COA=givenParams(constant.additive.model,tau=ALTS$COA$tau[60]),
-			      COM=givenParams(constant.multiplicative.model,tau=ALTS$COM$tau[60])
-			      )
+		  models=list("Add. Model"=givenParams(constant.additive.model,tau=ALTS$COA$tau[60]),
+			      "Mult. Model"=givenParams(constant.multiplicative.model,tau=ALTS$COM$tau[60])
+			      ),
+		  main="Normal Outcome"
 		  )
+
+text(4.5,11.8,as.expression(bquote(tau[0]==.(signif(ALTS$COA$tau[60],2)))),cex=.8)
+text(6.5,11.8,as.expression(bquote(tau[0]==.(signif(ALTS$COM$tau[60],4)))),cex=.8)
+
+text(c(4.5,4.5,6.5,6.5),c(7.3,7.8,7.3,7.8),
[…]

### 2015-12-23 simulation/ksvsssrpow.R
Off loading setup to a setup file that multiple files can use. Also trying to rationalize file names a bit

@@ -10,18 +10,12 @@ source("code/teststatistics.R")
 sampler <- simpleRandomSampler(total = n, treated = n/2)
 Zs <- sampler(REPETITIONS)$samples # N x nsims
 
-## Make two outcomes in control that both have 0 mean and 1 sd but with different distributions
-set.seed(20151130)
-tmpzif<-rgeom(n,prob=.7)
-y0zif<-jitter(as.vector(scale(tmpzif))+10)
-y0norm<-rnorm(n,mean=10)
+load("simulation/simpledat.rda")
+y0zif<-simpledat$y0zif
+y0norm<-simpledat$y0norm
 
 summary(y0zif)
 summary(y0norm)
-
-simpledat<-data.frame(y0zif=y0zif,y0norm=y0norm,Z=Zs)
-
-save(simpledat,file="simulation/simpledat.rda")
 
 ## Two models:
 ### One model is like a model of mean shift
@@ -36,48 +30,10 @@ constant.multiplicative.model <- UniformityModel( function(y, z, tau) {
 						   }
 }, function(y_0, z, tau) { y_0 * ( 1 + z * ( tau-1 ) )})
 
-## Simulation settings
-### Different parameter settings can mean the same thing in the two different models
-TRUTHAdditive<-list(tau=0)
-TRUTHMultiplicative<-list(tau=1)
-stopifnot(all.equal(constant.additive.model(y0norm,Zs[,1],0),constant.multiplicative.model(y0norm,Zs[,1],1)))
-stopifnot(all.equal(constant.additive.model(y0zif,Zs[,1],0),constant.multiplicative.model(y0zif,Zs[,1],1)))
-
-## Try to make the hypotheses in the multiplicative model similar in magnitude
-## to the hypotheses of the multiplicative model (we cannot make them all
-## exactly the same across all values of y0, but we can make them as close as
-## possible)
-
-ALTSAdditive<-list(tau=sort(unique(c(0,1,seq(-2,2,length=100)))))
-
-addvsmult<-function(x,tau){
-  addres<-invertModel(constant.additive.model,y0norm,Zs[,1],tau=tau)
-  mulres<-invertModel(constant.multiplicative.model,y0norm,Zs[,1],tau=x*tau)
-  median(abs(mulres-addres))
-}
-
[…]

### 2015-12-23 simulation/ksvsssrpow-results.R
Indentation

@@ -7,19 +7,19 @@ library(RItools) ##,lib.loc=".libraries")
 
 source("simulation/setup.R")
 
-load("simulation/ksvsssrpowresults.rda")
+load("simulation/ksvsssrpow.rda")
 load("simulation/simplealts.rda")
 load("simulation/simpletruth.rda")
 
 
 simPow2<-function(lst){
-	m <- sapply(lst,function(l){ l$p})
-	m <- m <= ALPHA
-	rowMeans(m)
+  m <- sapply(lst,function(l){ l$p})
+  m <- m <= ALPHA
+  rowMeans(m)
 }
 
 powComparison<-sapply(ls(patt="resultsCO"),function(nm){
-simPow2(get(nm))
+			simPow2(get(nm))
 })
 
 powCOA<-powComparison[,grep("COA",colnames(powComparison))]

### 2015-12-23 response.md
Added section saying that we were not going to use the 4 units example throughout.

@@
 When we set about adding the four unit example to the different sections of the
-paper, we realized that we were expanding the length of the research note.
+paper, we realized that we were not enhancing intuition: the two test
+statistics applied to an experiment of size 4 mainly reflected the fact that we
+had very little data. Instead, we added a discussion of the SSR and KS test
+statistics applied to a case with no interference and with very simple models
+of effect. We also added a figure and discussion that tries to show a bit more
+about how models of effects work in the context of sharp null hypothesis
+testing. We hope that this helps with the intuition so that the use of the SSR
+test statistic to an experiment on a social network with a complex model of
+treatment propagation is easier to understand.
 

### 2015-12-23 paper/introduction.tex
Working on the 'intuition' of SSR versus KS part
Added section describing the simple SSR versus KS comparison

@@
 \begin{figure}[H]\centering
-  \includegraphics[width=.5\textwidth]{figures/ksvsssroutcomes.pdf}
-  \caption{}\label{fig:simpleoutcomes}
+  \includegraphics[width=.5\textwidth]{ksvsssr-outcomes.pdf}
+  \caption{Two outcomes ($Y_i$) simulated with no interference, $N=256$, half
+    randomized to binary treatment by complete randomization. }\label{fig:simpleoutcomes}
 \end{figure}
 
-And we compared two models of effects --- a constant additive
-effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant
-multiplicative effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set
-the truth to be the sharp null of no effects $\tau=0$ for the additive model
-and $\tau=1$ for the multiplicative model. Our intuition was that the SSR
+And we compared two models of effects --- a constant additive effects model in
+which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant multiplicative effects
+model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set the truth to be the
+sharp null of no effects $\tau=0$ for the additive model and $\tau=1$ for the
+multiplicative model. To further explain the process of hypothesis testing and
+evaluation of the test statistics, we display the results from a power
+analysis of one alternative hypothesis in figure~\ref{fig:boxplot}. A given
+model and hypothesized parameter implies a distribution for observed outcomes
+in the treatment group and control groups. Here, since we have no
+interference, the hypotheses only have different implications for the control
+group distribution. In this case, we chose a two parameters for the additive
+and multiplicative models which produced very similar implications (as show by
+the similarity in the ``Add. Model'' and ``Mult.  Model'' boxplots in both
+panels of the figure). We see that when we tested the hypothesis of
+$\tau_0=.34$ from the additive model using the KS test statistic on Normal
+data, that we produced $p$-values lower than $\alpha=.05$ about 55\% of the
+time across 1000 simulations. That is, the KS test has a power of about .55
+to reject a false hypothesis with this Normal outcome and this constant
+additive effects model. The analogous power for the SSR test statistic was
+.73. For these particular parameters, we see the SSR performing better with
+Normal outcomes but not being very effected by the model of effects (which
+makes some sense here because we are choosing parameters at which both models
+imply more or less the same patterns in outcomes).
+
+\begin{figure}[H]\centering
+  \includegraphics[width=.5\textwidth]{ksvsssr-boxplot.pdf}
+  \caption{Distributions of the simulated data from left to right: $y_0$
+    outcome with no experiment, observed outcome after random assignment,
+    outcomes implied by a constant additive model of effects, outcomes implied by a constant
+    multiplicative model. The hypothesized model parameters $\tau_0$ that
+    produce the patterns shown are printed on the plot. The proportion of
+    simulated tests in which these values of $\tau_0$ with $p$-values less
+    than $\alpha=.05$ by the KS and SSR test statistics are printed on the plot as $\text{pow}_{KS}$ and
+  $\text{pow}_\text{SSR}$. The SSR test statistic has more power for the
+  Normal outcomes and less power for the skewed outcome.}\label{fig:boxplot}
+\end{figure}
+
+When we assessed power across many alternative hypotheses, our intuition was that the SSR
 would have more power than the KS test when the outcome was Normal and when
-the observational implication of the model of effects would shift means (i.e.
+the observational implication of the model of effects would shift means (i.e.\
 the additive model). We used direct simulation of the randomization
 distribution to generate $p$-values and repeated that process 1000 times to
-gauge the proportion of rejections of false hypotheses (i.e. the power of the
+gauge the proportion of rejections of false hypotheses (i.e.\ the power of the
 tests). The results, not shown here but part of the reproduction archive for
@@
 
-In the case where we know
-the fixed adjacency matrix of the network, $\bS$, and where we imagine that
-network attributes (like degree) of a node play a role in the mechanism by
-which treatment propagates, the idea of assessing model fit rather than
-closeness of distributions leads naturally to the sum-of-squared-residuals
-(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
-$\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned treatment)
-as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node). If we
-collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix $\bX$,
-and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
-coefficients $\bbeta$ when we could define the test statistic as:
+\subsection{The SSR Test Statistic with Network Information}
+
+In the case where we know the fixed adjacency matrix of a social network,
+$\bS$, and where we imagine that network attributes (like degree) of a node
+play a role in the mechanism by which treatment propagates, the idea of
+assessing model fit rather than closeness of distributions leads naturally to
+the sum-of-squared-residuals (SSR) from a least squares regression of $
+\widetilde y_{i,Z_i=0}$ on $Z_{i}$ and $\bz^{T} \bS$ (i.e.\ the number of
+directly connected nodes assigned treatment) as well as the $\mathbf{1}^{T}
+\bS$ (i.e.\ the degree of the node). If we collected $Z_{i}$,$\bz^{T} \bS$,
+and $\mathbf{1}^{T} \bS$ into a matrix $\bX$, and fit the $\widetilde
+y_{i,Z_i=0}$ as a linear function of $\bX$ with coefficients $\bbeta$ when we
+could define the test statistic as:
 

### 2015-12-23 simulation/ksvsssr-setup.R
File with setup for the ksvsssr stuff

@@ -1,0 +1,78 @@
+
+if(length(grep("TestStatRIInterference$",getwd(),ignore.case=TRUE))==0){setwd("..")}
+if(length(grep("TestStatRIInterference",.libPaths(),ignore.case=TRUE))==0){.libPaths(".libraries") }
+
+library(RItools)
+source("simulation/setup.R")
+source("code/teststatistics.R")
+
+sampler <- simpleRandomSampler(total = n, treated = n/2)
+Zs <- sampler(REPETITIONS)$samples # N x nsims
+
+## Make two outcomes in control that both have 0 mean and 1 sd but with different distributions
+set.seed(20151130)
+tmpzif<-rgeom(n,prob=.7)
+y0zif<-jitter(as.vector(scale(tmpzif))+10)
+y0norm<-rnorm(n,mean=10)
+
+summary(y0zif)
+summary(y0norm)
+
+simpledat<-data.frame(y0zif=y0zif,y0norm=y0norm,Z=Zs)
+
+save(simpledat,file="simulation/simpledat.rda")
+
+## Two models:
+### One model is like a model of mean shift
+constant.additive.model
+
+### One model focuses changes on the shape of the dist
+constant.multiplicative.model <- UniformityModel( function(y, z, tau) {
+						   if(tau==0){
+						     return(rep(0,length(y)))
+						   } else {
+						     y / ( 1 + z * ( tau-1 ) )
+						   }
+}, function(y_0, z, tau) { y_0 * ( 1 + z * ( tau-1 ) )})
+
+## Simulation settings
+### Different parameter settings can mean the same thing in the two different models
+TRUTHAdditive<-list(tau=0)
+TRUTHMultiplicative<-list(tau=1)
+stopifnot(all.equal(constant.additive.model(y0norm,Zs[,1],0),constant.multiplicative.model(y0norm,Zs[,1],1)))
+stopifnot(all.equal(constant.additive.model(y0zif,Zs[,1],0),constant.multiplicative.model(y0zif,Zs[,1],1)))
+
+## Try to make the hypotheses in the multiplicative model similar in magnitude
+## to the hypotheses of the multiplicative model (we cannot make them all
+## exactly the same across all values of y0, but we can make them as close as
+## possible)
+
[…]

### 2015-12-23 figures/ksvsssr-outcomes.R
File to plot the distributions of the outcomes in the simple comparison

@@ -1,0 +1,11 @@
+load("simulation/simpledat.rda")
+
+pdf(file="figures/ksvsssr-outcomes.pdf",width=8,height=4)
+par(mfrow=c(1,2),pty="s",mgp=c(1.5,.5,0),oma=rep(0,4),mar=c(3,3,2,0))
+plot(density(simpledat$y0norm),main="Normal Outcome")
+rug(simpledat$y0norm)
+plot(density(simpledat$y0zif),main="Geometric Outcome")
+rug(simpledat$y0zif)
+dev.off()
+
+

### 2015-12-28 response.md
Drafted response to reviewers

@@
 Another Look at Experimental Estimates of Legislator Responsiveness".
+
+# Response to Reviewer 1
+
+We appreciate the positive review.
+In response to this review and the comments from the editor and other reviewer, we now valuate the SSR test statistic for three non-interference models (the sharp null model, the constant additive effects model, and the constant multiplicative effects model) and one new interference model (the one used by Coppock in his *Journal of Experimental Political Science* article).
+We did not systematically search for other spillover models from political science, but agree that such a survey of models and now, test statistics, would be useful in the future.
+
+# Response to Reviewer 2
+
+We appreciate the positive review.
+We added a section in an attempt to provide more intuition about test statistics, models of effects, and their interaction.
+Continuing the 4 unit example did not, in the end, clarify Fisher/Rosenbaum-style hypothesis testing as we would have hoped so we only use that example in the beginning to illustrate the idea that each unit in a study may have many more than two potential outcomes.
+We did add a section showing a case where the KS test has more power than the SSR test -- in this case we used a skewed outcome with no spillover rather than heterogeneous exposure to spillover in order to keep things as simple as possible.

### 2015-12-29 response.md
Update response.md

@@
 In the paper we clarified that we are focusing on testing sharp null
-hypotheses arising from one toy model of how an experimental treatment might
+hypotheses arising from one model of how an experimental treatment might
 propagate across a social network. We have added some discussion to the paper
@@
 We appreciate the positive review.
-In response to this review and the comments from the editor and other reviewer, we now valuate the SSR test statistic for three non-interference models (the sharp null model, the constant additive effects model, and the constant multiplicative effects model) and one new interference model (the one used by Coppock in his *Journal of Experimental Political Science* article).
+In response to this review and the comments from the editor and other reviewer, we now evaluate the SSR test statistic for three non-interference models (the sharp null model, the constant additive effects model, and the constant multiplicative effects model) and one new interference model (the one used by Coppock in his *Journal of Experimental Political Science* article).
 We did not systematically search for other spillover models from political science, but agree that such a survey of models and now, test statistics, would be useful in the future.

### 2016-01-04 response.md
Final response to editors for R1

@@
 
-We really appreciate the constructive and timely feedback from you and the reviewers.
+We really appreciate the constructive and timely feedback from you and the
+reviewers. Here is our response to your questions and suggestions and those of
+the reviewers.
+
+# Overall
 
@@
 
-Since the point of the paper is to learn about test statistics,
-we do not explore alternative models of causal effects. The original
-paper (BFP) on this method compared the performance of tests when the data did
-and did not arise from the causal model and we point to that paper for those
-who would like to pursue those set of questions. We did add some discussion
-about the power of the SSR test statistic against different kinds of causal
-effects: the SSR test statistic ought to be particularly sensitive to changes
-in mean differences (after all, recall that the coefficient on a dichotomous
-variable in an least squares model with a continuous outcome that minimizes
-the SSR is the difference of means) and ought to be less sensitive to other
-differences.
+Since the point of the paper is to learn about test statistics, we do not systematically
+explore alternative models of the propagation of causal effects across
+networks. The original paper (BFP) on this method compared the performance of
+tests when the data did and did not arise from the causal model and we point to
+that paper for those who would like to pursue those set of questions. We did
+add some discussion about the power of the SSR test statistic against different
+kinds of simple causal effects and outcome distributions: the SSR test
+statistic ought to be particularly sensitive to changes in mean differences
+(after all, recall that the coefficient on a dichotomous variable in an least
+squares model with a continuous outcome that minimizes the SSR is the
+difference of means) and ought to be less sensitive to other differences than a
+statistic like the KS test statistic.
 
@@
 statistics applied to a case with no interference and with very simple models
-of effect. We also added a figure and discussion that tries to show a bit more
+of causal effect. We also added a figure and discussion that tries to show a bit more
 about how models of effects work in the context of sharp null hypothesis
 testing. We hope that this helps with the intuition so that the use of the SSR
-test statistic to an experiment on a social network with a complex model of
+test statistic in an experiment on a social network with a complex model of
 treatment propagation is easier to understand.
@@
 
-We appreciate the positive review.
-In response to this review and the comments from the editor and other reviewer, we now evaluate the SSR test statistic for three non-interference models (the sharp null model, the constant additive effects model, and the constant multiplicative effects model) and one new interference model (the one used by Coppock in his *Journal of Experimental Political Science* article).
-We did not systematically search for other spillover models from political science, but agree that such a survey of models and now, test statistics, would be useful in the future.
+We appreciate the positive review. In response to this review and the comments
+from the editor and other reviewer, we now evaluate the SSR test statistic for
+three non-interference models (the sharp null model, the constant additive
+effects model, and the constant multiplicative effects model) and one new
+interference model (the one used by Coppock in his *Journal of Experimental
+Political Science* article).  We did not systematically search for other
+spillover models from political science, but agree that such a survey of
+models and now, test statistics, would be useful in the future.
 
@@
 
-We appreciate the positive review.
-We added a section in an attempt to provide more intuition about test statistics, models of effects, and their interaction.
-Continuing the 4 unit example did not, in the end, clarify Fisher/Rosenbaum-style hypothesis testing as we would have hoped so we only use that example in the beginning to illustrate the idea that each unit in a study may have many more than two potential outcomes.
-We did add a section showing a case where the KS test has more power than the SSR test -- in this case we used a skewed outcome with no spillover rather than heterogeneous exposure to spillover in order to keep things as simple as possible.
+We appreciate the positive review. We added a section in an attempt to provide
+more intuition about test statistics, models of effects, and their
+interaction. Continuing the 4 unit example did not, in the end, clarify
+Fisher/Rosenbaum-style hypothesis testing as we would have hoped so we only
+use that example in the beginning to illustrate the idea that each unit in a
+study may have many more than two potential outcomes. We did add a section
+showing a case where the KS test has more power than the SSR test -- in this
+case we used a skewed outcome with no spillover rather than heterogeneous
+exposure to spillover in order to keep things as simple as possible.

### 2016-01-04 paper/introduction.tex
Fixing things to build the whole paper on keeling
Fixing up paper.
Edits pre resubmission

@@
 treatment is assigned to the \emph{whole} network. When the treatment
-assignment vector,$\bz$, provides treatment to persons 2 and 3,
+assignment vector, $\bz$, provides treatment to persons 2 and 3,
 $\bz=\{0,1,1,0\}$, person $i=1$ might respond one way,
@@
 points, etc), and the way that a test statistic summarizes the outcome (does
-it compare means, standard deviations, medians, qqplots, etc). In general,
+it compare means, standard deviations, medians, qqplots, etc).\footnote{Some
+  use the term `effective sample size' --- which we first saw in
+  \citet{kish65} ---
+  to highlight the fact that statistical power depends on more than the number
+  of rows in a given rectangular dataset. } In general,
 test statistics should be powerful against relevant alternatives. \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
@@
 unbiased test of the hypothesis of no effects or positive effects when the
-positive effects involve one parameter. Notice that this means that, a test of
+positive effects involve one parameter. Such results mean that a test of
 the sharp null of no effects
@@
 test-statistic to be valid test (in that the test should produce $p$-values
-less than $\alpha$ no more than 100$\alpha$ of the time when the null is true) even though different
+less than $\alpha$ no more than 100$\alpha \%$ of the time when the null is true) even though different
 test statistics may imply different power against false hypotheses.  Yet, when the models are complex and
@@
 case, and, if $Y_i$ is Normal or at least unimodal without major outliers,
-then this test might have optimal power. The complex model used as an example
+then this test using means as test statistics might have optimal power. The complex model used as an example
 by BFP involved adjustments to both control and treated outcomes --- some
@@
 statistic merely sums the difference between the mean of the outcome implied
-by the hypothesis and individual outcomes:
+by the hypothesis and individual outcomes thereby including the correlation
+between treated and control outcomes as a part of the distribution of the
+statistic:
 
@@
 
-And we compared two models of effects --- a constant additive effects model in
+We compared two models of effects --- a constant additive effects model in
 which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant multiplicative effects
-model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set the truth to be the
-sharp null of no effects $\tau=0$ for the additive model and $\tau=1$ for the
-multiplicative model. To further explain the process of hypothesis testing and
-evaluation of the test statistics, we display the results from a power
-analysis of one alternative hypothesis in figure~\ref{fig:boxplot}. A given
-model and hypothesized parameter implies a distribution for observed outcomes
-in the treatment group and control groups. Here, since we have no
-interference, the hypotheses only have different implications for the control
-group distribution. In this case, we chose a two parameters for the additive
-and multiplicative models which produced very similar implications (as show by
-the similarity in the ``Add. Model'' and ``Mult.  Model'' boxplots in both
-panels of the figure). We see that when we tested the hypothesis of
-$\tau_0=.34$ from the additive model using the KS test statistic on Normal
-data, that we produced $p$-values lower than $\alpha=.05$ about 55\% of the
-time across 1000 simulations. That is, the KS test has a power of about .55
-to reject a false hypothesis with this Normal outcome and this constant
-additive effects model. The analogous power for the SSR test statistic was
-.73. For these particular parameters, we see the SSR performing better with
-Normal outcomes but not being very effected by the model of effects (which
-makes some sense here because we are choosing parameters at which both models
-imply more or less the same patterns in outcomes).
+model in which $y_{i,Z_i=1}=y_{i,Z_i=0} \cdot \tau$. We set the truth to be
+the sharp null of no effects such that the true $\tau=0$ for the additive
+model and the true $\tau=1$ for the multiplicative model. To further explain
+the process of hypothesis testing and evaluation of the test statistics, we
+display the results from a power analysis of one alternative hypothesis in
+figure~\ref{fig:boxplot}. A given model and hypothesized parameter implies a
+distribution for observed outcomes in the treatment group and control groups.
+Here, since we have no interference, the hypotheses only have different
+implications for the control group distribution. In this case, we chose a two
+parameters for the additive and multiplicative models which produced very
+similar implications (as show by the similarity in the ``Add. Model'' and
+``Mult.  Model'' boxplots in both panels of the figure). We see that when we
+tested the hypothesis of $\tau_0=0.34$ from the additive model using the KS
+test statistic on Normal data, that we produced $p$-values lower than
+$\alpha=0.05$ about 55\% of the time across 1000 simulations. That is, the KS
+test has a power of about 0.55 to reject a false hypothesis with this Normal
+outcome and this constant additive effects model. The analogous power for the
+SSR test statistic was 0.73. For these particular parameters, we see the SSR
+performing better with Normal outcomes but not being very effected by the
+model of effects (which makes some sense here because we are choosing
+parameters at which both models imply more or less the same patterns in
+outcomes). On non-Normal outcomes (shown in the "Jittered Geometric Outcomes"
+panel), the SSR test statistic had less power. Again, the particular model of
+effect did not matter in this case because we chose the alternative hypothesis
+($\tau_0$ on the plots) to represent the case where both models implied
+similar outcome distributions.
 
 \begin{figure}[H]\centering
-  \includegraphics[width=.5\textwidth]{ksvsssr-boxplot.pdf}
+  \includegraphics[width=.9\textwidth]{ksvsssr-boxplots.pdf}
   \caption{Distributions of the simulated data from left to right: $y_0$
     outcome with no experiment, observed outcome after random assignment,
-    outcomes implied by a constant additive model of effects, outcomes implied by a constant
-    multiplicative model. The hypothesized model parameters $\tau_0$ that
+    outcomes implied by a constant additive model of effects ("Add. Model"), outcomes implied by a constant
+    multiplicative model ("Mult. Model"). The hypothesized model parameters $\tau_0$ that
     produce the patterns shown are printed on the plot. The proportion of
@@
 
-When we assessed power across many alternative hypotheses, our intuition was that the SSR
-would have more power than the KS test when the outcome was Normal and when
-the observational implication of the model of effects would shift means (i.e.\
-the additive model). We used direct simulation of the randomization
-distribution to generate $p$-values and repeated that process 1000 times to
-gauge the proportion of rejections of false hypotheses (i.e.\ the power of the
-tests). The results, not shown here but part of the reproduction archive for
-this paper, bear out this intuition: the SSR has slightly more power than KS
-for Normal outcomes in both the additive and multiplicative effects
+When we assessed power across many alternative hypotheses, our intuition was
+that the SSR would have more power than the KS test when the outcome was
+Normal and when the observational implication of the model of effects would
+shift means (i.e.\ the additive model). We used direct simulation of the
+randomization distribution to generate $p$-values and repeated that process
+1000 times to gauge the proportion of rejections of a range of false
+hypotheses  (i.e.\ the power of the tests at many different values of
+$\tau_0$). The results, not shown here but part of the reproduction archive
+for this paper, bear out this intuition: the SSR has slightly more power than
+KS for Normal outcomes in both the additive and multiplicative effects
 conditions. SSR has slightly less power than KS when the outcome is skewed for
@@
 As an example of the performance of these new statistics, we re-analyze the
-model and design from BFP. Their model of treatment propagation
-was:
+model and design from BFP. Their model of treatment propagation was:
 

### 2016-03-25 simulation/teststat-parallel.R
Small typos

@@ -4,7 +4,7 @@ if(length(grep("TestStatRIInterference$",getwd()))==0){setwd("..")}
 require(compiler)
 enableJIT(3)
 
-library(RItools,lib.loc=".libraries")
+library(RItools)
 options("RItools-lapply"=lapply)
 options("RItools-sapply"=sapply)
 

### 2016-03-26 figures/ksvsssr-boxplots.R
Because figure 2 as hard to read

@@ -13,7 +13,7 @@ load("simulation/simpledat.rda")
 apply(powCOA,2,summary)
 apply(powCOM,2,summary)
 
-t(powCOA[powCOA[,"resultsCOANormKS"]<1,])
+t(powCOA[powCOA[,"resultsCOANormKS"]<2,])
 t(powCOM[powCOM[,"resultsCOMNormKS"]<1,])
 
 ## SSR is slightly more power than KS for COA with Normal outcomes
@@ -45,8 +45,8 @@ constant.multiplicative.model <- UniformityModel( function(y, z, tau) {
 
 source("code/plotting.R")
 
-pdf(file="figures/ksvsssr-boxplots.pdf",width=8,height=4)
-par(mfrow=c(1,2),pty="s",mgp=c(1.5,.5,0),oma=rep(0,4),mar=c(3,3,2,0))
+pdf(file="figures/ksvsssr-boxplots.pdf",width=10,height=4)
+par(mfrow=c(1,2),pty="m",mgp=c(1.5,.5,0),oma=rep(0,4),mar=c(6,3,2,.5))##,xpd=NA)
 plotCompareModels(simpledat$y0norm,
 		  Z=simpledat$Z.1,
 		  make.data=function(y,Z){y},
@@ -55,17 +55,14 @@ plotCompareModels(simpledat$y0norm,
 			      ),
 		  main="Normal Outcome"
 		  )
-
-text(4.5,11.8,as.expression(bquote(tau[0]==.(signif(ALTS$COA$tau[60],2)))),cex=.8)
-text(6.5,11.8,as.expression(bquote(tau[0]==.(signif(ALTS$COM$tau[60],4)))),cex=.8)
-
-text(c(4.5,4.5,6.5,6.5),c(7.3,7.8,7.3,7.8),
+text(4.5,5.5,as.expression(bquote(tau[0]==.(signif(ALTS$COA$tau[60],2)))),cex=.8,xpd=NA)
+text(6.5,5.5,as.expression(bquote(tau[0]==.(signif(ALTS$COM$tau[60],4)))),cex=.8,xpd=NA)
+text(c(4.5,4.5,6.5,6.5),c(4.5,5,4.5,5),
      sapply(c(bquote(pow[KS] == .(signif(powCOA[60,"resultsCOANormKS"],2))),
               bquote(pow[SSR] ==  .(signif(powCOA[60,"resultsCOANormSSR"],2))),
[…]

### 2016-03-26 paper/introduction.tex
Fixing caption for Figure 2

@@
 \begin{figure}[H]\centering
-  \includegraphics[width=.9\textwidth]{ksvsssr-boxplots.pdf}
-  \caption{Distributions of the simulated data from left to right: $y_0$
-    outcome with no experiment, observed outcome after random assignment,
-    outcomes implied by a constant additive model of effects ("Add. Model"), outcomes implied by a constant
-    multiplicative model ("Mult. Model"). The hypothesized model parameters $\tau_0$ that
-    produce the patterns shown are printed on the plot. The proportion of
-    simulated tests in which these values of $\tau_0$ with $p$-values less
-    than $\alpha=.05$ by the KS and SSR test statistics are printed on the plot as $\text{pow}_{KS}$ and
-  $\text{pow}_\text{SSR}$. The SSR test statistic has more power for the
-  Normal outcomes and less power for the skewed outcome.}\label{fig:boxplot}
+  \includegraphics[width=.99\textwidth]{ksvsssr-boxplots.pdf}
+  \caption{The SSR test statistic has more power than the KS test statistic for the Normal outcome and
+    less power for the skewed outcome. Each panel shows distributions of simulated data from left to right:
+    outcome with no experiment ($y_0$), observed outcome after random assignment,
+    outcomes implied by a constant additive model of effects (``Add. Model''),
+    and outcomes implied by a constant multiplicative model (``Mult. Model''). The hypothesized model parameters $\tau_0$ that
+    produce the patterns, and the power of the tests using the KS and SSR test
+    statistics, are printed below the models.}\label{fig:boxplot}
 \end{figure}

### 2016-03-26 paper/introduction.tex
Because one of the reviewers pointed out typos, this commit is the result of fixing some of them.

@@
 power of a simple test of the sharp null hypothesis of no effects will vary as
-a function of the design of the study (total number of observations, proportion treated, blocking structure,
-etc), characteristics of the outcome (continuous, binary, skewed, extreme
-points, etc), and the way that a test statistic summarizes the outcome (does
-it compare means, standard deviations, medians, qqplots, etc).\footnote{Some
+a function of the design of the study (e.g. total number of observations, proportion treated, blocking structure), characteristics of the outcome (e.g. continuous, binary, skewed, extreme
+points), and the way that a test statistic summarizes the outcome (does
+it compare means, standard deviations, medians, qqplots, or something else?).\footnote{Some
   use the term `effective sample size' --- which we first saw in
@@
 small when the treated and control distributions in the adjusted data \ldots
-are similar, and large when the distributions diverge.'' \cite[Proposition 4
+are similar, and large when the distributions diverge.'' \citet[Proposition 4
 and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
@@
 sometimes a KS-test will have (1) no power to address the model at all such
-that all hypothesized parameters would receive the same implausibility
-assessment; (2) or might reject all such hypothesized parameters. Thus,
+that all hypothesized parameters would receive the same high $p$-value; (2) or might describe all such hypothesized parameters as
+implausible. Thus,
 although in theory one may assess sharp multiparameter hypotheses,
@@
 compare two distributions. Simple models imply that the distribution of the
-outcome in the control remains fixed. For example, $\widetilde
-y_{i,Z_i=0}=Y_i-Z_i \tau$ only changes the distribution of outcomes for units
+outcome in the control remains fixed. For example, the implication of the
+constant, additive effects model, $\widetilde
+y_{i,Z_i=0}=Y_i-Z_i \tau$, only changes the distribution of outcomes for units
 in the treated condition. Comparing the mean of $\widetilde y_{i,Z_i=1}|Z_i=1$
@@
 
-Yet, one can also think about the process of hypothesis testing as a process
+Although thinking about test statistics as comparing distributions is natural
+and simple, one can also think about the process of hypothesis testing as a process
 of assessing model fit, and there are usually better ways to evaluate the fit
@@
   each treatment group calculated without regard for the relationship between
-the treated and control distributions, thereby ignoring information that could
+the treated and control distributions, thereby ignoring information about the
+joint distribution that could
 increase the precision of the test. The simplest version of the SSR test
@@
 test statistic on Normal data, that we produced $p$-values lower than
-$\alpha=0.05$ about 55\% of the time across 1000 simulations. That is, the KS
-test has a power of about 0.55 to reject a false hypothesis with this Normal
+$\alpha=0.05$ about 54\% of the time across 1000 simulations. That is, the KS
+test has a power of about 0.54 to reject a false hypothesis with this Normal
 outcome and this constant additive effects model. The analogous power for the
-SSR test statistic was 0.73. For these particular parameters, we see the SSR
-performing better with Normal outcomes but not being very effected by the
+SSR test statistic was 0.70. For these particular parameters, we see the
+performance of the SSR test statistic as better than the KS test statistic
+with Normal outcomes, but not very effected by the
 model of effects (which makes some sense here because we are choosing
 parameters at which both models imply more or less the same patterns in
-outcomes). On non-Normal outcomes (shown in the "Jittered Geometric Outcomes"
+outcomes). On non-Normal outcomes (shown in the ``Jittered Geometric Outcomes"
 panel), the SSR test statistic had less power. Again, the particular model of
@@
 
-When we assessed power across many alternative hypotheses, our intuition was
+When we assessed power across many alternative hypotheses in results not shown
+here but part of the reproduction archive
+for this paper, our intuition was
 that the SSR would have more power than the KS test when the outcome was
@@
 hypotheses  (i.e.\ the power of the tests at many different values of
-$\tau_0$). The results, not shown here but part of the reproduction archive
-for this paper, bear out this intuition: the SSR has slightly more power than
+$\tau_0$). The results bear out this intuition: the SSR has slightly more power than
 KS for Normal outcomes in both the additive and multiplicative effects
@@
 
-In the case where we know the fixed adjacency matrix of a social network,
-$\bS$, and where we imagine that network attributes (like degree) of a node
-play a role in the mechanism by which treatment propagates, the idea of
-assessing model fit rather than closeness of distributions leads naturally to
-the sum-of-squared-residuals (SSR) from a least squares regression of $
-\widetilde y_{i,Z_i=0}$ on $Z_{i}$ and $\bz^{T} \bS$ (i.e.\ the number of
-directly connected nodes assigned treatment) as well as the $\mathbf{1}^{T}
-\bS$ (i.e.\ the degree of the node). If we collected $Z_{i}$,$\bz^{T} \bS$,
-and $\mathbf{1}^{T} \bS$ into a matrix $\bX$, and fit the $\widetilde
-y_{i,Z_i=0}$ as a linear function of $\bX$ with coefficients $\bbeta$ when we
-could define the test statistic as:
+In the case where we know the fixed binary adjacency matrix of a social
+network, $\bS$, where a given entry in the $n \times n$ matrix is 1 if two
+units are adjacent and 0 if two units are not adjacent, and where we imagine
+that network attributes (like degree) of a node play a role in the mechanism
+by which treatment propagates, the idea of assessing model fit rather than
+closeness of distributions leads naturally to the sum-of-squared-residuals
+(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$
+and $\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned
+treatment) as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node).
+If we collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix
+$\bX$, and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
+coefficients $\bbeta$ when we could define the test statistic as:
 
@@
 connected in an undirected, random graph with node degree ranging from 0 to 10
-(mean degree 4, 95\% of nodes with degree between 1 and 8, five nodes with
-degree 0 [i.e.\ unconnected]).  Treatment is assigned to 50\% of the nodes 
+(mean degree 4, 95\% of nodes with degree between 1 and 8, five unconnected nodes with
+degree 0).  Treatment is assigned to 50\% of the nodes
 completely at random in the BFP example.
@@
 discussion in their \S~5.2.} However, we hope that the results from the
-examples presented here improves the application of the BFP approach and raises new
+examples presented here improve the application of the BFP approach and raise new
 questions for research.  BFP are correct in the assertion that, regardless of
@@
 used, and not necessarily reflect the plausibility of the model at hand. That
-is, the results of any hypothesis test (or confidence interval creation) tell
+is, the results of any hypothesis test (or confidence interval) tell
 us \emph{both} about the test statistic \emph{and} about the causal model under scrutiny.
@@
 probability weights so as to ensure representative samples of potential
-outcomes. This suggests a conjecture: that the $SSR$ from an \emph{inverse-probability-weighted} least squares regression is more generally a
+outcomes. This suggests a conjecture: that the $SSR$ from an
+\emph{inverse-probability-weighted} least squares regression is a more generally
 sensible test statistic for models that include

### 2016-06-14 paper/introduction.tex
Version sent to PA for production at first

@@
 here but part of the reproduction archive
-for this paper, our intuition was
+for this paper \citep{bowers2016dataverse}, our intuition was
 that the SSR would have more power than the KS test when the outcome was
@@
 hypotheses  (i.e.\ the power of the tests at many different values of
-$\tau_0$). The results bear out this intuition: the SSR has slightly more power than
+$\tau_0$).\footnote{Those interested in the code for this document can find it
+at \url{https://github.com/jwbowers/TestStatRIInterference}. } The results bear out this intuition: the SSR has slightly more power than
 KS for Normal outcomes in both the additive and multiplicative effects

### 2016-06-14 paper/paper.tex
Version sent to PA for production at first

@@
 
-\author{Jake Bowers \thanks{Associate Professor, Dept of Political Science and Statistics, University of Illinois @ Urbana-Champaign (\texttt{jwbowers@illinois.edu}).} \and Mark Fredrickson \thanks{Ph.D.
+\author{Jake Bowers \thanks{Associate Professor, Dept of Political Science and
+		Statistics, University of Illinois @ Urbana-Champaign
+		(\texttt{jwbowers@illinois.edu}). Data and code to reproduce this document
+	can be found at \citet{bowers2016dataverse}} \and Mark Fredrickson \thanks{Ph.D.
 Student in Statistics and Political Science, University of Illinois @ Urbana-Champaign.} \and Peter M.
